---
title: "Demonstrating `predict_evo` and `alfak` in alfakR"
author: "RJB"
output: rmarkdown::github_document
vignette: >
  %\VignetteIndexEntry{Demonstrating predict_evo and alfak in alfakR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# Ensure the outDir for alfak examples is cleaned up or unique if run multiple times
outDir_vignette <- tempfile(pattern = "alfakR_tmp_vignette_")
dir.create(outDir_vignette)
original_wd <- getwd()

```

## Introduction

This vignette demonstrates key functionalities of the `alfakR` package. We will primarily focus on:

1.  **`predict_evo`**: Simulating karyotype evolution using both Ordinary Differential Equation (ODE) models and Agent-Based Models (ABM). We will compare the `predict_evo` ODE output to a custom ODE implementation based on Kimmel et al. (2023) to serve as a ground truth.
2.  **`alfak`**: Recovering fitness estimates from karyotype count data.


## Demonstrating `predict_evo`

We will simulate the evolution of karyotype frequencies under different missegregation rates ($\beta$) and population ceilings ($N_{max}$).

### Setup: Parameters and Libraries

First, we define the simulation parameters and load necessary libraries.

```{r params_libs}
misseg_rates <- c(0, 0.005, 0.01)
abm_max_vec  <- c(1e4, 1e5)
replicates   <- 10 # Number of ABM replicates
sim_times    <- seq(0, 50, by = 1)

library(alfakR); library(deSolve); library(ggplot2); library(utils)
```

The following helper functions are used to construct a custom ODE model for comparison purposes, based on the model described in Kimmel, G. J., et al. (2023). *PLOS Computational Biology*, *19*(1), e1010815.

```{r helper_funcs}
pij <- function(i,j,beta){ if(abs(i-j)>i) return(0); if(j==0) j <- 2*i;
sum(choose(i, seq(abs(i-j), i, 2))*beta^(seq(abs(i-j), i, 2))*
    (1-beta)^(i-seq(abs(i-j), i, 2))*0.5^seq(abs(i-j), i, 2)*
    choose(seq(abs(i-j), i, 2), (seq(abs(i-j), i, 2)+i-j)/2)) }
get_A <- function(kary_list,g,beta) matrix(sapply(1:2, function(i)
  sapply(1:2, function(j){ q<-prod(sapply(1:length(kary_list[[1]]),
                                        function(k) pij(kary_list[[i]][k], kary_list[[j]][k], beta))); if(i==j)
                                          g[i]*(2*q-1) else g[i]*2*q })),2,2,byrow=TRUE)
chrmod <- function(t,state,parms) list(state%*%parms$A)
vec <- function(s) as.numeric(strsplit(s,"[.]")[[1]])
```

### Landscape and Initial State

We define a simple fitness landscape with two karyotypes and their initial frequencies.

```{r landscape_initial}
landscape_df <- data.frame(k=c("2.2.2.2","2.2.2.3"), mean=c(0.1,0.12))
init <- c(0.9,0.1); names(init) <- landscape_df$k
kary_list <- lapply(landscape_df$k, vec); g <- landscape_df$mean 
```

### Simulation Loop

We now loop through the defined missegregation rates and population ceilings, running simulations with `predict_evo` (ODE and ABM) and the custom ODE model.

```{r sim_loop, message=FALSE, warning=FALSE}
results <- list(); idx <- 1
for(beta_val in misseg_rates) for(max_pop_val in abm_max_vec){ 
  pop_size <- max_pop_val/10
  
  ode1 <- predict_evo(lscape=landscape_df,p=beta_val,times=sim_times,x0=init,
                      prediction_type="ODE")
  ode1_df <- data.frame(time=sim_times,freq=ode1[,2],method="predict_evo_ODE",
                        beta=beta_val,beta_lbl=paste0("β=",beta_val),
                        pop=max_pop_val,pop_lbl=paste0("Nmax=",max_pop_val),rep=max_pop_val)
  
  A <- get_A(kary_list,g,beta_val)
  ode2 <- deSolve::ode(y=init,func=chrmod,parms=list(A=A),times=sim_times)
  # Ensure the correct column is selected by name and then normalized
  ode2_df <- data.frame(time=sim_times,freq=ode2[,2]/rowSums(ode2[,2:3]),
                        method="custom_ODE",beta=beta_val,beta_lbl=paste0("β=",beta_val),
                        pop=max_pop_val,pop_lbl=paste0("Nmax=",max_pop_val),rep=max_pop_val)
  results[[idx]] <- rbind(ode1_df,ode2_df); idx <- idx+1
  
  for(r_val in 1:replicates){ 
    abm <- predict_evo(lscape=landscape_df,p=beta_val,times=sim_times,x0=init,
                       prediction_type="ABM",abm_pop_size=pop_size,
                       abm_delta_t=0.01,abm_max_pop=max_pop_val,abm_record_interval=100)
    results[[idx]] <- data.frame(time=sim_times,freq=unlist(abm[,2]),method="ABM",
                                 beta=beta_val,beta_lbl=paste0("β=",beta_val),
                                 pop=max_pop_val,pop_lbl=paste0("Nmax=",max_pop_val),
                                 rep=r_val) 
    idx <- idx+1 }
}
res <- do.call(rbind, results)
```

### Plotting Results

The combined results are plotted to visualize the frequency dynamics. The plot will show the frequency of the first karyotype (`r landscape_df$k[1]`).

```{r plot_predict_evo, fig.width=8, fig.height=6}
sc <- scale_linetype_manual(values=c(ABM="solid",predict_evo_ODE="dashed",
                                     custom_ODE="dotdash"))
ac <- scale_alpha_manual(values=c(ABM=0.3,predict_evo_ODE=1,custom_ODE=1))
ss <- scale_size_manual(values=c(ABM=0.3,predict_evo_ODE=0.8,custom_ODE=0.8))

ggplot(res,aes(time,freq,group=interaction(method,rep),color=method))+
  geom_line(aes(linetype=method, alpha=method, size=method))+
  facet_grid(beta_lbl~pop_lbl)+
  labs(x="Time",y=paste("Frequency of karyotype", landscape_df$k[1]))+
  theme_bw() + sc + ac + ss
```

### Interpretation of `predict_evo` Output

The figure above displays the frequency of karyotype `r landscape_df$k[1]` over time under various conditions.

-   The **`predict_evo_ODE`** (dashed lines) represents the deterministic predictions from `alfakR`.
-   The **`custom_ODE`** (dot-dashed lines) represents predictions from the ground truth model based on Kimmel et al. (2023).
-   The **`ABM`** (solid, semi-transparent lines) shows individual stochastic simulations from `predict_evo`.

As can be seen, the `predict_evo_ODE` results closely match the `custom_ODE` outputs, validating the ODE implementation within `predict_evo`. The ABM simulations show stochastic variation around these deterministic trajectories, as expected.

## Exemplar: Using `alfak` to Recover Fitness Estimates

The `alfak` function can be used to infer karyotype fitness values from time-series count data. Here, we simulate data using the custom ODE model and then attempt to recover the original fitness parameters.

### Data Simulation and Preparation

We first simulate karyotype dynamics using the custom ODE model. For this example, we will use $\beta = `r tail(misseg_rates,1)`$.

```{r alfak_data_prep1}
beta_for_alfak <- tail(misseg_rates,1) 
A_alfak <- get_A(kary_list,g,beta_for_alfak) 
x_sim <- deSolve::ode(y=init,func=chrmod,parms=list(A=A_alfak),times=sim_times)
```

We then sample from these ODE-predicted frequencies to generate plausible counts data, mimicking experimental observations. We select time points from 5 to 45, with a step of 5.

```{r alfak_counts_prep1}
time_indices_for_counts <- sim_times %in% seq(5,45,5)
# Ensure x_sim has a time column to exclude if it's part of the output from deSolve::ode
# The default output of ode includes time as the first column.
# We need to select states corresponding to landscape_df$k
# And ensure probabilities for rmultinom are non-negative and sum to 1.
x_states_all <- x_sim[, -1, drop=FALSE] # Exclude time column
x_sampled <- x_states_all[time_indices_for_counts, , drop=FALSE] 

# Ensure probabilities are correctly normalized for rmultinom
prob_func <- function(row_vec) {
  row_vec_norm <- pmax(0, row_vec) # Ensure non-negative
  if(sum(row_vec_norm) == 0) { # Handle case where all probs are zero (e.g. after pmax)
    return(rep(1/length(row_vec_norm), length(row_vec_norm))) # Assign equal probability
  }
  return(row_vec_norm / sum(row_vec_norm))
}

counts <- data.frame(apply(x_sampled,1,function(xi) rmultinom(n=1,size=200,prob=prob_func(xi)))) 
colnames(counts) <- seq(5,45,5) 
rownames(counts) <- landscape_df$k
yi <- list(x=counts,dt=1) 
```

### Running `alfak` (Initial Attempt)

Now, we run `alfak` on this simulated dataset. The `outdir` for `alfak` will create a subdirectory containing output within our  working directory.

```{r alfak_run1, message=FALSE, warning=FALSE}
alfak_output_subdir1 <- "alfakR_tmp_run1" 
console_out <- utils::capture.output(xv <- alfak(yi,outdir=alfak_output_subdir1,passage_times = as.numeric(colnames(yi$x)),minobs = 10,nboot=20,pm=beta_for_alfak,n0=1e5,nb=1e7))
print(head(console_out))
```

This is a toy example. The cross-validation score (`xv$score`) is likely low, and the kriging-fitted landscape is likely unreliable. During the execution of `alfak`, some warnings related to the kriging process (specifically, "Grid searches over lambda ...") may appear. These indicate that the optimization for variance components in the spatial smoothing model found its best solution at the edge of the parameter search space. This is not uncommon in examples with sparse data or simple underlying fitness landscapes, as is the case with this toy example. 

```{r alfak_results1_landscape}
landscape_rds_path1 <- file.path(alfak_output_subdir1, "landscape.Rds")
if(file.exists(landscape_rds_path1)){
  lscape_alfak1 <- readRDS(landscape_rds_path1)
  print("Kriging fitted landscape (Initial):")
  head(lscape_alfak1)
} else {
  print(paste("landscape.Rds not found in", landscape_rds_path1))
}
```

The relative stage 1 fitness estimates can still be informative.

```{r alfak_results1_fitness}
boot_res_rds_path1 <- file.path(alfak_output_subdir1, "bootstrap_res.Rds")
if(file.exists(boot_res_rds_path1)){
  boot_res1 <- readRDS(boot_res_rds_path1)
  print("Median inferred fitness (Stage 1 - Initial):")
  print(apply(boot_res1$final_fitness,2,median))
  print("Original fitness values:")
  print(landscape_df)
} else {
  print(paste("bootstrap_res.Rds not found in", boot_res_rds_path1))
}
```

Note that the absolute values of the stage 1 fitness estimates might not match the true values directly. `alfakR` uses passage times (`dt`) to calibrate absolute fitness.

### Calibrating Fitness Estimates with Realistic Passage Times

The growth rate of the dominant cell population can inform `dt`. If the population grows exponentially from $N_0$ to $N_B$ before passaging, and its growth rate is $m$, then $dt = \log(N_B/N_0)/m$.

```{r alfak_dt_calc}
mean_growth_rate <- landscape_df$mean[1] 
nb_val <- 1e6 
n0_val <- 1e5 
dt_realistic <- log(nb_val/n0_val)/mean_growth_rate
print(paste("Calculated dt:", dt_realistic))
```

Let's re-simulate the ODE model with these more "realistic" passage intervals.

```{r alfak_data_prep2}
A_alfak_realistic <- get_A(kary_list,g,beta_for_alfak) 
total_sim_time_realistic <- 5 * dt_realistic # Simulate for a few passages
passage_sim_times <- seq(0, total_sim_time_realistic, by=dt_realistic)
if(length(passage_sim_times) < 2) passage_sim_times <- c(0, dt_realistic) # Ensure at least one interval

x_sim_realistic <- deSolve::ode(y=init,func=chrmod,parms=list(A=A_alfak_realistic),times=passage_sim_times)
```

Prepare the data for `alfak` again.

```{r alfak_counts_prep2}
# Exclude initial state at t=0 and the time column
x_states_realistic_all <- x_sim_realistic[-1, -1, drop=FALSE] 
counts_realistic <- data.frame(apply(x_states_realistic_all,1,function(xi) rmultinom(n=1,size=200,prob=prob_func(xi))))
colnames(counts_realistic) <- 1:ncol(counts_realistic) 
rownames(counts_realistic) <- landscape_df$k
yi_realistic <- list(x=counts_realistic,dt=dt_realistic)
```

### Running `alfak` with Calibrated `dt`

```{r alfak_run2, message=FALSE, warning=FALSE}
alfak_output_subdir2 <- "alfakR_tmp_run2"
console_out <- utils::capture.output(xv_realistic <- alfak(yi_realistic,outdir=alfak_output_subdir2,passage_times = as.numeric(colnames(yi_realistic$x)),minobs = 10,nboot=20,pm=beta_for_alfak,n0=n0_val,nb=nb_val))

boot_res_rds_path2 <- file.path(alfak_output_subdir2, "bootstrap_res.Rds")
if(file.exists(boot_res_rds_path2)){
  boot_res_realistic <- readRDS(boot_res_rds_path2)
  print("Median inferred fitness (Stage 1 - Calibrated dt):")
  print(apply(boot_res_realistic$final_fitness,2,median))
  print("Original fitness values:")
  print(landscape_df)
} else {
  print(paste("bootstrap_res.Rds not found in", boot_res_rds_path2))
}
```

With `dt` calibrated, the absolute fitness estimates from `alfak` should now be closer to the original fitness values defined in `landscape_df`.

## Demonstrating `correct_efflux`

When missegregation rates are high, the population growth rate can be dampened by the constant "efflux" of cells into unfit or lethal aneuploid states. `alfak` includes an optional `correct_efflux` parameter to account for this phenomenon when estimating fitness.

### Setup and Data Generation

We define a landscape with a high missegregation rate (`beta = 0.02`) and simulate population dynamics.

```{r efflux_setup}
# Define a simple fitness landscape with 2 chromosome types
# "2.2" is the wild type, "2.3" and "2.4" are neighbors.
ground_truth_landscape <- data.frame(
  k = c("2.2", "2.3", "2.4"),
  mean = c(0.10, 0.11, 0.13) # Fitness values
)

# Simulation settings
beta_val <- 0.02           # High missegregation rate to make the efflux effect visible
n0_val <- 1e5              # Start of passage population
nb_val <- 1e7              # End of passage population
mean_growth_rate <- 0.10   # Approx growth rate of WT

# Calculate realistic passage time (dt)
dt_calc <- log(nb_val / n0_val) / mean_growth_rate

# Time points (simulate 15 passages)
n_passages <- 15
sim_times <- seq(0, n_passages * dt_calc, by = dt_calc)

# Initial condition: 100% Wild Type (2.2)
init_freq <- c(1, 0, 0)
names(init_freq) <- ground_truth_landscape$k

# Run the simulation using predict_evo
ode_res <- predict_evo(
  lscape = ground_truth_landscape,
  p = beta_val,
  times = sim_times,
  x0 = init_freq,
  prediction_type = "ODE"
)

# Convert frequencies to counts (mimicking sequencing depth of ~200 cells)
# We skip the first time point (t=0) as alfak typically expects data after selection
sample_depth <- 200
ode_probs <- ode_res[-1, -1] # Remove time column and t=0 row
counts_matrix <- apply(ode_probs, 1, function(probs) {
  rmultinom(1, size = sample_depth, prob = probs)
})

# Format for alfak (rows = karyotypes, cols = passage numbers)
rownames(counts_matrix) <- colnames(ode_res)[-1]
colnames(counts_matrix) <- 1:n_passages

# Create the input object 'yi'
yi_data <- list(
  x = counts_matrix,
  dt = dt_calc
)
```

### Running `alfak` with and without Correction

We now compare the fitness recovery when `correct_efflux` is enabled versus disabled.

```{r efflux_comparison_run, message=FALSE, warning=FALSE}
# Create temporary directories for output
dir_false <- tempfile("alfak_test_false_")
dir_true  <- tempfile("alfak_test_true_")

# A) Run without correction (Default behavior)
cat("Running alfak with correct_efflux = FALSE...\n")
invisible(utils::capture.output(
  res_false <- alfak(
    yi = yi_data,
    outdir = dir_false,
    passage_times = as.numeric(colnames(yi_data$x)), # Passage numbers
    minobs = 5,      # Low threshold for this small toy data
    nboot = 20,      # Low bootstrap count for speed
    n0 = n0_val,
    nb = nb_val,
    pm = beta_val,
    correct_efflux = FALSE
  )
))

# B) Run with correction
cat("Running alfak with correct_efflux = TRUE...\n")
invisible(utils::capture.output(
  res_true <- alfak(
    yi = yi_data,
    outdir = dir_true,
    passage_times = as.numeric(colnames(yi_data$x)),
    minobs = 5,
    nboot = 20,
    n0 = n0_val,
    nb = nb_val,
    pm = beta_val,
    correct_efflux = TRUE
  )
))
```

### Comparison of Results

We compare the median inferred fitness values against the ground truth.

```{r efflux_results}
# Load the bootstrap results
boot_false <- readRDS(file.path(dir_false, "bootstrap_res.Rds"))
boot_true  <- readRDS(file.path(dir_true, "bootstrap_res.Rds"))

# Extract median fitness estimates
est_false <- apply(boot_false$final_fitness, 2, median)
est_true  <- apply(boot_true$final_fitness, 2, median)

# Combine into a comparison table
comparison <- data.frame(
  Karyotype = names(est_false),
  Truth = ground_truth_landscape$mean[match(names(est_false), ground_truth_landscape$k)],
  Est_False = est_false,
  Est_True = est_true
)

comparison$Diff_False <- abs(comparison$Est_False - comparison$Truth)
comparison$Diff_True  <- abs(comparison$Est_True - comparison$Truth)

print("Comparison of Fitness Estimates:")
print(comparison)

# Cleanup
unlink(dir_false, recursive = TRUE)
unlink(dir_true, recursive = TRUE)
```

## Conclusion

This vignette has demonstrated the use of `predict_evo` for simulating karyotype evolution and validated its ODE component against a custom model. It also provided an example of how to use the `alfak` function to infer fitness landscapes from simulated count data, highlighting the importance of calibrating passage times (`dt`) and correcting for efflux (`correct_efflux`) for accurate absolute fitness estimation.


```{r session_info_chunk, echo=FALSE}
sessionInfo()
```
