---
title: "ALFA-K Benchmark"
subtitle: "Two-round benchmark for nn_prior_grid_n and nn_prior"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
params:
  patient_subset: !r character(0)
  minobs_values: [5, 10, 20]
  pm_values: [5e-05]
  n_cores: 8
  nboot: 45
  n0: 100000
  nb: 10000000
  benchmark_seed: 31415926
  correct_efflux: true
  grid_values: [41, 81, 161]
  round2_priors: ["none", "empirical"]
  include_posterior_comparison: true
  top_shift_n: 15
  force_refit_round1: false
  force_refit_round2: false
  rebuild_inputs: false
  run_round1: true
  run_round2: true
  render_figures: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = TRUE,
  fig.width = 10,
  fig.height = 6,
  dpi = 150
)
```

## Overview

This notebook benchmarks the current local `alfakR` code under `benchmark/data`
using a two-round task design.

- Round 1 fixes `nn_prior = "empirical_censored"` and sweeps
  `nn_prior_grid_n = 41, 81, 161`.
- Round 1 then compares the resulting landscapes for each
  `patient x minobs x pm`, summarizes the global tradeoff, and picks one
  benchmark-wide `nn_prior_grid_n`.
- Round 2 reuses the selected `empirical_censored` fit from Round 1, and only
  runs the missing `nn_prior = "none"` and `nn_prior = "empirical"` tasks.
- Round 2 then compares the three prior modes using the same pairwise
  comparison framework.

The ALFA-K input construction intentionally follows the Figure S5 reanalysis
pipeline in `cloneidR/analysis/FigureS5/FigureS_UseCase_GBM.Rmd`: metadata are
read from `meta_data.xlsx`, `ID_alt` is preferred over `ID`, transcriptome
profiles are converted into karyotype strings, and the primary-to-recurrent
interval is encoded as `Primary = 1`, `Recurrent = abs(Age_recurrent -
Age_primary) * 360` days.

```{r paths-and-libraries}
suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(knitr)
  library(parallel)
  library(pkgload)
  library(purrr)
  library(readxl)
  library(tibble)
  library(tidyr)
})

resolve_repo_dir <- function() {
  candidates <- unique(c(
    tryCatch(dirname(knitr::current_input(dir = TRUE)), error = function(e) NA_character_),
    getwd()
  ))

  for (cand in candidates) {
    if (is.na(cand) || !nzchar(cand)) {
      next
    }
    cand_norm <- normalizePath(cand, winslash = "/", mustWork = FALSE)
    if (file.exists(file.path(cand_norm, "DESCRIPTION")) &&
        dir.exists(file.path(cand_norm, "benchmark"))) {
      return(cand_norm)
    }
    parent2 <- normalizePath(file.path(cand_norm, "..", ".."), winslash = "/", mustWork = FALSE)
    if (file.exists(file.path(parent2, "DESCRIPTION")) &&
        dir.exists(file.path(parent2, "benchmark"))) {
      return(parent2)
    }
  }

  stop("Could not locate the alfakR repository root from the current knitting context.")
}

repo_dir <- resolve_repo_dir()
benchmark_dir <- file.path(repo_dir, "benchmark")
data_dir <- file.path(benchmark_dir, "data")
meta_path <- file.path(data_dir, "meta_data.xlsx")
results_dir <- file.path(benchmark_dir, "results")
input_dir <- file.path(results_dir, "inputs")
grid_fit_dir <- file.path(results_dir, "fits", "grid_round")
prior_fit_dir <- file.path(results_dir, "fits", "prior_round")
tables_dir <- file.path(results_dir, "tables")
figures_dir <- file.path(results_dir, "figures")
cache_dir <- file.path(results_dir, "cache")

dir.create(results_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(input_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(grid_fit_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(prior_fit_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(tables_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(figures_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(cache_dir, recursive = TRUE, showWarnings = FALSE)

pkgload::load_all(repo_dir, quiet = TRUE)

patient_subset_use <- unique(trimws(as.character(unlist(params$patient_subset))))
patient_subset_use <- patient_subset_use[nzchar(patient_subset_use)]
if (!length(patient_subset_use)) {
  patient_subset_use <- NULL
}

minobs_values_use <- sort(unique(as.integer(unlist(params$minobs_values))))
minobs_values_use <- minobs_values_use[is.finite(minobs_values_use) & minobs_values_use > 0L]
if (!length(minobs_values_use)) {
  stop("No valid minobs values supplied.")
}

pm_values_use <- sort(unique(as.numeric(unlist(params$pm_values))))
pm_values_use <- pm_values_use[is.finite(pm_values_use) & pm_values_use > 0]
if (!length(pm_values_use)) {
  stop("No valid pm values supplied.")
}

grid_values_use <- sort(unique(as.integer(unlist(params$grid_values))))
grid_values_use <- grid_values_use[is.finite(grid_values_use) & grid_values_use >= 3L]
if (!length(grid_values_use)) {
  stop("No valid nn_prior_grid_n values supplied.")
}

round2_priors_use <- unique(as.character(unlist(params$round2_priors)))
round2_priors_use <- round2_priors_use[nzchar(round2_priors_use)]
if (!all(round2_priors_use %in% c("none", "empirical"))) {
  stop("Round 2 priors must be chosen from: none, empirical")
}

n_cores_use <- suppressWarnings(as.integer(params$n_cores))
if (!is.finite(n_cores_use) || n_cores_use < 1L) {
  detected_cores <- suppressWarnings(parallel::detectCores(logical = FALSE))
  if (!is.finite(detected_cores) || detected_cores < 1L) {
    detected_cores <- 1L
  }
  n_cores_use <- max(1L, detected_cores - 1L)
}

nboot_use <- as.integer(params$nboot)
n0_use <- as.numeric(params$n0)
nb_use <- as.numeric(params$nb)
benchmark_seed_use <- suppressWarnings(as.integer(params$benchmark_seed))
if (!is.finite(benchmark_seed_use)) {
  stop("benchmark_seed must be a finite integer.")
}
correct_efflux_use <- isTRUE(params$correct_efflux)
force_refit_round1 <- isTRUE(params$force_refit_round1)
force_refit_round2 <- isTRUE(params$force_refit_round2)
rebuild_inputs_use <- isTRUE(params$rebuild_inputs)
run_round1_use <- isTRUE(params$run_round1)
run_round2_use <- isTRUE(params$run_round2)
include_posterior_use <- isTRUE(params$include_posterior_comparison)
render_figures_use <- isTRUE(params$render_figures)
top_shift_n_use <- max(1L, as.integer(params$top_shift_n))

reference_grid_n <- max(grid_values_use)
diploid_state <- paste(rep(2, 22), collapse = ".")
stage_levels <- c("Primary", "Recurrent")

set.seed(benchmark_seed_use)
options(mc.cores = n_cores_use)
```

## Helper Functions

```{r helpers}
sort_pid_levels <- function(x) {
  x <- unique(as.character(x))
  ord_num <- suppressWarnings(as.integer(sub("^P", "", x)))
  x[order(ifelse(is.na(ord_num), Inf, ord_num), x)]
}

pm_to_label <- function(pm) {
  format(pm, scientific = FALSE, trim = TRUE)
}

write_tsv_base <- function(x, path) {
  utils::write.table(
    x,
    file = path,
    sep = "\t",
    quote = FALSE,
    row.names = FALSE,
    col.names = TRUE
  )
  invisible(path)
}

save_table_bundle <- function(x, stem) {
  saveRDS(x, paste0(stem, ".rds"))
  if (is.data.frame(x)) {
    write_tsv_base(x, paste0(stem, ".tsv"))
  }
  invisible(stem)
}

load_saved_table <- function(stem) {
  rds_path <- paste0(stem, ".rds")
  if (file.exists(rds_path)) {
    return(readRDS(rds_path))
  }
  NULL
}

render_tbl <- function(x, caption = NULL, digits = 4) {
  fmt <- if (knitr::is_latex_output()) "latex" else "pipe"
  knitr::kable(
    x,
    format = fmt,
    digits = digits,
    caption = caption,
    booktabs = knitr::is_latex_output(),
    longtable = knitr::is_latex_output()
  )
}

empty_note_tbl <- function(note) {
  tibble(note = note)
}

alfak_log <- function(...) {
  msg <- paste0(...)
  timestamp <- format(Sys.time(), "%Y-%m-%d %H:%M:%S")
  message("[", timestamp, "] ", msg)
  flush.console()
}

normalize_01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  if (!all(is.finite(rng)) || diff(rng) == 0) {
    return(rep(0, length(x)))
  }
  (x - rng[1]) / diff(rng)
}

available_pid_dirs <- function(base_dir) {
  dd <- list.dirs(base_dir, recursive = FALSE, full.names = TRUE)
  dd <- dd[file.info(dd)$isdir %in% TRUE]
  dd <- dd[grepl("^P[0-9]+$", basename(dd))]
  sort_pid_levels(basename(dd))
}

read_transcriptome_perspective <- function(rds_path) {
  obj <- readRDS(rds_path)
  stopifnot(is.list(obj), "profile" %in% names(obj))
  profile <- as.matrix(obj$profile)
  storage.mode(profile) <- "numeric"
  list(
    passaging_id = obj$passaging_id,
    perspective_type = obj$perspective_type,
    profile = round(profile)
  )
}

compute_delta_time_days <- function(df_pid) {
  df_pid <- df_pid %>%
    mutate(Stage = factor(Stage, levels = stage_levels)) %>%
    arrange(Stage, effective_id)

  age_primary <- suppressWarnings(as.numeric(df_pid$Age[df_pid$Stage == "Primary"][1]))
  age_recurrent <- suppressWarnings(as.numeric(df_pid$Age[df_pid$Stage == "Recurrent"][1]))

  recurrent_days <- if (is.finite(age_primary) && is.finite(age_recurrent)) {
    as.integer(round(abs(age_recurrent - age_primary) * 360))
  } else {
    NA_integer_
  }
  if (!is.finite(recurrent_days) || recurrent_days <= 1L) {
    recurrent_days <- 180L
  }

  df_pid %>%
    mutate(
      Delta_Time = ifelse(as.character(Stage) == "Primary", 1L, recurrent_days)
    )
}

build_patient_manifest <- function(meta_tbl, base_dir, patient_subset = NULL) {
  pid_dirs <- available_pid_dirs(base_dir)
  manifest <- meta_tbl %>%
    mutate(
      effective_id = ifelse(!is.na(ID_alt) & nzchar(ID_alt), ID_alt, ID),
      pid = as.character(pid),
      Stage = as.character(Stage),
      Age = suppressWarnings(as.numeric(Age))
    ) %>%
    filter(pid %in% pid_dirs)

  if (!is.null(patient_subset)) {
    manifest <- manifest %>% filter(pid %in% patient_subset)
  }

  manifest %>%
    group_by(pid) %>%
    group_modify(function(.x, .y) compute_delta_time_days(.x)) %>%
    ungroup() %>%
    mutate(
      rds_path = file.path(base_dir, pid, "perspectives", effective_id, "TranscriptomePerspective.rds"),
      exists = file.exists(rds_path)
    ) %>%
    arrange(factor(pid, levels = sort_pid_levels(pid)), factor(Stage, levels = stage_levels))
}

profile_to_karyotype_strings <- function(profile_mat) {
  if (nrow(profile_mat) != 22) {
    warning("Expected 22 chromosome rows; got ", nrow(profile_mat))
  }
  x <- t(round(as.matrix(profile_mat)))
  apply(x, 1, paste, collapse = ".")
}

build_count_matrix_from_profiles <- function(patient_manifest) {
  stopifnot(length(unique(patient_manifest$pid)) == 1L)
  rows_ok <- patient_manifest[patient_manifest$exists, , drop = FALSE]
  rows_ok <- rows_ok[order(match(rows_ok$Stage, stage_levels)), , drop = FALSE]
  if (nrow(rows_ok) < 2) {
    stop("Need at least two available samples for patient ", unique(patient_manifest$pid))
  }

  long_df <- do.call(rbind, lapply(seq_len(nrow(rows_ok)), function(i) {
    rr <- rows_ok[i, , drop = FALSE]
    obj <- read_transcriptome_perspective(rr$rds_path)
    if (!is.null(obj$passaging_id) &&
        !identical(as.character(obj$passaging_id), as.character(rr$effective_id))) {
      warning("passaging_id mismatch for ", rr$pid, " / ", rr$effective_id)
    }
    tibble(
      karyotype = profile_to_karyotype_strings(obj$profile),
      Delta_Time = rr$Delta_Time,
      Stage = rr$Stage,
      effective_id = rr$effective_id
    )
  }))

  delta_levels <- sort(unique(rows_ok$Delta_Time))
  count_mat <- as.data.frame.matrix(table(
    long_df$karyotype,
    factor(long_df$Delta_Time, levels = delta_levels)
  ))
  colnames(count_mat) <- as.character(delta_levels)
  count_mat <- count_mat[order(rowSums(count_mat), decreasing = TRUE), , drop = FALSE]

  list(
    count_matrix = count_mat,
    long_df = long_df
  )
}

save_alfak_input <- function(count_matrix, out_path) {
  yi <- list(
    x = count_matrix,
    pop.fitness = NULL,
    dt = 1
  )
  saveRDS(yi, out_path)
  invisible(out_path)
}

build_benchmark_inputs <- function(meta_tbl, base_dir, input_dir, rebuild_inputs = FALSE, patient_subset = NULL) {
  manifest_tbl <- build_patient_manifest(meta_tbl, base_dir, patient_subset = patient_subset)
  save_table_bundle(manifest_tbl, file.path(tables_dir, "benchmark_patient_manifest"))

  patient_ids <- sort_pid_levels(unique(manifest_tbl$pid))
  if (!length(patient_ids)) {
    stop("No benchmark patients discovered under ", base_dir)
  }

  rows <- lapply(patient_ids, function(pid) {
    patient_manifest <- manifest_tbl %>% filter(pid == !!pid)
    input_rds <- file.path(input_dir, paste0(pid, ".Rds"))
    count_matrix_path <- file.path(input_dir, paste0(pid, "_count_matrix.tsv"))
    long_df_path <- file.path(input_dir, paste0(pid, "_profile_long.tsv"))

    if (rebuild_inputs || !file.exists(input_rds) || !file.exists(count_matrix_path) || !file.exists(long_df_path)) {
      built <- build_count_matrix_from_profiles(patient_manifest)
      count_export <- data.frame(
        karyotype = rownames(built$count_matrix),
        built$count_matrix,
        check.names = FALSE
      )
      save_alfak_input(built$count_matrix, input_rds)
      write_tsv_base(count_export, count_matrix_path)
      write_tsv_base(built$long_df, long_df_path)
    }

    yi <- readRDS(input_rds)
    yi$x <- as.data.frame(yi$x)

    tibble(
      patient_id = pid,
      input_rds = input_rds,
      count_matrix_path = count_matrix_path,
      long_df_path = long_df_path,
      n_timepoints = ncol(yi$x),
      n_karyotypes = nrow(yi$x),
      total_cells = sum(as.matrix(yi$x), na.rm = TRUE),
      delta_time_labels = paste(colnames(yi$x), collapse = ","),
      effective_ids = paste(patient_manifest$effective_id[patient_manifest$exists], collapse = ","),
      stage_sequence = paste(as.character(patient_manifest$Stage[patient_manifest$exists]), collapse = " -> ")
    )
  })

  bind_rows(rows) %>%
    arrange(factor(patient_id, levels = sort_pid_levels(patient_id)))
}

extract_xval_metrics <- function(xv) {
  metrics <- list(
    xval_r2 = NA_real_,
    xval_cor = NA_real_,
    xval_rmse = NA_real_,
    xval_mae = NA_real_,
    n_xval = NA_integer_
  )

  if (is.null(xv)) {
    return(metrics)
  }

  if (is.atomic(xv) && is.numeric(xv) && is.null(dim(xv)) && length(xv) >= 1L) {
    metrics$xval_r2 <- suppressWarnings(as.numeric(xv[[1]]))
    return(metrics)
  }

  if (!is.list(xv)) {
    return(metrics)
  }

  if (!is.null(xv$R2R)) {
    metrics$xval_r2 <- suppressWarnings(as.numeric(xv$R2R))
  }

  if (is.null(xv$tmp)) {
    return(metrics)
  }

  tmp_df <- as.data.frame(xv$tmp)
  if (ncol(tmp_df) < 2) {
    return(metrics)
  }

  colnames(tmp_df)[1:2] <- c("f_est", "f_xv")
  tmp_df <- tmp_df[, c("f_est", "f_xv"), drop = FALSE]
  tmp_df$f_est <- suppressWarnings(as.numeric(tmp_df$f_est))
  tmp_df$f_xv <- suppressWarnings(as.numeric(tmp_df$f_xv))
  tmp_df <- tmp_df[is.finite(tmp_df$f_est) & is.finite(tmp_df$f_xv), , drop = FALSE]

  metrics$n_xval <- nrow(tmp_df)
  if (!nrow(tmp_df)) {
    return(metrics)
  }

  diff_vec <- tmp_df$f_est - tmp_df$f_xv
  metrics$xval_rmse <- sqrt(mean(diff_vec^2))
  metrics$xval_mae <- mean(abs(diff_vec))
  if (nrow(tmp_df) >= 2) {
    metrics$xval_cor <- suppressWarnings(stats::cor(tmp_df$f_est, tmp_df$f_xv, use = "complete.obs"))
  }
  metrics
}

run_alfak_fit <- function(patient_id,
                          input_rds,
                          outdir,
                          minobs,
                          pm,
                          nboot,
                          n0,
                          nb,
                          benchmark_seed,
                          correct_efflux = TRUE,
                          nn_prior = "none",
                          nn_prior_grid_n = 81L,
                          force_refit = FALSE) {
  landscape_path <- file.path(outdir, "landscape.Rds")
  bootstrap_path <- file.path(outdir, "bootstrap_res.Rds")
  posterior_path <- file.path(outdir, "landscape_posterior_samples.Rds")
  xval_path <- file.path(outdir, "xval.Rds")
  summary_path <- file.path(outdir, "fit_task_result.rds")
  warning_log_path <- file.path(outdir, "fit_warnings.log")
  pm_label <- pm_to_label(pm)
  task_tag <- paste0(
    patient_id,
    " | minobs=", minobs,
    " | pm=", pm_label,
    " | nn_prior=", nn_prior,
    " | grid=", nn_prior_grid_n
  )

  if (!force_refit && file.exists(summary_path)) {
    cached <- readRDS(summary_path)
    cached$benchmark_seed <- benchmark_seed
    cached$cached <- TRUE
    return(as_tibble(cached))
  }

  if (!force_refit && file.exists(landscape_path) && file.exists(xval_path)) {
    xv <- tryCatch(readRDS(xval_path), error = function(e) NULL)
    cached <- c(
      list(
        patient_id = patient_id,
        outdir = outdir,
        pm = pm,
        pm_label = pm_label,
        minobs = minobs,
        benchmark_seed = benchmark_seed,
        nn_prior = nn_prior,
        nn_prior_grid_n = nn_prior_grid_n,
        status = "ok",
        cached = TRUE,
        error_message = NA_character_,
        elapsed_sec = NA_real_,
        warning_count = if (file.exists(warning_log_path)) length(readLines(warning_log_path, warn = FALSE)) else 0L,
        lambda_endpoint_warning_count = if (file.exists(warning_log_path)) {
          sum(grepl("^Grid searches over lambda", readLines(warning_log_path, warn = FALSE)))
        } else {
          0L
        },
        warning_messages = if (file.exists(warning_log_path)) paste(readLines(warning_log_path, warn = FALSE), collapse = " || ") else NA_character_,
        landscape_path = landscape_path,
        bootstrap_path = if (file.exists(bootstrap_path)) bootstrap_path else NA_character_,
        posterior_path = if (file.exists(posterior_path)) posterior_path else NA_character_,
        xval_path = xval_path
      ),
      extract_xval_metrics(xv)
    )
    saveRDS(cached, summary_path)
    return(as_tibble(cached))
  }

  alfak_log("ALFA-K start: ", task_tag)
  dir.create(outdir, recursive = TRUE, showWarnings = FALSE)

  yi <- readRDS(input_rds)
  yi$x <- as.data.frame(yi$x)
  if (diploid_state %in% rownames(yi$x)) {
    yi$x <- yi$x[rownames(yi$x) != diploid_state, , drop = FALSE]
  }
  if (!nrow(yi$x)) {
    stop("All rows were filtered out for input: ", input_rds)
  }
  if (max(rowSums(yi$x), na.rm = TRUE) < minobs) {
    stop("No frequent karyotypes reach minobs=", minobs, " for input: ", input_rds)
  }

  warning_messages <- character()
  started_at <- Sys.time()

  fit_row <- tryCatch({
    set.seed(benchmark_seed)
    withCallingHandlers({
      alfakR::alfak(
        yi = yi,
        outdir = outdir,
        passage_times = NULL,
        minobs = minobs,
        nboot = nboot,
        n0 = n0,
        nb = nb,
        pm = pm,
        correct_efflux = correct_efflux,
        nn_prior = nn_prior,
        nn_prior_grid_n = nn_prior_grid_n
      )
    }, warning = function(w) {
      warning_messages <<- c(warning_messages, conditionMessage(w))
      invokeRestart("muffleWarning")
    })

    xv <- readRDS(xval_path)
    elapsed_sec <- as.numeric(difftime(Sys.time(), started_at, units = "secs"))
    warning_messages <- unique(warning_messages)
    if (length(warning_messages)) {
      writeLines(warning_messages, warning_log_path)
    }

    c(
      list(
        patient_id = patient_id,
        outdir = outdir,
        pm = pm,
        pm_label = pm_label,
        minobs = minobs,
        benchmark_seed = benchmark_seed,
        nn_prior = nn_prior,
        nn_prior_grid_n = nn_prior_grid_n,
        status = "ok",
        cached = FALSE,
        error_message = NA_character_,
        elapsed_sec = elapsed_sec,
        warning_count = length(warning_messages),
        lambda_endpoint_warning_count = sum(grepl("^Grid searches over lambda", warning_messages)),
        warning_messages = if (length(warning_messages)) paste(warning_messages, collapse = " || ") else NA_character_,
        landscape_path = landscape_path,
        bootstrap_path = if (file.exists(bootstrap_path)) bootstrap_path else NA_character_,
        posterior_path = if (file.exists(posterior_path)) posterior_path else NA_character_,
        xval_path = xval_path
      ),
      extract_xval_metrics(xv)
    )
  }, error = function(e) {
    elapsed_sec <- as.numeric(difftime(Sys.time(), started_at, units = "secs"))
    warning_messages <- unique(warning_messages)
    if (length(warning_messages)) {
      writeLines(warning_messages, warning_log_path)
    }

    alfak_log("ALFA-K error: ", task_tag, " | ", conditionMessage(e))
    list(
      patient_id = patient_id,
      outdir = outdir,
      pm = pm,
      pm_label = pm_label,
      minobs = minobs,
      benchmark_seed = benchmark_seed,
      nn_prior = nn_prior,
      nn_prior_grid_n = nn_prior_grid_n,
      status = "error",
      cached = FALSE,
      error_message = conditionMessage(e),
      elapsed_sec = elapsed_sec,
      warning_count = length(warning_messages),
      lambda_endpoint_warning_count = sum(grepl("^Grid searches over lambda", warning_messages)),
      warning_messages = if (length(warning_messages)) paste(warning_messages, collapse = " || ") else NA_character_,
      xval_r2 = NA_real_,
      xval_cor = NA_real_,
      xval_rmse = NA_real_,
      xval_mae = NA_real_,
      n_xval = NA_integer_,
      landscape_path = if (file.exists(landscape_path)) landscape_path else NA_character_,
      bootstrap_path = if (file.exists(bootstrap_path)) bootstrap_path else NA_character_,
      posterior_path = if (file.exists(posterior_path)) posterior_path else NA_character_,
      xval_path = if (file.exists(xval_path)) xval_path else NA_character_
    )
  })

  saveRDS(fit_row, summary_path)
  if (identical(fit_row$status, "ok")) {
    alfak_log("ALFA-K done: ", task_tag, " | xval_r2=", signif(fit_row$xval_r2, 4))
  }
  as_tibble(fit_row)
}

task_outdir_grid <- function(root_dir, patient_id, minobs, pm, grid_n) {
  file.path(
    root_dir,
    "nn_prior_empirical_censored",
    paste0("grid_", grid_n),
    paste0("pm_", pm_to_label(pm)),
    paste0("MINOBS_", minobs),
    patient_id
  )
}

task_outdir_prior <- function(root_dir, patient_id, minobs, pm, nn_prior, grid_n) {
  file.path(
    root_dir,
    paste0("nn_prior_", nn_prior),
    paste0("grid_", grid_n),
    paste0("pm_", pm_to_label(pm)),
    paste0("MINOBS_", minobs),
    patient_id
  )
}

build_grid_tasks <- function(input_index_tbl,
                             fit_root,
                             minobs_values,
                             pm_values,
                             grid_values,
                             nboot,
                             n0,
                             nb,
                             benchmark_seed,
                             correct_efflux,
                             force_refit) {
  tidyr::crossing(
    input_index_tbl %>% select(patient_id, input_rds),
    minobs = minobs_values,
    pm = pm_values,
    nn_prior_grid_n = grid_values
  ) %>%
    mutate(
      nn_prior = "empirical_censored",
      outdir = purrr::pmap_chr(
        list(patient_id, minobs, pm, nn_prior_grid_n),
        ~ task_outdir_grid(fit_root, ..1, ..2, ..3, ..4)
      ),
      nboot = nboot,
      n0 = n0,
      nb = nb,
      benchmark_seed = benchmark_seed,
      correct_efflux = correct_efflux,
      force_refit = force_refit,
      round_label = "grid_round"
    ) %>%
    arrange(factor(patient_id, levels = sort_pid_levels(patient_id)), minobs, pm, nn_prior_grid_n)
}

build_prior_tasks <- function(input_index_tbl,
                              fit_root,
                              minobs_values,
                              pm_values,
                              best_grid_n,
                              priors,
                              nboot,
                              n0,
                              nb,
                              benchmark_seed,
                              correct_efflux,
                              force_refit) {
  tidyr::crossing(
    input_index_tbl %>% select(patient_id, input_rds),
    minobs = minobs_values,
    pm = pm_values,
    nn_prior = priors
  ) %>%
    mutate(
      nn_prior_grid_n = best_grid_n,
      outdir = purrr::pmap_chr(
        list(patient_id, minobs, pm, nn_prior),
        ~ task_outdir_prior(fit_root, ..1, ..2, ..3, ..4, best_grid_n)
      ),
      nboot = nboot,
      n0 = n0,
      nb = nb,
      benchmark_seed = benchmark_seed,
      correct_efflux = correct_efflux,
      force_refit = force_refit,
      round_label = "prior_round"
    ) %>%
    arrange(factor(patient_id, levels = sort_pid_levels(patient_id)), minobs, pm, nn_prior)
}

run_task_table_parallel <- function(task_tbl) {
  if (!nrow(task_tbl)) {
    return(tibble())
  }

  if (.Platform$OS.type == "unix" && n_cores_use > 1L) {
    res <- parallel::mclapply(
      seq_len(nrow(task_tbl)),
      function(i) {
        rr <- task_tbl[i, , drop = FALSE]
        run_alfak_fit(
          patient_id = rr$patient_id,
          input_rds = rr$input_rds,
          outdir = rr$outdir,
          minobs = rr$minobs,
          pm = rr$pm,
          nboot = rr$nboot,
          n0 = rr$n0,
          nb = rr$nb,
          benchmark_seed = rr$benchmark_seed,
          correct_efflux = rr$correct_efflux,
          nn_prior = rr$nn_prior,
          nn_prior_grid_n = rr$nn_prior_grid_n,
          force_refit = rr$force_refit
        )
      },
      mc.cores = n_cores_use,
      mc.preschedule = FALSE,
      mc.set.seed = FALSE
    )
  } else {
    res <- lapply(seq_len(nrow(task_tbl)), function(i) {
      rr <- task_tbl[i, , drop = FALSE]
      run_alfak_fit(
        patient_id = rr$patient_id,
        input_rds = rr$input_rds,
        outdir = rr$outdir,
        minobs = rr$minobs,
        pm = rr$pm,
        nboot = rr$nboot,
        n0 = rr$n0,
        nb = rr$nb,
        benchmark_seed = rr$benchmark_seed,
        correct_efflux = rr$correct_efflux,
        nn_prior = rr$nn_prior,
        nn_prior_grid_n = rr$nn_prior_grid_n,
        force_refit = rr$force_refit
      )
    })
  }

  bind_rows(res)
}

summarize_fit_results <- function(fit_results_tbl, group_cols) {
  if (!nrow(fit_results_tbl)) {
    return(tibble())
  }

  fit_results_tbl %>%
    group_by(across(all_of(group_cols))) %>%
    summarise(
      n_tasks = n(),
      n_ok = sum(status == "ok", na.rm = TRUE),
      n_error = sum(status == "error", na.rm = TRUE),
      success_rate = n_ok / n_tasks,
      mean_xval_r2 = mean(xval_r2, na.rm = TRUE),
      median_xval_r2 = median(xval_r2, na.rm = TRUE),
      mean_elapsed_sec = mean(elapsed_sec, na.rm = TRUE),
      median_elapsed_sec = median(elapsed_sec, na.rm = TRUE),
      total_lambda_endpoint_warnings = sum(lambda_endpoint_warning_count, na.rm = TRUE),
      .groups = "drop"
    )
}

safe_read_rds <- function(path) {
  if (is.na(path) || !nzchar(path) || !file.exists(path)) {
    return(NULL)
  }
  readRDS(path)
}

read_fit_bundle <- function(fit_dir) {
  list(
    bootstrap = safe_read_rds(file.path(fit_dir, "bootstrap_res.Rds")),
    landscape = safe_read_rds(file.path(fit_dir, "landscape.Rds")),
    posterior = safe_read_rds(file.path(fit_dir, "landscape_posterior_samples.Rds")),
    xval = safe_read_rds(file.path(fit_dir, "xval.Rds"))
  )
}

flatten_pair_values <- function(lhs, rhs) {
  if (is.null(lhs) || is.null(rhs)) {
    return(list(lhs = numeric(), rhs = numeric(), lhs_total = 0L, rhs_total = 0L, n_used = 0L))
  }

  if (!is.null(dim(lhs)) && !is.null(dim(rhs))) {
    lhs_mat <- as.matrix(lhs)
    rhs_mat <- as.matrix(rhs)

    if (!is.null(colnames(lhs_mat)) && !is.null(colnames(rhs_mat))) {
      common_cols <- intersect(colnames(lhs_mat), colnames(rhs_mat))
      if (length(common_cols)) {
        lhs_mat <- lhs_mat[, common_cols, drop = FALSE]
        rhs_mat <- rhs_mat[, common_cols, drop = FALSE]
      }
    }

    n_rows <- min(nrow(lhs_mat), nrow(rhs_mat))
    if (n_rows < 1L) {
      return(list(lhs = numeric(), rhs = numeric(), lhs_total = length(lhs_mat), rhs_total = length(rhs_mat), n_used = 0L))
    }
    lhs_vec <- as.numeric(lhs_mat[seq_len(n_rows), , drop = FALSE])
    rhs_vec <- as.numeric(rhs_mat[seq_len(n_rows), , drop = FALSE])
  } else {
    lhs_vec <- as.numeric(lhs)
    rhs_vec <- as.numeric(rhs)
  }

  n_used <- min(length(lhs_vec), length(rhs_vec))
  if (n_used < 1L) {
    return(list(lhs = numeric(), rhs = numeric(), lhs_total = length(lhs_vec), rhs_total = length(rhs_vec), n_used = 0L))
  }

  list(
    lhs = lhs_vec[seq_len(n_used)],
    rhs = rhs_vec[seq_len(n_used)],
    lhs_total = length(lhs_vec),
    rhs_total = length(rhs_vec),
    n_used = n_used
  )
}

compare_numeric_objects <- function(lhs, rhs, metric_name) {
  aligned <- flatten_pair_values(lhs, rhs)
  lhs_vec <- aligned$lhs
  rhs_vec <- aligned$rhs
  if (!length(lhs_vec) || !length(rhs_vec)) {
    return(tibble(
      metric = metric_name,
      n = 0L,
      lhs_mean = NA_real_,
      rhs_mean = NA_real_,
      mean_diff = NA_real_,
      mean_abs_diff = NA_real_,
      max_abs_diff = NA_real_,
      correlation = NA_real_,
      lhs_negative = NA_integer_,
      rhs_negative = NA_integer_,
      lhs_total = aligned$lhs_total,
      rhs_total = aligned$rhs_total
    ))
  }

  diff_vec <- rhs_vec - lhs_vec
  tibble(
    metric = metric_name,
    n = aligned$n_used,
    lhs_mean = mean(lhs_vec, na.rm = TRUE),
    rhs_mean = mean(rhs_vec, na.rm = TRUE),
    mean_diff = mean(diff_vec, na.rm = TRUE),
    mean_abs_diff = mean(abs(diff_vec), na.rm = TRUE),
    max_abs_diff = max(abs(diff_vec), na.rm = TRUE),
    correlation = if (aligned$n_used >= 2L) {
      suppressWarnings(stats::cor(lhs_vec, rhs_vec, use = "complete.obs"))
    } else {
      NA_real_
    },
    lhs_negative = sum(lhs_vec < 0, na.rm = TRUE),
    rhs_negative = sum(rhs_vec < 0, na.rm = TRUE),
    lhs_total = aligned$lhs_total,
    rhs_total = aligned$rhs_total
  )
}

summarize_sign_changes <- function(lhs, rhs, metric_name) {
  aligned <- flatten_pair_values(lhs, rhs)
  lhs_vec <- aligned$lhs
  rhs_vec <- aligned$rhs
  if (!length(lhs_vec) || !length(rhs_vec)) {
    return(tibble(
      metric = metric_name,
      n = 0L,
      lhs_negative = NA_integer_,
      rhs_negative = NA_integer_,
      neg_to_nonneg = NA_integer_,
      nonneg_to_neg = NA_integer_,
      across_zero = NA_integer_
    ))
  }

  tibble(
    metric = metric_name,
    n = aligned$n_used,
    lhs_negative = sum(lhs_vec < 0, na.rm = TRUE),
    rhs_negative = sum(rhs_vec < 0, na.rm = TRUE),
    neg_to_nonneg = sum(lhs_vec < 0 & rhs_vec >= 0, na.rm = TRUE),
    nonneg_to_neg = sum(lhs_vec >= 0 & rhs_vec < 0, na.rm = TRUE),
    across_zero = sum(lhs_vec * rhs_vec < 0, na.rm = TRUE)
  )
}

merge_landscape_pair <- function(lhs_landscape, rhs_landscape) {
  lhs_tbl <- lhs_landscape %>%
    transmute(
      k,
      lhs_mean = mean,
      lhs_median = median,
      lhs_sd = sd,
      lhs_fq = as.logical(fq),
      lhs_nn = as.logical(nn)
    )
  rhs_tbl <- rhs_landscape %>%
    transmute(
      k,
      rhs_mean = mean,
      rhs_median = median,
      rhs_sd = sd,
      rhs_fq = as.logical(fq),
      rhs_nn = as.logical(nn)
    )

  inner_join(lhs_tbl, rhs_tbl, by = "k") %>%
    mutate(
      state_group = case_when(
        lhs_fq | rhs_fq ~ "fq",
        lhs_nn | rhs_nn ~ "nn",
        TRUE ~ "other"
      )
    )
}

summarize_landscape_groups <- function(merged_landscape, lhs_col, rhs_col, metric_name) {
  if (!nrow(merged_landscape)) {
    return(tibble())
  }

  diff_vec <- merged_landscape[[rhs_col]] - merged_landscape[[lhs_col]]
  tmp <- merged_landscape %>%
    mutate(
      lhs_value = .data[[lhs_col]],
      rhs_value = .data[[rhs_col]],
      diff_value = diff_vec
    ) %>%
    group_by(state_group) %>%
    summarise(
      metric = metric_name,
      n = n(),
      lhs_mean = mean(lhs_value, na.rm = TRUE),
      rhs_mean = mean(rhs_value, na.rm = TRUE),
      mean_diff = mean(diff_value, na.rm = TRUE),
      mean_abs_diff = mean(abs(diff_value), na.rm = TRUE),
      max_abs_diff = max(abs(diff_value), na.rm = TRUE),
      correlation = if (n() >= 2L) {
        suppressWarnings(stats::cor(lhs_value, rhs_value, use = "complete.obs"))
      } else {
        NA_real_
      },
      .groups = "drop"
    )

  tmp %>% select(metric, state_group, everything())
}

top_landscape_shifts <- function(merged_landscape, top_n = 15L) {
  if (!nrow(merged_landscape)) {
    return(tibble())
  }

  merged_landscape %>%
    transmute(
      k,
      state_group,
      lhs_mean,
      rhs_mean,
      diff = rhs_mean - lhs_mean
    ) %>%
    arrange(desc(abs(diff))) %>%
    slice_head(n = top_n)
}

compare_fit_pair <- function(lhs_dir,
                             rhs_dir,
                             patient_id,
                             minobs,
                             pm,
                             lhs_label,
                             rhs_label,
                             lhs_setting,
                             rhs_setting,
                             comparison_set,
                             top_n = 15L,
                             include_posterior = TRUE) {
  lhs_bundle <- read_fit_bundle(lhs_dir)
  rhs_bundle <- read_fit_bundle(rhs_dir)

  component_tbl <- bind_rows(
    compare_numeric_objects(lhs_bundle$bootstrap$initial_fitness, rhs_bundle$bootstrap$initial_fitness, "bootstrap_initial_fitness"),
    compare_numeric_objects(lhs_bundle$bootstrap$final_fitness, rhs_bundle$bootstrap$final_fitness, "bootstrap_final_fitness"),
    compare_numeric_objects(lhs_bundle$bootstrap$initial_frequencies, rhs_bundle$bootstrap$initial_frequencies, "bootstrap_initial_frequencies"),
    compare_numeric_objects(lhs_bundle$bootstrap$final_frequencies, rhs_bundle$bootstrap$final_frequencies, "bootstrap_final_frequencies"),
    compare_numeric_objects(lhs_bundle$bootstrap$nn_fitness, rhs_bundle$bootstrap$nn_fitness, "bootstrap_nn_fitness"),
    compare_numeric_objects(lhs_bundle$landscape$mean, rhs_bundle$landscape$mean, "landscape_mean"),
    compare_numeric_objects(lhs_bundle$landscape$median, rhs_bundle$landscape$median, "landscape_median"),
    compare_numeric_objects(lhs_bundle$landscape$sd, rhs_bundle$landscape$sd, "landscape_sd"),
    if (include_posterior) {
      compare_numeric_objects(lhs_bundle$posterior, rhs_bundle$posterior, "landscape_posterior_samples")
    },
    compare_numeric_objects(
      extract_xval_metrics(lhs_bundle$xval)$xval_r2,
      extract_xval_metrics(rhs_bundle$xval)$xval_r2,
      "xval_r2"
    )
  ) %>%
    mutate(
      comparison_set = comparison_set,
      patient_id = patient_id,
      minobs = minobs,
      pm = pm,
      lhs_label = lhs_label,
      rhs_label = rhs_label,
      lhs_setting = lhs_setting,
      rhs_setting = rhs_setting
    ) %>%
    relocate(comparison_set, patient_id, minobs, pm, lhs_label, rhs_label, lhs_setting, rhs_setting, metric)

  sign_tbl <- bind_rows(
    summarize_sign_changes(lhs_bundle$bootstrap$nn_fitness, rhs_bundle$bootstrap$nn_fitness, "bootstrap_nn_fitness"),
    summarize_sign_changes(lhs_bundle$landscape$mean, rhs_bundle$landscape$mean, "landscape_mean"),
    summarize_sign_changes(lhs_bundle$landscape$median, rhs_bundle$landscape$median, "landscape_median")
  ) %>%
    mutate(
      comparison_set = comparison_set,
      patient_id = patient_id,
      minobs = minobs,
      pm = pm,
      lhs_label = lhs_label,
      rhs_label = rhs_label,
      lhs_setting = lhs_setting,
      rhs_setting = rhs_setting
    ) %>%
    relocate(comparison_set, patient_id, minobs, pm, lhs_label, rhs_label, lhs_setting, rhs_setting, metric)

  merged_landscape <- merge_landscape_pair(lhs_bundle$landscape, rhs_bundle$landscape)
  landscape_group_tbl <- bind_rows(
    summarize_landscape_groups(merged_landscape, "lhs_mean", "rhs_mean", "landscape_mean"),
    summarize_landscape_groups(merged_landscape, "lhs_median", "rhs_median", "landscape_median"),
    summarize_landscape_groups(merged_landscape, "lhs_sd", "rhs_sd", "landscape_sd")
  ) %>%
    mutate(
      comparison_set = comparison_set,
      patient_id = patient_id,
      minobs = minobs,
      pm = pm,
      lhs_label = lhs_label,
      rhs_label = rhs_label,
      lhs_setting = lhs_setting,
      rhs_setting = rhs_setting
    ) %>%
    relocate(comparison_set, patient_id, minobs, pm, lhs_label, rhs_label, lhs_setting, rhs_setting, metric, state_group)

  top_shift_tbl <- top_landscape_shifts(merged_landscape, top_n = top_n) %>%
    mutate(
      comparison_set = comparison_set,
      patient_id = patient_id,
      minobs = minobs,
      pm = pm,
      lhs_label = lhs_label,
      rhs_label = rhs_label,
      lhs_setting = lhs_setting,
      rhs_setting = rhs_setting
    ) %>%
    relocate(comparison_set, patient_id, minobs, pm, lhs_label, rhs_label, lhs_setting, rhs_setting)

  list(
    component = component_tbl,
    sign = sign_tbl,
    landscape_group = landscape_group_tbl,
    top_shift = top_shift_tbl
  )
}

build_pairwise_comparisons <- function(results_tbl,
                                       pair_values,
                                       setting_col,
                                       comparison_set,
                                       top_n = 15L,
                                       include_posterior = TRUE) {
  if (!nrow(results_tbl) || length(pair_values) < 2L) {
    return(list(
      component = tibble(),
      sign = tibble(),
      landscape_group = tibble(),
      top_shift = tibble()
    ))
  }

  pair_defs <- combn(pair_values, 2, simplify = FALSE)
  all_component <- list()
  all_sign <- list()
  all_group <- list()
  all_shift <- list()

  for (pair in pair_defs) {
    lhs_val <- pair[[1]]
    rhs_val <- pair[[2]]

    lhs_idx <- results_tbl %>%
      filter(status == "ok", .data[[setting_col]] == lhs_val) %>%
      transmute(
        patient_id,
        minobs,
        pm,
        lhs_setting = .data[[setting_col]],
        lhs_label = if (setting_col == "nn_prior_grid_n") paste0("grid_", .data[[setting_col]]) else as.character(.data[[setting_col]]),
        outdir_lhs = outdir
      )

    rhs_idx <- results_tbl %>%
      filter(status == "ok", .data[[setting_col]] == rhs_val) %>%
      transmute(
        patient_id,
        minobs,
        pm,
        rhs_setting = .data[[setting_col]],
        rhs_label = if (setting_col == "nn_prior_grid_n") paste0("grid_", .data[[setting_col]]) else as.character(.data[[setting_col]]),
        outdir_rhs = outdir
      )

    joined <- inner_join(lhs_idx, rhs_idx, by = c("patient_id", "minobs", "pm"))
    if (!nrow(joined)) {
      next
    }

    pair_res <- lapply(seq_len(nrow(joined)), function(i) {
      rr <- joined[i, , drop = FALSE]
      compare_fit_pair(
        lhs_dir = rr$outdir_lhs,
        rhs_dir = rr$outdir_rhs,
        patient_id = rr$patient_id,
        minobs = rr$minobs,
        pm = rr$pm,
        lhs_label = rr$lhs_label,
        rhs_label = rr$rhs_label,
        lhs_setting = rr$lhs_setting,
        rhs_setting = rr$rhs_setting,
        comparison_set = comparison_set,
        top_n = top_n,
        include_posterior = include_posterior
      )
    })

    all_component <- c(all_component, lapply(pair_res, `[[`, "component"))
    all_sign <- c(all_sign, lapply(pair_res, `[[`, "sign"))
    all_group <- c(all_group, lapply(pair_res, `[[`, "landscape_group"))
    all_shift <- c(all_shift, lapply(pair_res, `[[`, "top_shift"))
  }

  list(
    component = bind_rows(all_component),
    sign = bind_rows(all_sign),
    landscape_group = bind_rows(all_group),
    top_shift = bind_rows(all_shift)
  )
}

compute_grid_recommendation <- function(grid_results_tbl,
                                        grid_component_tbl,
                                        reference_grid_n,
                                        runtime_weight = 0.5) {
  if (!nrow(grid_results_tbl)) {
    return(tibble())
  }

  runtime_tbl <- grid_results_tbl %>%
    filter(status == "ok") %>%
    group_by(nn_prior_grid_n) %>%
    summarise(
      success_n = n(),
      mean_xval_r2 = mean(xval_r2, na.rm = TRUE),
      median_elapsed_sec = median(elapsed_sec, na.rm = TRUE),
      total_lambda_endpoint_warnings = sum(lambda_endpoint_warning_count, na.rm = TRUE),
      .groups = "drop"
    )

  ref_component_tbl <- grid_component_tbl %>%
    filter(rhs_setting == reference_grid_n, metric %in% c("landscape_mean", "bootstrap_nn_fitness", "xval_r2")) %>%
    group_by(lhs_setting, metric) %>%
    summarise(
      mean_abs_diff = mean(mean_abs_diff, na.rm = TRUE),
      max_abs_diff = max(max_abs_diff, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(lhs_setting = as.integer(lhs_setting)) %>%
    pivot_wider(
      names_from = metric,
      values_from = c(mean_abs_diff, max_abs_diff),
      names_sep = "__"
    ) %>%
    rename(nn_prior_grid_n = lhs_setting)

  recommendation_tbl <- tibble(nn_prior_grid_n = sort(unique(grid_results_tbl$nn_prior_grid_n))) %>%
    left_join(runtime_tbl, by = "nn_prior_grid_n") %>%
    left_join(ref_component_tbl, by = "nn_prior_grid_n")

  needed_cols <- c(
    "mean_abs_diff__landscape_mean",
    "mean_abs_diff__bootstrap_nn_fitness",
    "mean_abs_diff__xval_r2",
    "max_abs_diff__landscape_mean",
    "max_abs_diff__bootstrap_nn_fitness",
    "max_abs_diff__xval_r2"
  )
  for (col in needed_cols) {
    if (!col %in% names(recommendation_tbl)) {
      recommendation_tbl[[col]] <- NA_real_
    }
  }

  recommendation_tbl <- recommendation_tbl %>%
    mutate(
      mean_abs_diff__landscape_mean = ifelse(nn_prior_grid_n == reference_grid_n, 0, mean_abs_diff__landscape_mean),
      mean_abs_diff__bootstrap_nn_fitness = ifelse(nn_prior_grid_n == reference_grid_n, 0, mean_abs_diff__bootstrap_nn_fitness),
      mean_abs_diff__xval_r2 = ifelse(nn_prior_grid_n == reference_grid_n, 0, mean_abs_diff__xval_r2)
    )

  max_success_n <- max(recommendation_tbl$success_n, na.rm = TRUE)
  eligible_tbl <- recommendation_tbl %>% filter(success_n == max_success_n)
  if (!nrow(eligible_tbl)) {
    return(recommendation_tbl)
  }

  scale_landscape <- max(eligible_tbl$mean_abs_diff__landscape_mean, na.rm = TRUE)
  scale_nn <- max(eligible_tbl$mean_abs_diff__bootstrap_nn_fitness, na.rm = TRUE)
  scale_xval <- max(eligible_tbl$mean_abs_diff__xval_r2, na.rm = TRUE)
  min_runtime <- min(eligible_tbl$median_elapsed_sec, na.rm = TRUE)

  if (!is.finite(scale_landscape) || scale_landscape <= 0) {
    scale_landscape <- 1
  }
  if (!is.finite(scale_nn) || scale_nn <= 0) {
    scale_nn <- 1
  }
  if (!is.finite(scale_xval) || scale_xval <= 0) {
    scale_xval <- 1
  }
  if (!is.finite(min_runtime) || min_runtime <= 0) {
    min_runtime <- 1
  }

  recommendation_tbl <- recommendation_tbl %>%
    mutate(
      landscape_relative = coalesce_value(mean_abs_diff__landscape_mean / scale_landscape, 0),
      nn_relative = coalesce_value(mean_abs_diff__bootstrap_nn_fitness / scale_nn, 0),
      xval_relative = coalesce_value(mean_abs_diff__xval_r2 / scale_xval, 0),
      runtime_relative = coalesce_value(median_elapsed_sec / min_runtime, 1),
      efficiency_score = landscape_relative + nn_relative + xval_relative + runtime_weight * runtime_relative
    ) %>%
    arrange(desc(success_n), efficiency_score, nn_prior_grid_n) %>%
    mutate(is_recommended = row_number() == 1L)

  recommendation_tbl
}

coalesce_value <- function(x, y) {
  ifelse(is.na(x), y, x)
}
```

## Build ALFA-K Inputs

```{r build-inputs}
meta_tbl <- readxl::read_xlsx(meta_path)
input_index_tbl <- build_benchmark_inputs(
  meta_tbl = meta_tbl,
  base_dir = data_dir,
  input_dir = input_dir,
  rebuild_inputs = rebuild_inputs_use,
  patient_subset = patient_subset_use
)
save_table_bundle(input_index_tbl, file.path(tables_dir, "benchmark_input_index"))

input_index_tbl
```

```{r input-summary}
render_tbl(
  input_index_tbl %>%
    select(patient_id, n_timepoints, n_karyotypes, total_cells, delta_time_labels, stage_sequence),
  caption = "Benchmark input summary"
)
```

## Round 1: `nn_prior_grid_n` Sweep

```{r round1-tasks}
grid_tasks_tbl <- build_grid_tasks(
  input_index_tbl = input_index_tbl,
  fit_root = grid_fit_dir,
  minobs_values = minobs_values_use,
  pm_values = pm_values_use,
  grid_values = grid_values_use,
  nboot = nboot_use,
  n0 = n0_use,
  nb = nb_use,
  benchmark_seed = benchmark_seed_use,
  correct_efflux = correct_efflux_use,
  force_refit = force_refit_round1
)
save_table_bundle(grid_tasks_tbl, file.path(tables_dir, "grid_round_tasks"))

grid_tasks_tbl %>%
  count(nn_prior_grid_n, name = "n_tasks")
```

```{r round1-run}
grid_results_path <- file.path(tables_dir, "grid_round_fit_results")

if (run_round1_use) {
  grid_results_tbl <- run_task_table_parallel(grid_tasks_tbl)
  save_table_bundle(grid_results_tbl, grid_results_path)
} else {
  grid_results_tbl <- load_saved_table(grid_results_path)
  if (is.null(grid_results_tbl)) {
    grid_results_tbl <- tibble()
  }
}

grid_results_tbl
```

```{r round1-summary}
grid_fit_summary_tbl <- summarize_fit_results(grid_results_tbl, group_cols = c("nn_prior_grid_n"))
save_table_bundle(grid_fit_summary_tbl, file.path(tables_dir, "grid_round_fit_summary"))

if (nrow(grid_fit_summary_tbl)) {
  render_tbl(grid_fit_summary_tbl, caption = "Round 1 fit summary by nn_prior_grid_n")
} else {
  render_tbl(empty_note_tbl("No Round 1 fit results are available yet."), caption = "Round 1 fit summary")
}
```

## Round 1: Grid Comparisons and Recommendation

```{r round1-compare}
grid_pair_results <- build_pairwise_comparisons(
  results_tbl = grid_results_tbl,
  pair_values = grid_values_use,
  setting_col = "nn_prior_grid_n",
  comparison_set = "grid",
  top_n = top_shift_n_use,
  include_posterior = include_posterior_use
)

save_table_bundle(grid_pair_results$component, file.path(tables_dir, "grid_pair_component_summary"))
save_table_bundle(grid_pair_results$sign, file.path(tables_dir, "grid_pair_sign_summary"))
save_table_bundle(grid_pair_results$landscape_group, file.path(tables_dir, "grid_pair_landscape_group_summary"))
save_table_bundle(grid_pair_results$top_shift, file.path(tables_dir, "grid_pair_top_landscape_shifts"))

grid_pair_overview_tbl <- if (nrow(grid_pair_results$component)) {
  grid_pair_results$component %>%
    filter(metric %in% c("landscape_mean", "bootstrap_nn_fitness", "xval_r2")) %>%
    group_by(lhs_label, rhs_label, metric) %>%
    summarise(
      n_pairs = n(),
      mean_abs_diff = mean(mean_abs_diff, na.rm = TRUE),
      max_abs_diff = max(max_abs_diff, na.rm = TRUE),
      mean_correlation = mean(correlation, na.rm = TRUE),
      .groups = "drop"
    )
} else {
  tibble()
}
save_table_bundle(grid_pair_overview_tbl, file.path(tables_dir, "grid_pair_overview"))

grid_recommendation_tbl <- compute_grid_recommendation(
  grid_results_tbl = grid_results_tbl,
  grid_component_tbl = grid_pair_results$component,
  reference_grid_n = reference_grid_n
)
save_table_bundle(grid_recommendation_tbl, file.path(tables_dir, "grid_recommendation"))

best_grid_n <- if (nrow(grid_recommendation_tbl)) {
  grid_recommendation_tbl$nn_prior_grid_n[which(grid_recommendation_tbl$is_recommended)][1]
} else {
  NA_integer_
}
saveRDS(best_grid_n, file.path(cache_dir, "best_grid_n.rds"))

list(
  pair_overview = grid_pair_overview_tbl,
  recommendation = grid_recommendation_tbl,
  best_grid_n = best_grid_n
)
```

```{r round1-compare-render}
if (nrow(grid_pair_overview_tbl)) {
  render_tbl(grid_pair_overview_tbl, caption = "Round 1 pairwise comparison overview")
} else {
  render_tbl(empty_note_tbl("No successful Round 1 pairwise comparisons are available yet."), caption = "Round 1 pairwise comparison overview")
}
```

```{r round1-recommendation-render}
if (nrow(grid_recommendation_tbl)) {
  render_tbl(grid_recommendation_tbl, caption = "Global nn_prior_grid_n recommendation summary")
} else {
  render_tbl(empty_note_tbl("No grid recommendation could be computed because Round 1 results are unavailable."), caption = "Global nn_prior_grid_n recommendation summary")
}
```

```{r round1-best-grid-text, results='asis'}
if (is.finite(best_grid_n)) {
  cat(
    sprintf(
      "**Recommended benchmark-wide `nn_prior_grid_n`: `%d`**\n\n",
      best_grid_n
    )
  )
} else {
  cat("**Recommended benchmark-wide `nn_prior_grid_n`: unavailable**\n\n")
}
```

```{r round1-top-shifts}
grid_top_shift_preview_tbl <- if (nrow(grid_pair_results$top_shift)) {
  grid_pair_results$top_shift %>%
    group_by(lhs_label, rhs_label) %>%
    slice_head(n = min(10L, top_shift_n_use)) %>%
    ungroup()
} else {
  tibble()
}

if (nrow(grid_top_shift_preview_tbl)) {
  render_tbl(grid_top_shift_preview_tbl, caption = "Top landscape shifts for Round 1 grid comparisons")
} else {
  render_tbl(empty_note_tbl("No top-shift table is available for Round 1."), caption = "Top landscape shifts for Round 1 grid comparisons")
}
```

```{r round1-figures, eval=render_figures_use}
if (nrow(grid_fit_summary_tbl)) {
  p_grid_runtime <- ggplot(grid_results_tbl %>% filter(status == "ok"), aes(x = factor(nn_prior_grid_n), y = elapsed_sec)) +
    geom_boxplot(outlier.shape = NA, fill = "#D9EAF7", color = "#2C4C7A") +
    geom_jitter(width = 0.15, alpha = 0.5, size = 1.6, color = "#2C4C7A") +
    labs(
      title = "Round 1 runtime by nn_prior_grid_n",
      x = "nn_prior_grid_n",
      y = "Elapsed seconds"
    ) +
    theme_bw(base_size = 12)
  print(p_grid_runtime)
  ggsave(file.path(figures_dir, "round1_runtime_by_grid.png"), p_grid_runtime, width = 8, height = 5, dpi = 150)

  p_grid_xval <- ggplot(grid_results_tbl %>% filter(status == "ok"), aes(x = factor(nn_prior_grid_n), y = xval_r2)) +
    geom_boxplot(outlier.shape = NA, fill = "#F4E3B2", color = "#7A5C00") +
    geom_jitter(width = 0.15, alpha = 0.5, size = 1.6, color = "#7A5C00") +
    labs(
      title = "Round 1 xval by nn_prior_grid_n",
      x = "nn_prior_grid_n",
      y = "xval_r2"
    ) +
    theme_bw(base_size = 12)
  print(p_grid_xval)
  ggsave(file.path(figures_dir, "round1_xval_by_grid.png"), p_grid_xval, width = 8, height = 5, dpi = 150)
}
```

## Round 2: `nn_prior` Benchmark at the Selected Grid

```{r round2-tasks}
if (is.finite(best_grid_n)) {
  prior_tasks_tbl <- build_prior_tasks(
    input_index_tbl = input_index_tbl,
    fit_root = prior_fit_dir,
    minobs_values = minobs_values_use,
    pm_values = pm_values_use,
    best_grid_n = best_grid_n,
    priors = round2_priors_use,
    nboot = nboot_use,
    n0 = n0_use,
    nb = nb_use,
    benchmark_seed = benchmark_seed_use,
    correct_efflux = correct_efflux_use,
    force_refit = force_refit_round2
  )
} else {
  prior_tasks_tbl <- tibble()
}

save_table_bundle(prior_tasks_tbl, file.path(tables_dir, "prior_round_tasks"))
prior_tasks_tbl
```

```{r round2-run}
prior_results_path <- file.path(tables_dir, "prior_round_fit_results")

if (run_round2_use && nrow(prior_tasks_tbl)) {
  prior_results_new_tbl <- run_task_table_parallel(prior_tasks_tbl)
  save_table_bundle(prior_results_new_tbl, prior_results_path)
} else {
  prior_results_new_tbl <- load_saved_table(prior_results_path)
  if (is.null(prior_results_new_tbl)) {
    prior_results_new_tbl <- tibble()
  }
}

empirical_censored_reuse_tbl <- if (nrow(grid_results_tbl) && is.finite(best_grid_n)) {
  grid_results_tbl %>%
    filter(status == "ok", nn_prior == "empirical_censored", nn_prior_grid_n == best_grid_n) %>%
    mutate(round_label = "prior_round_reused")
} else {
  tibble()
}

prior_results_all_tbl <- bind_rows(
  empirical_censored_reuse_tbl,
  prior_results_new_tbl
)
if (nrow(prior_results_all_tbl)) {
  prior_results_all_tbl <- prior_results_all_tbl %>%
    arrange(factor(patient_id, levels = sort_pid_levels(patient_id)), minobs, pm, nn_prior)
}

save_table_bundle(prior_results_all_tbl, file.path(tables_dir, "prior_round_all_results"))
prior_results_all_tbl
```

```{r round2-summary}
prior_fit_summary_tbl <- summarize_fit_results(prior_results_all_tbl, group_cols = c("nn_prior"))
save_table_bundle(prior_fit_summary_tbl, file.path(tables_dir, "prior_round_fit_summary"))

if (nrow(prior_fit_summary_tbl)) {
  render_tbl(prior_fit_summary_tbl, caption = "Round 2 fit summary by nn_prior")
} else {
  render_tbl(empty_note_tbl("No Round 2 fit results are available yet."), caption = "Round 2 fit summary")
}
```

## Round 2: Prior Comparisons

```{r round2-compare}
prior_pair_values_use <- c("empirical_censored", round2_priors_use)

prior_pair_results <- build_pairwise_comparisons(
  results_tbl = prior_results_all_tbl,
  pair_values = prior_pair_values_use,
  setting_col = "nn_prior",
  comparison_set = "prior",
  top_n = top_shift_n_use,
  include_posterior = include_posterior_use
)

save_table_bundle(prior_pair_results$component, file.path(tables_dir, "prior_pair_component_summary"))
save_table_bundle(prior_pair_results$sign, file.path(tables_dir, "prior_pair_sign_summary"))
save_table_bundle(prior_pair_results$landscape_group, file.path(tables_dir, "prior_pair_landscape_group_summary"))
save_table_bundle(prior_pair_results$top_shift, file.path(tables_dir, "prior_pair_top_landscape_shifts"))

prior_pair_overview_tbl <- if (nrow(prior_pair_results$component)) {
  prior_pair_results$component %>%
    filter(metric %in% c("landscape_mean", "bootstrap_nn_fitness", "xval_r2")) %>%
    group_by(lhs_label, rhs_label, metric) %>%
    summarise(
      n_pairs = n(),
      mean_abs_diff = mean(mean_abs_diff, na.rm = TRUE),
      max_abs_diff = max(max_abs_diff, na.rm = TRUE),
      mean_correlation = mean(correlation, na.rm = TRUE),
      .groups = "drop"
    )
} else {
  tibble()
}
save_table_bundle(prior_pair_overview_tbl, file.path(tables_dir, "prior_pair_overview"))

prior_pair_overview_tbl
```

```{r round2-overview-render}
if (nrow(prior_pair_overview_tbl)) {
  render_tbl(prior_pair_overview_tbl, caption = "Round 2 pairwise comparison overview")
} else {
  render_tbl(empty_note_tbl("No successful Round 2 pairwise comparisons are available yet."), caption = "Round 2 pairwise comparison overview")
}
```

```{r round2-top-shifts}
prior_top_shift_preview_tbl <- if (nrow(prior_pair_results$top_shift)) {
  prior_pair_results$top_shift %>%
    group_by(lhs_label, rhs_label) %>%
    slice_head(n = min(10L, top_shift_n_use)) %>%
    ungroup()
} else {
  tibble()
}

if (nrow(prior_top_shift_preview_tbl)) {
  render_tbl(prior_top_shift_preview_tbl, caption = "Top landscape shifts for Round 2 prior comparisons")
} else {
  render_tbl(empty_note_tbl("No top-shift table is available for Round 2."), caption = "Top landscape shifts for Round 2 prior comparisons")
}
```

```{r round2-figures, eval=render_figures_use}
if (nrow(prior_fit_summary_tbl)) {
  p_prior_runtime <- ggplot(prior_results_all_tbl %>% filter(status == "ok"), aes(x = nn_prior, y = elapsed_sec)) +
    geom_boxplot(outlier.shape = NA, fill = "#E8D8E8", color = "#633A63") +
    geom_jitter(width = 0.15, alpha = 0.5, size = 1.6, color = "#633A63") +
    labs(
      title = paste0("Round 2 runtime by nn_prior (grid = ", best_grid_n, ")"),
      x = "nn_prior",
      y = "Elapsed seconds"
    ) +
    theme_bw(base_size = 12)
  print(p_prior_runtime)
  ggsave(file.path(figures_dir, "round2_runtime_by_prior.png"), p_prior_runtime, width = 8, height = 5, dpi = 150)

  p_prior_xval <- ggplot(prior_results_all_tbl %>% filter(status == "ok"), aes(x = nn_prior, y = xval_r2)) +
    geom_boxplot(outlier.shape = NA, fill = "#D8EBCF", color = "#31572C") +
    geom_jitter(width = 0.15, alpha = 0.5, size = 1.6, color = "#31572C") +
    labs(
      title = paste0("Round 2 xval by nn_prior (grid = ", best_grid_n, ")"),
      x = "nn_prior",
      y = "xval_r2"
    ) +
    theme_bw(base_size = 12)
  print(p_prior_xval)
  ggsave(file.path(figures_dir, "round2_xval_by_prior.png"), p_prior_xval, width = 8, height = 5, dpi = 150)
}
```

## Artifact Index

```{r artifact-index}
artifact_index_tbl <- tibble(
  artifact = c(
    "benchmark_input_index",
    "grid_round_tasks",
    "grid_round_fit_results",
    "grid_round_fit_summary",
    "grid_pair_component_summary",
    "grid_pair_landscape_group_summary",
    "grid_pair_sign_summary",
    "grid_pair_top_landscape_shifts",
    "grid_recommendation",
    "prior_round_tasks",
    "prior_round_fit_results",
    "prior_round_all_results",
    "prior_round_fit_summary",
    "prior_pair_component_summary",
    "prior_pair_landscape_group_summary",
    "prior_pair_sign_summary",
    "prior_pair_top_landscape_shifts"
  ),
  path = file.path(
    tables_dir,
    c(
      "benchmark_input_index.tsv",
      "grid_round_tasks.tsv",
      "grid_round_fit_results.tsv",
      "grid_round_fit_summary.tsv",
      "grid_pair_component_summary.tsv",
      "grid_pair_landscape_group_summary.tsv",
      "grid_pair_sign_summary.tsv",
      "grid_pair_top_landscape_shifts.tsv",
      "grid_recommendation.tsv",
      "prior_round_tasks.tsv",
      "prior_round_fit_results.tsv",
      "prior_round_all_results.tsv",
      "prior_round_fit_summary.tsv",
      "prior_pair_component_summary.tsv",
      "prior_pair_landscape_group_summary.tsv",
      "prior_pair_sign_summary.tsv",
      "prior_pair_top_landscape_shifts.tsv"
    )
  )
)
save_table_bundle(artifact_index_tbl, file.path(tables_dir, "artifact_index"))

render_tbl(artifact_index_tbl, caption = "Benchmark artifact index")
```
