---
title: "ALFA-K Benchmark"
subtitle: "Single-round benchmark for nn_prior comparison"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    theme: flatly
params:
  patient_subset: !r character(0)
  minobs_values: [5, 10, 20]
  pm_values: [5e-05]
  parameter_labels: ["nn_prior_empirical_censored", "nn_prior_empirical_censored_weighted", "nn_prior_none", "nn_prior_empirical"]
  n_cores: 18
  nboot: 45
  n0: 100000
  nb: 10000000
  benchmark_seed: 31415926
  correct_efflux: true
  nn_prior_grid_n: 161
  nn_prior_fit_subset: "hybrid"
  nn_prior_zero_exposure_quantile: 0.10
  nn_prior_zero_weight_scale: 0.50
  nn_prior_zero_weight_cap_ratio: 0.05
  nn_prior_zero_birth_fallback_weight: 0.02
  include_posterior_comparison: true
  top_shift_n: 15
  force_refit: false
  rebuild_inputs: false
  run_benchmark: true
  render_figures: true
  focus_pids: ["P2"]
  focus_pid: ""
  focus_minobs: 20
  focus_pm: 5e-05
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6,
  dpi = 150
)

resolve_repo_dir_light <- function() {
  candidates <- unique(c(
    tryCatch(dirname(knitr::current_input(dir = TRUE)), error = function(e) NA_character_),
    getwd()
  ))

  for (cand in candidates) {
    if (is.na(cand) || !nzchar(cand)) {
      next
    }
    cand_norm <- normalizePath(cand, winslash = "/", mustWork = FALSE)
    if (file.exists(file.path(cand_norm, "DESCRIPTION")) &&
        dir.exists(file.path(cand_norm, "benchmark"))) {
      return(cand_norm)
    }
    parent2 <- normalizePath(file.path(cand_norm, "..", ".."), winslash = "/", mustWork = FALSE)
    if (file.exists(file.path(parent2, "DESCRIPTION")) &&
        dir.exists(file.path(parent2, "benchmark"))) {
      return(parent2)
    }
  }

  stop("Could not locate the alfakR repository root from the current knitting context.")
}

repo_dir <- resolve_repo_dir_light()
sys.source(file.path(repo_dir, "benchmark", "scr", "run_benchmark_pipeline.R"), envir = knitr::knit_global())
benchmark_run <- run_benchmark_pipeline_entry(params = params, repo_dir = repo_dir)

ctx <- benchmark_run$ctx
focus_results <- benchmark_run$focus_results
```

```{r html-table-style, results='asis'}
if (knitr::is_html_output()) {
  cat(
    "<style>\n",
    "table.three-line-table {\n",
    "  width: auto;\n",
    "  margin-left: 0;\n",
    "  border-collapse: collapse;\n",
    "  border-top: 2px solid #222 !important;\n",
    "  border-bottom: 2px solid #222 !important;\n",
    "}\n",
    "table.three-line-table thead th {\n",
    "  border-bottom: 1px solid #222 !important;\n",
    "}\n",
    "table.three-line-table th,\n",
    "table.three-line-table td {\n",
    "  border-left: none !important;\n",
    "  border-right: none !important;\n",
    "  padding: 0.35em 0.65em;\n",
    "  background: transparent !important;\n",
    "}\n",
    "</style>\n",
    sep = ""
  )
}
```

## Overview

This report renders benchmark artifacts generated by the modular pipeline under
`benchmark/R/*.R`. The R Markdown file only orchestrates the pipeline and
displays the saved results.

- Fixed benchmark-wide `nn_prior_grid_n`: `r ctx$selected_grid_n_use`
- Compared `nn_prior` labels: `r paste(ctx$parameter_levels_use, collapse = ", ")`
- Focus samples: `r paste(ctx$focus_pids_use, collapse = ", ")`
- Focus comparison settings: `minobs = r ctx$focus_minobs_use`, `pm = r pm_to_label(ctx$focus_pm_use)`

## Input Summary

```{r input-summary}
print(
  render_tbl(
    benchmark_run$input_index_tbl %>%
      dplyr::select(patient_id, n_timepoints, n_karyotypes, total_cells, delta_time_labels, stage_sequence),
    caption = "Benchmark input summary"
  )
)
```

## Parameter Benchmark

```{r parameter-summary}
parameter_fit_summary_tbl <- benchmark_run$parameter_fit_summary_tbl
if (nrow(parameter_fit_summary_tbl)) {
  print(render_tbl(parameter_fit_summary_tbl, caption = "Fit summary by parameter label"))
} else {
  print(render_tbl(empty_note_tbl("No parameter-label fit results are available yet."), caption = "Fit summary by parameter label"))
}
```

## Selected Fits for All-Sample Panels

```{r parameter-selected-fits-render}
parameter_selected_fit_tbl <- benchmark_run$parameter_selected_fit_tbl
parameter_selected_fit_render_tbl <- if (nrow(parameter_selected_fit_tbl)) {
  parameter_selected_fit_tbl %>%
    dplyr::transmute(
      parameter_label,
      patient_id,
      nn_prior,
      minobs,
      pm,
      xval_r2,
      elapsed_sec,
      cached
    )
} else {
  tibble::tibble()
}

if (nrow(parameter_selected_fit_render_tbl)) {
  print(
    render_tbl(
      parameter_selected_fit_render_tbl,
      caption = "Best successful fit per patient within each parameter label, used for all-sample heatmaps"
    )
  )
} else {
  print(
    render_tbl(
      empty_note_tbl("No successful patient-level fits are available for the all-sample heatmaps yet."),
      caption = "Best successful fit per patient within each parameter label"
    )
  )
}
```

## All-Sample Beneficial Karyotype Heatmaps

```{r parameter-beneficial-artifacts-render}
parameter_beneficial_artifact_tbl <- benchmark_run$parameter_beneficial_artifact_tbl
parameter_beneficial_summary_render_tbl <- if (nrow(parameter_beneficial_artifact_tbl)) {
  parameter_beneficial_artifact_tbl %>%
    dplyr::select(parameter_label, n_patients)
} else {
  tibble::tibble()
}

if (nrow(parameter_beneficial_summary_render_tbl)) {
  print(
    render_tbl(
      parameter_beneficial_summary_render_tbl,
      caption = "All-sample beneficial-heatmap coverage by parameter label"
    )
  )
} else {
  print(
    render_tbl(
      empty_note_tbl("No all-sample beneficial heatmaps are available yet."),
      caption = "All-sample beneficial-heatmap coverage by parameter label"
    )
  )
}
```

```{r parameter-beneficial-heatmaps-display, results='asis'}
if (nrow(parameter_beneficial_artifact_tbl)) {
  for (i in seq_len(nrow(parameter_beneficial_artifact_tbl))) {
    rr <- parameter_beneficial_artifact_tbl[i, , drop = FALSE]
    cat("### ", rr$parameter_label, "\n\n", sep = "")
    if (is.na(rr$heatmap_png_path) || !nzchar(rr$heatmap_png_path) || !file.exists(rr$heatmap_png_path)) {
      cat("No all-sample beneficial heatmap is available for this parameter label.\n\n")
      next
    }
    emit_report_image(
      rr$heatmap_png_path,
      alt = paste("All-sample beneficial heatmap for", rr$parameter_label)
    )
  }
} else {
  cat("No all-sample beneficial heatmaps are available yet.\n")
}
```

## Parameter-Label Pairwise Comparisons

```{r parameter-overview-render}
parameter_pair_overview_tbl <- benchmark_run$parameter_pair_overview_tbl
if (nrow(parameter_pair_overview_tbl)) {
  print(render_tbl(parameter_pair_overview_tbl, caption = "Pairwise comparison overview by parameter label"))
} else {
  print(render_tbl(empty_note_tbl("No successful parameter-label pairwise comparisons are available yet."), caption = "Pairwise comparison overview by parameter label"))
}
```

```{r parameter-top-shifts-render}
parameter_top_shift_preview_tbl <- benchmark_run$parameter_top_shift_preview_tbl
if (nrow(parameter_top_shift_preview_tbl)) {
  print(render_tbl(parameter_top_shift_preview_tbl, caption = "Top landscape shifts for parameter-label comparisons"))
} else {
  print(render_tbl(empty_note_tbl("No top-shift table is available for the parameter-label benchmark."), caption = "Top landscape shifts for parameter-label comparisons"))
}
```

## Global Landscape and Beneficial Comparison

```{r parameter-global-render}
if (nrow(benchmark_run$parameter_global_landscape_overview_tbl)) {
  print(render_tbl(benchmark_run$parameter_global_landscape_overview_tbl, caption = "Benchmark-wide landscape and beneficial-proportion comparison summary"))
} else {
  print(render_tbl(empty_note_tbl("No benchmark-wide landscape summary is available yet."), caption = "Benchmark-wide landscape summary"))
}

if (nrow(benchmark_run$parameter_global_landscape_state_tbl)) {
  print(render_tbl(benchmark_run$parameter_global_landscape_state_tbl, caption = "Benchmark-wide landscape mean differences split by fq / nn / other states"))
} else {
  print(render_tbl(empty_note_tbl("No state-group landscape summary is available for the benchmark yet."), caption = "Benchmark-wide landscape mean by state group"))
}
```

## Benchmark Figures

```{r parameter-figures-display, results='asis'}
parameter_figure_paths <- benchmark_run$parameter_figure_paths
figure_titles <- c(
  runtime = "Runtime by Parameter Label",
  xval = "xval by Parameter Label",
  global_diff = "Benchmark-wide Mean Absolute Differences",
  global_state_diff = "Benchmark-wide Landscape Mean Differences by State Group"
)

for (nm in names(parameter_figure_paths)) {
  cat("### ", figure_titles[[nm]], "\n\n", sep = "")
  emit_report_image(parameter_figure_paths[[nm]], alt = figure_titles[[nm]])
}
```

## Focus Comparisons

```{r focus-sections, results='asis'}
if (!length(focus_results)) {
  cat("No focus sample is configured.\n")
} else {
  for (focus_pid in names(focus_results)) {
    focus_res <- focus_results[[focus_pid]]
    cat("### ", focus_pid, "\n\n", sep = "")
    cat(
      sprintf(
        "Focus comparison uses patient `%s`, `minobs = %d`, and `pm = %s`.\n\n",
        focus_pid,
        ctx$focus_minobs_use,
        pm_to_label(ctx$focus_pm_use)
      )
    )

    if (nrow(focus_res$focus_fit_index_tbl)) {
      print(
        render_tbl(
          focus_res$focus_fit_index_tbl %>%
            dplyr::transmute(
              parameter_label,
              nn_prior,
              status,
              xval_r2,
              elapsed_sec,
              cached,
              warning_count,
              error_message
            ),
          caption = paste("Focus fit index:", focus_pid)
        )
      )
    } else {
      print(render_tbl(empty_note_tbl("The focus patient combination is not available in the benchmark results yet."), caption = paste("Focus fit index:", focus_pid)))
    }

    if (nrow(focus_res$focus_landscape_summary_tbl)) {
      print(render_tbl(focus_res$focus_landscape_summary_tbl, caption = paste("Focus landscape summary:", focus_pid)))
    } else {
      print(render_tbl(empty_note_tbl("No focus-patient landscape summary is available yet."), caption = paste("Focus landscape summary:", focus_pid)))
    }

    cat("#### Focus UMAP\n\n")
    if (nrow(focus_res$focus_umap_artifact_tbl)) {
      for (i in seq_len(nrow(focus_res$focus_umap_artifact_tbl))) {
        rr <- focus_res$focus_umap_artifact_tbl[i, , drop = FALSE]
        cat("##### ", rr$parameter_label, "\n\n", sep = "")
        if (is.na(rr$png_path) || !nzchar(rr$png_path) || !file.exists(rr$png_path)) {
          cat("No UMAP is available for this parameter label.\n\n")
          next
        }
        emit_report_image(rr$png_path, alt = paste("Focus patient UMAP for", focus_pid, rr$parameter_label))
      }
    } else {
      cat("No focus-patient UMAP plots are available yet.\n\n")
    }

    if (nrow(focus_res$focus_pairwise_component_tbl)) {
      print(render_tbl(focus_res$focus_pairwise_component_tbl, caption = paste("Focus pairwise landscape, xval, and beneficial summary:", focus_pid)))
    } else {
      print(render_tbl(empty_note_tbl("No pairwise focus-patient comparison is available yet."), caption = paste("Focus pairwise comparison summary:", focus_pid)))
    }

    if (nrow(focus_res$focus_pairwise_landscape_group_tbl)) {
      print(render_tbl(focus_res$focus_pairwise_landscape_group_tbl, caption = paste("Focus pairwise landscape mean differences by state group:", focus_pid)))
    } else {
      print(render_tbl(empty_note_tbl("No focus-patient state-group landscape summary is available yet."), caption = paste("Focus landscape mean by state group:", focus_pid)))
    }

    if (nrow(focus_res$focus_landscape_variation_tbl)) {
      print(render_tbl(focus_res$focus_landscape_variation_tbl, caption = paste("Focus karyotypes with the largest landscape spread:", focus_pid)))
    } else {
      print(render_tbl(empty_note_tbl("No focus-patient landscape spread table is available yet."), caption = paste("Focus top variable karyotypes:", focus_pid)))
    }

    if (nrow(focus_res$focus_top_shift_tbl)) {
      print(render_tbl(focus_res$focus_top_shift_tbl, caption = paste("Top focus-patient pairwise landscape mean shifts:", focus_pid)))
    } else {
      print(render_tbl(empty_note_tbl("No focus-patient top-shift table is available yet."), caption = paste("Focus top landscape shifts:", focus_pid)))
    }

    if (nrow(focus_res$focus_beneficial_summary_tbl)) {
      print(render_tbl(focus_res$focus_beneficial_summary_tbl, caption = paste("Focus beneficial-karyotype summary:", focus_pid)))
    } else {
      print(render_tbl(empty_note_tbl("No focus-patient beneficial-karyotype summary is available yet."), caption = paste("Focus beneficial-karyotype summary:", focus_pid)))
    }

    if (nrow(focus_res$focus_beneficial_shift_tbl)) {
      print(render_tbl(focus_res$focus_beneficial_shift_tbl, caption = paste("Focus beneficial-proportion shifts:", focus_pid)))
    } else {
      print(render_tbl(empty_note_tbl("No focus-patient beneficial-proportion shift table is available yet."), caption = paste("Focus beneficial-proportion shifts:", focus_pid)))
    }

    cat("#### Focus Figures\n\n")
    emit_report_image(focus_res$focus_figure_paths$density, alt = paste("Focus landscape density for", focus_pid))
    emit_report_image(focus_res$focus_figure_paths$parity, alt = paste("Focus parity plot for", focus_pid))
    emit_report_image(focus_res$focus_figure_paths$beneficial_heatmap, alt = paste("Focus beneficial heatmap for", focus_pid))
  }
}
```

## Artifact Index

```{r artifact-index}
print(render_tbl(benchmark_run$artifact_index_tbl, caption = "Benchmark artifact index"))
```
