---
title: "ALFA-K Benchmark"
subtitle: "Single-round benchmark for nn_prior comparison"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    theme: flatly
params:
  patient_subset: !r character(0)
  minobs_values: [5, 10, 20]
  pm_values: [5e-05]
  parameter_labels: ["nn_prior_empirical_censored", "nn_prior_empirical_censored_weighted", "nn_prior_none", "nn_prior_empirical"]
  n_cores: 18
  nboot: 45
  n0: 100000
  nb: 10000000
  benchmark_seed: 31415926
  correct_efflux: true
  nn_prior_grid_n: 161
  nn_prior_fit_subset: "hybrid"
  nn_prior_zero_exposure_quantile: 0.10
  nn_prior_zero_weight_scale: 0.50
  nn_prior_zero_weight_cap_ratio: 0.05
  nn_prior_zero_birth_fallback_weight: 0.02
  include_posterior_comparison: true
  top_shift_n: 15
  force_refit: false
  rebuild_inputs: false
  run_benchmark: true
  render_figures: true
  focus_pid: "P2"
  focus_minobs: 20
  focus_pm: 5e-05
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = TRUE,
  fig.width = 10,
  fig.height = 6,
  dpi = 150
)
```

```{r html-table-style, echo=FALSE, results='asis'}
if (knitr::is_html_output()) {
  cat(
    "<style>\n",
    "table.three-line-table {\n",
    "  width: auto;\n",
    "  margin-left: 0;\n",
    "  border-collapse: collapse;\n",
    "  border-top: 2px solid #222 !important;\n",
    "  border-bottom: 2px solid #222 !important;\n",
    "}\n",
    "table.three-line-table thead th {\n",
    "  border-bottom: 1px solid #222 !important;\n",
    "}\n",
    "table.three-line-table th,\n",
    "table.three-line-table td {\n",
    "  border-left: none !important;\n",
    "  border-right: none !important;\n",
    "  padding: 0.35em 0.65em;\n",
    "  background: transparent !important;\n",
    "}\n",
    "</style>\n",
    sep = ""
  )
}
```

## Overview

This notebook benchmarks the current local `alfakR` code under `benchmark/data`
using a single-round nn_prior comparison design.

- `nn_prior_grid_n` is fixed to a user-validated benchmark-wide value.
- The benchmark compares the current supported `nn_prior` labels.
- The current code adds `nn_prior_empirical_censored_weighted`, so the
  default benchmark compares:
  `nn_prior_empirical_censored`,
  `nn_prior_empirical_censored_weighted`,
  `nn_prior_none`,
  and `nn_prior_empirical`.
- The report summarizes benchmark-wide landscape differences and produces
  a detailed four-way comparison for one focus patient.

The ALFA-K input construction intentionally follows the Figure S5 reanalysis
pipeline in `cloneidR/analysis/FigureS5/FigureS_UseCase_GBM.Rmd`: metadata are
read from `meta_data.xlsx`, `ID_alt` is preferred over `ID`, transcriptome
profiles are converted into karyotype strings, and the primary-to-recurrent
interval is encoded as `Primary = 1`, `Recurrent = abs(Age_recurrent -
Age_primary) * 360` days.

```{r paths-and-libraries}
suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(knitr)
  library(parallel)
  library(pkgload)
  library(purrr)
  library(readxl)
  library(tibble)
  library(tidyr)
})

resolve_repo_dir <- function() {
  candidates <- unique(c(
    tryCatch(dirname(knitr::current_input(dir = TRUE)), error = function(e) NA_character_),
    getwd()
  ))

  for (cand in candidates) {
    if (is.na(cand) || !nzchar(cand)) {
      next
    }
    cand_norm <- normalizePath(cand, winslash = "/", mustWork = FALSE)
    if (file.exists(file.path(cand_norm, "DESCRIPTION")) &&
        dir.exists(file.path(cand_norm, "benchmark"))) {
      return(cand_norm)
    }
    parent2 <- normalizePath(file.path(cand_norm, "..", ".."), winslash = "/", mustWork = FALSE)
    if (file.exists(file.path(parent2, "DESCRIPTION")) &&
        dir.exists(file.path(parent2, "benchmark"))) {
      return(parent2)
    }
  }

  stop("Could not locate the alfakR repository root from the current knitting context.")
}

repo_dir <- resolve_repo_dir()
benchmark_dir <- file.path(repo_dir, "benchmark")
data_dir <- file.path(benchmark_dir, "data")
meta_path <- file.path(data_dir, "meta_data.xlsx")
results_dir <- file.path(benchmark_dir, "results")
input_dir <- file.path(results_dir, "inputs")
fit_dir <- file.path(results_dir, "fits")
tables_dir <- file.path(results_dir, "tables")
figures_dir <- file.path(results_dir, "figures")
cache_dir <- file.path(results_dir, "cache")

dir.create(results_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(input_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(fit_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(tables_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(figures_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(cache_dir, recursive = TRUE, showWarnings = FALSE)

pkgload::load_all(repo_dir, quiet = TRUE)

patient_subset_use <- unique(trimws(as.character(unlist(params$patient_subset))))
patient_subset_use <- patient_subset_use[nzchar(patient_subset_use)]
if (!length(patient_subset_use)) {
  patient_subset_use <- NULL
}

minobs_values_use <- sort(unique(as.integer(unlist(params$minobs_values))))
minobs_values_use <- minobs_values_use[is.finite(minobs_values_use) & minobs_values_use > 0L]
if (!length(minobs_values_use)) {
  stop("No valid minobs values supplied.")
}

pm_values_use <- sort(unique(as.numeric(unlist(params$pm_values))))
pm_values_use <- pm_values_use[is.finite(pm_values_use) & pm_values_use > 0]
if (!length(pm_values_use)) {
  stop("No valid pm values supplied.")
}

allowed_nn_prior_values <- eval(formals(alfakR::alfak)$nn_prior)
allowed_parameter_labels <- paste0("nn_prior_", allowed_nn_prior_values)
parameter_labels_use <- unique(as.character(unlist(params$parameter_labels)))
parameter_labels_use <- parameter_labels_use[nzchar(parameter_labels_use)]
parameter_labels_use <- ifelse(
  parameter_labels_use %in% allowed_nn_prior_values,
  paste0("nn_prior_", parameter_labels_use),
  parameter_labels_use
)
if (!length(parameter_labels_use) || !all(parameter_labels_use %in% allowed_parameter_labels)) {
  stop(
    "parameter_labels must be chosen from: ",
    paste(allowed_parameter_labels, collapse = ", ")
  )
}

selected_grid_n_use <- suppressWarnings(as.integer(params$nn_prior_grid_n))
if (!is.finite(selected_grid_n_use) || selected_grid_n_use < 3L) {
  stop("nn_prior_grid_n must be a finite integer >= 3.")
}

n_cores_use <- suppressWarnings(as.integer(params$n_cores))
if (!is.finite(n_cores_use) || n_cores_use < 1L) {
  detected_cores <- suppressWarnings(parallel::detectCores(logical = FALSE))
  if (!is.finite(detected_cores) || detected_cores < 1L) {
    detected_cores <- 1L
  }
  n_cores_use <- max(1L, detected_cores - 1L)
}

nboot_use <- as.integer(params$nboot)
n0_use <- as.numeric(params$n0)
nb_use <- as.numeric(params$nb)
benchmark_seed_use <- suppressWarnings(as.integer(params$benchmark_seed))
if (!is.finite(benchmark_seed_use)) {
  stop("benchmark_seed must be a finite integer.")
}
correct_efflux_use <- isTRUE(params$correct_efflux)
nn_prior_fit_subset_use <- match.arg(as.character(params$nn_prior_fit_subset), c("hybrid", "all"))
nn_prior_zero_exposure_quantile_use <- suppressWarnings(as.numeric(params$nn_prior_zero_exposure_quantile))
if (!is.finite(nn_prior_zero_exposure_quantile_use) ||
    nn_prior_zero_exposure_quantile_use < 0 ||
    nn_prior_zero_exposure_quantile_use > 1) {
  stop("nn_prior_zero_exposure_quantile must be a finite number in [0, 1].")
}
nn_prior_zero_weight_scale_use <- suppressWarnings(as.numeric(params$nn_prior_zero_weight_scale))
if (!is.finite(nn_prior_zero_weight_scale_use) ||
    nn_prior_zero_weight_scale_use < 0 ||
    nn_prior_zero_weight_scale_use > 1) {
  stop("nn_prior_zero_weight_scale must be a finite number in [0, 1].")
}
nn_prior_zero_weight_cap_ratio_raw <- params$nn_prior_zero_weight_cap_ratio
if (is.null(nn_prior_zero_weight_cap_ratio_raw) ||
    (length(nn_prior_zero_weight_cap_ratio_raw) == 1L && is.na(nn_prior_zero_weight_cap_ratio_raw))) {
  nn_prior_zero_weight_cap_ratio_use <- NA_real_
} else {
  nn_prior_zero_weight_cap_ratio_use <- suppressWarnings(as.numeric(nn_prior_zero_weight_cap_ratio_raw))
  if (!is.finite(nn_prior_zero_weight_cap_ratio_use) || nn_prior_zero_weight_cap_ratio_use < 0) {
    stop("nn_prior_zero_weight_cap_ratio must be NULL/NA or a finite non-negative number.")
  }
}
nn_prior_zero_birth_fallback_weight_use <- suppressWarnings(as.numeric(params$nn_prior_zero_birth_fallback_weight))
if (!is.finite(nn_prior_zero_birth_fallback_weight_use) ||
    nn_prior_zero_birth_fallback_weight_use < 0 ||
    nn_prior_zero_birth_fallback_weight_use > 1) {
  stop("nn_prior_zero_birth_fallback_weight must be a finite number in [0, 1].")
}
force_refit_use <- isTRUE(params$force_refit)
rebuild_inputs_use <- isTRUE(params$rebuild_inputs)
run_benchmark_use <- isTRUE(params$run_benchmark)
include_posterior_use <- isTRUE(params$include_posterior_comparison)
render_figures_use <- isTRUE(params$render_figures)
top_shift_n_use <- max(1L, as.integer(params$top_shift_n))
focus_pid_use <- trimws(as.character(params$focus_pid))
if (!nzchar(focus_pid_use)) {
  focus_pid_use <- "P2"
}
focus_minobs_use <- suppressWarnings(as.integer(params$focus_minobs))
if (!is.finite(focus_minobs_use) || !focus_minobs_use %in% minobs_values_use) {
  focus_minobs_use <- max(minobs_values_use)
}
focus_pm_use <- suppressWarnings(as.numeric(params$focus_pm))
if (!is.finite(focus_pm_use)) {
  focus_pm_use <- pm_values_use[1]
} else {
  focus_pm_use <- pm_values_use[which.min(abs(pm_values_use - focus_pm_use))][1]
}

diploid_state <- paste(rep(2, 22), collapse = ".")
stage_levels <- c("Primary", "Recurrent")
beneficial_move_levels <- as.vector(rbind(paste0(seq_len(22), "+"), paste0(seq_len(22), "-")))
parameter_levels_use <- parameter_labels_use

set.seed(benchmark_seed_use)
options(mc.cores = n_cores_use)
theme_set(theme_bw(base_size = 12))
```

## Helper Functions

```{r helpers}
sort_pid_levels <- function(x) {
  x <- unique(as.character(x))
  ord_num <- suppressWarnings(as.integer(sub("^P", "", x)))
  x[order(ifelse(is.na(ord_num), Inf, ord_num), x)]
}

pm_to_label <- function(pm) {
  format(pm, scientific = FALSE, trim = TRUE)
}

write_tsv_base <- function(x, path) {
  utils::write.table(
    x,
    file = path,
    sep = "\t",
    quote = FALSE,
    row.names = FALSE,
    col.names = TRUE
  )
  invisible(path)
}

save_table_bundle <- function(x, stem) {
  saveRDS(x, paste0(stem, ".rds"))
  if (is.data.frame(x)) {
    write_tsv_base(x, paste0(stem, ".tsv"))
  }
  invisible(stem)
}

load_saved_table <- function(stem) {
  rds_path <- paste0(stem, ".rds")
  if (file.exists(rds_path)) {
    return(readRDS(rds_path))
  }
  NULL
}

render_tbl <- function(x, caption = NULL, digits = 4) {
  if (knitr::is_html_output()) {
    return(
      knitr::kable(
        x,
        format = "html",
        digits = digits,
        caption = caption,
        table.attr = 'class="three-line-table"'
      )
    )
  }

  knitr::kable(
    x,
    format = "latex",
    digits = digits,
    caption = caption,
    booktabs = TRUE,
    longtable = TRUE
  )
}

html_escape_attr <- function(x) {
  x <- as.character(x)
  x <- gsub("&", "&amp;", x, fixed = TRUE)
  x <- gsub("\"", "&quot;", x, fixed = TRUE)
  x <- gsub("<", "&lt;", x, fixed = TRUE)
  gsub(">", "&gt;", x, fixed = TRUE)
}

emit_report_image <- function(path, alt = "") {
  if (is.na(path) || !nzchar(path) || !file.exists(path)) {
    return(invisible(FALSE))
  }

  if (knitr::is_html_output()) {
    cat(
      sprintf(
        "<div class=\"report-image\"><img src=\"%s\" alt=\"%s\" style=\"max-width:100%%; height:auto; display:block;\" /></div>\n\n",
        knitr::image_uri(path),
        html_escape_attr(alt)
      )
    )
  } else {
    cat(sprintf("![](%s)\n\n", path))
  }

  invisible(TRUE)
}

empty_note_tbl <- function(note) {
  tibble(note = note)
}

alfak_log <- function(...) {
  msg <- paste0(...)
  timestamp <- format(Sys.time(), "%Y-%m-%d %H:%M:%S")
  message("[", timestamp, "] ", msg)
  flush.console()
}

has_complete_alfak_outputs <- function(outdir) {
  required_paths <- file.path(
    outdir,
    c(
      "landscape.Rds",
      "bootstrap_res.Rds",
      "landscape_posterior_samples.Rds",
      "xval.Rds"
    )
  )
  all(file.exists(required_paths))
}

same_optional_numeric <- function(lhs, rhs, tol = 1e-12) {
  lhs_missing <- is.null(lhs) || !length(lhs) || (length(lhs) == 1L && is.na(lhs))
  rhs_missing <- is.null(rhs) || !length(rhs) || (length(rhs) == 1L && is.na(rhs))
  if (lhs_missing && rhs_missing) {
    return(TRUE)
  }
  if (lhs_missing || rhs_missing) {
    return(FALSE)
  }
  lhs_num <- suppressWarnings(as.numeric(lhs[[1]]))
  rhs_num <- suppressWarnings(as.numeric(rhs[[1]]))
  is.finite(lhs_num) && is.finite(rhs_num) && abs(lhs_num - rhs_num) <= tol
}

weighted_prior_cache_matches <- function(cached,
                                         nn_prior,
                                         nn_prior_fit_subset,
                                         nn_prior_zero_exposure_quantile,
                                         nn_prior_zero_weight_scale,
                                         nn_prior_zero_weight_cap_ratio,
                                         nn_prior_zero_birth_fallback_weight) {
  if (!identical(nn_prior, "empirical_censored_weighted")) {
    return(TRUE)
  }

  identical(as.character(cached$nn_prior_fit_subset), nn_prior_fit_subset) &&
    same_optional_numeric(cached$nn_prior_zero_exposure_quantile, nn_prior_zero_exposure_quantile) &&
    same_optional_numeric(cached$nn_prior_zero_weight_scale, nn_prior_zero_weight_scale) &&
    same_optional_numeric(cached$nn_prior_zero_weight_cap_ratio, nn_prior_zero_weight_cap_ratio) &&
    same_optional_numeric(cached$nn_prior_zero_birth_fallback_weight, nn_prior_zero_birth_fallback_weight)
}

beneficial_cnv_profile <- function(landscape_df) {
  fq_flag <- landscape_df$fq
  fq_keep <- if (is.logical(fq_flag)) {
    fq_flag
  } else {
    as.character(fq_flag) %in% c("TRUE", "T", "true", "1")
  }

  k_vec <- as.character(landscape_df$k)
  prop <- rep(NA_real_, length(beneficial_move_levels))
  valid_n <- integer(length(beneficial_move_levels))
  beneficial_n <- integer(length(beneficial_move_levels))
  names(prop) <- beneficial_move_levels
  names(valid_n) <- beneficial_move_levels
  names(beneficial_n) <- beneficial_move_levels

  if (!length(k_vec)) {
    return(list(
      proportion = prop,
      beneficial_n = beneficial_n,
      valid_n = valid_n
    ))
  }

  k_parts <- strsplit(k_vec, ".", fixed = TRUE)
  part_lengths <- lengths(k_parts)
  if (any(part_lengths != 22L)) {
    bad_idx <- which(part_lengths != 22L)[1]
    stop("Found karyotype with != 22 entries in beneficial_cnv_profile(): ", k_vec[bad_idx])
  }

  k_mat <- matrix(as.integer(unlist(k_parts, use.names = FALSE)), ncol = 22, byrow = TRUE)
  fitness_vec <- as.numeric(landscape_df$mean)
  starts <- which(fq_keep %in% TRUE)
  if (!length(starts)) {
    return(list(
      proportion = prop,
      beneficial_n = beneficial_n,
      valid_n = valid_n
    ))
  }

  edge_rows <- vector("list", length(starts))
  edge_idx <- 0L
  for (from in starts) {
    v <- k_mat[from, ]
    neigh_mat <- matrix(rep(v, each = length(beneficial_move_levels)), nrow = length(beneficial_move_levels), ncol = 22)
    for (chr in seq_len(22)) {
      plus_row <- 2L * chr - 1L
      minus_row <- 2L * chr
      neigh_mat[plus_row, chr] <- neigh_mat[plus_row, chr] + 1L
      neigh_mat[minus_row, chr] <- neigh_mat[minus_row, chr] - 1L
    }

    neigh_str <- do.call(paste, c(as.data.frame(neigh_mat), sep = "."))
    to <- match(neigh_str, k_vec)
    valid_cn <- apply(neigh_mat, 1, function(row) all(row >= 1L))
    valid <- valid_cn & !is.na(to)
    if (!any(valid)) {
      next
    }

    to_valid <- to[valid]
    move_valid <- beneficial_move_levels[valid]
    delta <- fitness_vec[to_valid] - fitness_vec[from]
    ok <- is.finite(delta)
    benefit <- rep(NA, length(delta))
    benefit[ok] <- delta[ok] > 0

    edge_idx <- edge_idx + 1L
    edge_rows[[edge_idx]] <- data.frame(
      move = move_valid,
      ok = ok,
      benefit = benefit,
      stringsAsFactors = FALSE
    )
  }

  if (edge_idx > 0L) {
    edges_df <- dplyr::bind_rows(edge_rows[seq_len(edge_idx)])
    for (mv in beneficial_move_levels) {
      ii <- which(edges_df$move == mv)
      if (!length(ii)) {
        next
      }
      sub <- edges_df[ii, , drop = FALSE]
      ok <- sub$ok %in% TRUE
      valid_n[mv] <- sum(ok)
      if (!valid_n[mv]) {
        next
      }
      beneficial_n[mv] <- sum(sub$benefit[ok] %in% TRUE, na.rm = TRUE)
      prop[mv] <- beneficial_n[mv] / valid_n[mv]
    }
  }

  list(
    proportion = prop,
    beneficial_n = beneficial_n,
    valid_n = valid_n
  )
}

landscape_to_patient_df <- function(landscape_df, patient_id) {
  k_mat <- do.call(rbind, lapply(landscape_df$k, function(xx) {
    as.numeric(unlist(strsplit(xx, split = "\\.")))
  }))
  colnames(k_mat) <- as.character(seq_len(ncol(k_mat)))

  tibble::as_tibble(k_mat) %>%
    mutate(
      karyotypes = as.character(landscape_df$k),
      fitness = as.numeric(landscape_df$mean),
      Patient = patient_id
    )
}

ensure_center_in_limits <- function(limits, center = 1) {
  c(min(limits[1], center), max(limits[2], center))
}

make_focus_umap_plot <- function(landscape_df, patient_id, parameter_label, scale_mode = c("relative", "absolute")) {
  scale_mode <- match.arg(scale_mode)
  k_df <- landscape_to_patient_df(landscape_df, patient_id)

  set.seed(benchmark_seed_use)
  umap_mat <- uwot::umap2(as.matrix(k_df[, as.character(seq_len(22))]), n_components = 2, min_dist = 0.9)
  umap_df <- as.data.frame(umap_mat)
  colnames(umap_df) <- c("UMAP1", "UMAP2")

  if (scale_mode == "relative") {
    dip_idx <- which(k_df$karyotypes == diploid_state)
    if (!length(dip_idx)) {
      stop("No diploid karyotype found for patient ", patient_id, " under ", parameter_label)
    }
    dip_mean <- mean(k_df$fitness[dip_idx], na.rm = TRUE)
    if (!is.finite(dip_mean) || dip_mean == 0) {
      stop("Diploid mean fitness is invalid for patient ", patient_id, " under ", parameter_label)
    }

    umap_df$plot_fitness <- k_df$fitness / dip_mean
    legend_name <- "Relative fitness\n(vs diploid)"
    color_center <- 1
    color_limits <- c(min(umap_df$plot_fitness, na.rm = TRUE), max(umap_df$plot_fitness, na.rm = TRUE))
    color_limits <- ensure_center_in_limits(color_limits, center = color_center)
    color_knots <- c(
      seq(color_limits[1], color_center, length.out = 6),
      seq(color_center, color_limits[2], length.out = 6)[-1]
    )
    color_breaks <- c(color_limits[1], color_center, color_limits[2])
  } else {
    umap_df$plot_fitness <- k_df$fitness
    legend_name <- "Absolute fitness"
    color_limits <- c(min(umap_df$plot_fitness, na.rm = TRUE), max(umap_df$plot_fitness, na.rm = TRUE))
    color_knots <- seq(color_limits[1], color_limits[2], length.out = 11)
    color_breaks <- c(color_limits[1], mean(color_limits), color_limits[2])
  }

  umap_df <- umap_df[order(umap_df$plot_fitness, decreasing = FALSE, na.last = TRUE), , drop = FALSE]
  if (diff(color_limits) == 0) {
    if (scale_mode == "relative") {
      color_limits <- c(color_center - 5e-07, color_center + 5e-07)
      color_knots <- c(
        seq(color_limits[1], color_center, length.out = 6),
        seq(color_center, color_limits[2], length.out = 6)[-1]
      )
      color_breaks <- c(color_limits[1], color_center, color_limits[2])
    } else {
      color_limits[2] <- color_limits[2] + 1e-06
      color_knots <- seq(color_limits[1], color_limits[2], length.out = 11)
      color_breaks <- c(color_limits[1], mean(color_limits), color_limits[2])
    }
  }

  color_values <- scales::rescale(color_knots, from = color_limits)

  ggplot(umap_df, aes(x = UMAP1, y = UMAP2, color = plot_fitness)) +
    geom_jitter(width = 0.1, height = 0.1, alpha = 1, size = 1, shape = 16) +
    scale_color_gradientn(
      colours = rev(RColorBrewer::brewer.pal(11, "RdYlBu")),
      values = color_values,
      limits = color_limits,
      breaks = color_breaks,
      labels = signif(color_breaks, 3),
      oob = scales::squish,
      name = legend_name
    ) +
    labs(
      title = paste0(patient_id, " UMAP: ", parameter_label),
      subtitle = if (scale_mode == "relative") "Relative to diploid fitness" else "Absolute landscape mean fitness",
      x = "UMAP1",
      y = "UMAP2"
    ) +
    theme_minimal(base_size = 11)
}

build_focus_parameter_bundles <- function(results_tbl, focus_pid, focus_minobs, focus_pm, parameter_levels) {
  focus_idx <- results_tbl %>%
    filter(
      status == "ok",
      patient_id == focus_pid,
      minobs == focus_minobs,
      dplyr::near(pm, focus_pm),
      parameter_label %in% parameter_levels
    ) %>%
    distinct(parameter_label, .keep_all = TRUE) %>%
    arrange(match(parameter_label, parameter_levels))

  if (!nrow(focus_idx)) {
    return(list())
  }

  bundle_list <- lapply(focus_idx$outdir, read_fit_bundle)
  names(bundle_list) <- focus_idx$parameter_label
  bundle_list
}

landscape_long_from_bundles <- function(bundle_list) {
  if (!length(bundle_list)) {
    return(tibble())
  }

  bind_rows(lapply(names(bundle_list), function(parameter_label_name) {
    landscape_df <- bundle_list[[parameter_label_name]]$landscape
    if (is.null(landscape_df) || !nrow(landscape_df)) {
      return(tibble())
    }

    tibble(
      parameter_label = parameter_label_name,
      k = as.character(landscape_df$k),
      mean = as.numeric(landscape_df$mean),
      median = as.numeric(landscape_df$median),
      sd = as.numeric(landscape_df$sd),
      fq = as.logical(landscape_df$fq),
      nn = as.logical(landscape_df$nn)
    ) %>%
      mutate(
        state_group = case_when(
          fq %in% TRUE ~ "fq",
          nn %in% TRUE ~ "nn",
          TRUE ~ "other"
        )
      )
  }))
}

summarize_landscape_by_parameter <- function(landscape_long) {
  if (!nrow(landscape_long)) {
    return(tibble())
  }

  landscape_long %>%
    group_by(parameter_label) %>%
    summarise(
      n_karyotypes = n(),
      n_fq = sum(fq %in% TRUE, na.rm = TRUE),
      n_nn = sum(nn %in% TRUE, na.rm = TRUE),
      mean_landscape_mean = mean(mean, na.rm = TRUE),
      median_landscape_mean = median(mean, na.rm = TRUE),
      mean_landscape_sd = mean(sd, na.rm = TRUE),
      min_landscape_mean = min(mean, na.rm = TRUE),
      max_landscape_mean = max(mean, na.rm = TRUE),
      negative_mean_n = sum(mean < 0, na.rm = TRUE),
      negative_median_n = sum(median < 0, na.rm = TRUE),
      .groups = "drop"
    )
}

summarize_focus_landscape_variation <- function(landscape_long, parameter_levels, top_n = 15L) {
  if (!nrow(landscape_long) || !length(parameter_levels)) {
    return(tibble())
  }

  wide_tbl <- landscape_long %>%
    select(parameter_label, k, mean) %>%
    distinct() %>%
    pivot_wider(names_from = parameter_label, values_from = mean)

  for (parameter_name in parameter_levels) {
    if (!parameter_name %in% names(wide_tbl)) {
      wide_tbl[[parameter_name]] <- NA_real_
    }
  }
  wide_tbl <- wide_tbl[, c("k", parameter_levels), drop = FALSE]
  mean_mat <- as.matrix(wide_tbl[, parameter_levels, drop = FALSE])

  state_tbl <- landscape_long %>%
    group_by(k) %>%
    summarise(
      state_group = case_when(
        any(fq %in% TRUE, na.rm = TRUE) ~ "fq",
        any(nn %in% TRUE, na.rm = TRUE) ~ "nn",
        TRUE ~ "other"
      ),
      .groups = "drop"
    )

  out <- wide_tbl %>%
    left_join(state_tbl, by = "k") %>%
    mutate(
      n_parameter_label = rowSums(is.finite(mean_mat)),
      mean_range = apply(mean_mat, 1, function(x) {
        if (sum(is.finite(x)) >= 2L) diff(range(x, na.rm = TRUE)) else NA_real_
      }),
      mean_sd_across = apply(mean_mat, 1, function(x) {
        if (sum(is.finite(x)) >= 2L) stats::sd(x, na.rm = TRUE) else NA_real_
      }),
      min_parameter_label = apply(mean_mat, 1, function(x) {
        if (!sum(is.finite(x))) {
          return(NA_character_)
        }
        parameter_levels[which.min(replace(x, !is.finite(x), Inf))][1]
      }),
      max_parameter_label = apply(mean_mat, 1, function(x) {
        if (!sum(is.finite(x))) {
          return(NA_character_)
        }
        parameter_levels[which.max(replace(x, !is.finite(x), -Inf))][1]
      })
    ) %>%
    select(k, state_group, n_parameter_label, all_of(parameter_levels), mean_range, mean_sd_across, min_parameter_label, max_parameter_label) %>%
    arrange(desc(mean_range), desc(mean_sd_across), k) %>%
    slice_head(n = top_n)

  out
}

build_focus_parity_tbl <- function(bundle_list, parameter_levels, value_col = "mean") {
  if (length(parameter_levels) < 2L) {
    return(tibble())
  }

  pair_defs <- combn(parameter_levels, 2, simplify = FALSE)
  bind_rows(lapply(pair_defs, function(pair) {
    lhs_name <- pair[[1]]
    rhs_name <- pair[[2]]
    lhs_landscape <- bundle_list[[lhs_name]]$landscape
    rhs_landscape <- bundle_list[[rhs_name]]$landscape

    if (is.null(lhs_landscape) || is.null(rhs_landscape)) {
      return(tibble())
    }

    lhs_tbl <- lhs_landscape %>%
      transmute(
        k,
        lhs_value = .data[[value_col]],
        lhs_fq = as.logical(fq),
        lhs_nn = as.logical(nn)
      )
    rhs_tbl <- rhs_landscape %>%
      transmute(
        k,
        rhs_value = .data[[value_col]],
        rhs_fq = as.logical(fq),
        rhs_nn = as.logical(nn)
      )

    inner_join(lhs_tbl, rhs_tbl, by = "k") %>%
      mutate(
        lhs_label = lhs_name,
        rhs_label = rhs_name,
        comparison = paste(lhs_name, rhs_name, sep = " vs "),
        state_group = case_when(
          lhs_fq | rhs_fq ~ "fq",
          lhs_nn | rhs_nn ~ "nn",
          TRUE ~ "other"
        )
      ) %>%
      select(comparison, lhs_label, rhs_label, k, state_group, lhs_value, rhs_value)
  }))
}

beneficial_long_from_profiles <- function(beneficial_profiles, parameter_levels = names(beneficial_profiles)) {
  if (!length(beneficial_profiles)) {
    return(tibble())
  }

  out <- bind_rows(lapply(parameter_levels, function(parameter_label_name) {
    prof <- beneficial_profiles[[parameter_label_name]]
    if (is.null(prof)) {
      return(tibble())
    }

    tibble(
      parameter_label = parameter_label_name,
      move = beneficial_move_levels,
      proportion = as.numeric(prof$proportion[beneficial_move_levels]),
      beneficial_n = as.integer(prof$beneficial_n[beneficial_move_levels]),
      valid_n = as.integer(prof$valid_n[beneficial_move_levels])
    )
  }))

  if (!nrow(out) || !"move" %in% names(out)) {
    return(tibble())
  }

  out %>%
    mutate(
      chromosome = suppressWarnings(as.integer(sub("[+-]$", "", move))),
      direction = ifelse(grepl("\\+$", move), "gain", "loss"),
      move = factor(move, levels = beneficial_move_levels),
      parameter_label = factor(parameter_label, levels = parameter_levels)
    )
}

summarize_beneficial_by_parameter <- function(beneficial_long) {
  if (!nrow(beneficial_long)) {
    return(tibble())
  }

  beneficial_long %>%
    group_by(parameter_label) %>%
    summarise(
      n_move_types = sum(!is.na(proportion)),
      mean_proportion = mean(proportion, na.rm = TRUE),
      median_proportion = median(proportion, na.rm = TRUE),
      total_beneficial_n = sum(beneficial_n, na.rm = TRUE),
      total_valid_n = sum(valid_n, na.rm = TRUE),
      weighted_proportion = ifelse(sum(valid_n, na.rm = TRUE) > 0, sum(beneficial_n, na.rm = TRUE) / sum(valid_n, na.rm = TRUE), NA_real_),
      .groups = "drop"
    )
}

build_focus_beneficial_shift_tbl <- function(beneficial_long, parameter_levels, top_n = 15L) {
  if (!nrow(beneficial_long) || length(parameter_levels) < 2L) {
    return(tibble())
  }

  pair_defs <- combn(parameter_levels, 2, simplify = FALSE)
  bind_rows(lapply(pair_defs, function(pair) {
    lhs_name <- pair[[1]]
    rhs_name <- pair[[2]]
    wide_tbl <- beneficial_long %>%
      filter(parameter_label %in% c(lhs_name, rhs_name)) %>%
      select(parameter_label, move, chromosome, direction, proportion, valid_n) %>%
      pivot_wider(names_from = parameter_label, values_from = c(proportion, valid_n), names_sep = "__")

    lhs_prop_col <- paste0("proportion__", lhs_name)
    rhs_prop_col <- paste0("proportion__", rhs_name)
    lhs_valid_col <- paste0("valid_n__", lhs_name)
    rhs_valid_col <- paste0("valid_n__", rhs_name)
    missing_cols <- setdiff(c(lhs_prop_col, rhs_prop_col, lhs_valid_col, rhs_valid_col), names(wide_tbl))
    for (col in missing_cols) {
      wide_tbl[[col]] <- NA_real_
    }

    wide_tbl %>%
      transmute(
        lhs_label = lhs_name,
        rhs_label = rhs_name,
        move,
        chromosome,
        direction,
        lhs_prop = .data[[lhs_prop_col]],
        rhs_prop = .data[[rhs_prop_col]],
        diff = rhs_prop - lhs_prop,
        lhs_valid_n = .data[[lhs_valid_col]],
        rhs_valid_n = .data[[rhs_valid_col]]
      ) %>%
      arrange(desc(abs(diff)), move) %>%
      slice_head(n = top_n)
  }))
}

available_pid_dirs <- function(base_dir) {
  dd <- list.dirs(base_dir, recursive = FALSE, full.names = TRUE)
  dd <- dd[file.info(dd)$isdir %in% TRUE]
  dd <- dd[grepl("^P[0-9]+$", basename(dd))]
  sort_pid_levels(basename(dd))
}

read_transcriptome_perspective <- function(rds_path) {
  obj <- readRDS(rds_path)
  stopifnot(is.list(obj), "profile" %in% names(obj))
  profile <- as.matrix(obj$profile)
  storage.mode(profile) <- "numeric"
  list(
    passaging_id = obj$passaging_id,
    perspective_type = obj$perspective_type,
    profile = round(profile)
  )
}

compute_delta_time_days <- function(df_pid) {
  df_pid <- df_pid %>%
    mutate(Stage = factor(Stage, levels = stage_levels)) %>%
    arrange(Stage, effective_id)

  age_primary <- suppressWarnings(as.numeric(df_pid$Age[df_pid$Stage == "Primary"][1]))
  age_recurrent <- suppressWarnings(as.numeric(df_pid$Age[df_pid$Stage == "Recurrent"][1]))

  recurrent_days <- if (is.finite(age_primary) && is.finite(age_recurrent)) {
    as.integer(round(abs(age_recurrent - age_primary) * 360))
  } else {
    NA_integer_
  }
  if (!is.finite(recurrent_days) || recurrent_days <= 1L) {
    recurrent_days <- 180L
  }

  df_pid %>%
    mutate(
      Delta_Time = ifelse(as.character(Stage) == "Primary", 1L, recurrent_days)
    )
}

build_patient_manifest <- function(meta_tbl, base_dir, patient_subset = NULL) {
  pid_dirs <- available_pid_dirs(base_dir)
  manifest <- meta_tbl %>%
    mutate(
      effective_id = ifelse(!is.na(ID_alt) & nzchar(ID_alt), ID_alt, ID),
      pid = as.character(pid),
      Stage = as.character(Stage),
      Age = suppressWarnings(as.numeric(Age))
    ) %>%
    filter(pid %in% pid_dirs)

  if (!is.null(patient_subset)) {
    manifest <- manifest %>% filter(pid %in% patient_subset)
  }

  manifest %>%
    group_by(pid) %>%
    group_modify(function(.x, .y) compute_delta_time_days(.x)) %>%
    ungroup() %>%
    mutate(
      rds_path = file.path(base_dir, pid, "perspectives", effective_id, "TranscriptomePerspective.rds"),
      exists = file.exists(rds_path)
    ) %>%
    arrange(factor(pid, levels = sort_pid_levels(pid)), factor(Stage, levels = stage_levels))
}

profile_to_karyotype_strings <- function(profile_mat) {
  if (nrow(profile_mat) != 22) {
    warning("Expected 22 chromosome rows; got ", nrow(profile_mat))
  }
  x <- t(round(as.matrix(profile_mat)))
  apply(x, 1, paste, collapse = ".")
}

build_count_matrix_from_profiles <- function(patient_manifest) {
  stopifnot(length(unique(patient_manifest$pid)) == 1L)
  rows_ok <- patient_manifest[patient_manifest$exists, , drop = FALSE]
  rows_ok <- rows_ok[order(match(rows_ok$Stage, stage_levels)), , drop = FALSE]
  if (nrow(rows_ok) < 2) {
    stop("Need at least two available samples for patient ", unique(patient_manifest$pid))
  }

  long_df <- do.call(rbind, lapply(seq_len(nrow(rows_ok)), function(i) {
    rr <- rows_ok[i, , drop = FALSE]
    obj <- read_transcriptome_perspective(rr$rds_path)
    if (!is.null(obj$passaging_id) &&
        !identical(as.character(obj$passaging_id), as.character(rr$effective_id))) {
      warning("passaging_id mismatch for ", rr$pid, " / ", rr$effective_id)
    }
    tibble(
      karyotype = profile_to_karyotype_strings(obj$profile),
      Delta_Time = rr$Delta_Time,
      Stage = rr$Stage,
      effective_id = rr$effective_id
    )
  }))

  delta_levels <- sort(unique(rows_ok$Delta_Time))
  count_mat <- as.data.frame.matrix(table(
    long_df$karyotype,
    factor(long_df$Delta_Time, levels = delta_levels)
  ))
  colnames(count_mat) <- as.character(delta_levels)
  count_mat <- count_mat[order(rowSums(count_mat), decreasing = TRUE), , drop = FALSE]

  list(
    count_matrix = count_mat,
    long_df = long_df
  )
}

save_alfak_input <- function(count_matrix, out_path) {
  yi <- list(
    x = count_matrix,
    pop.fitness = NULL,
    dt = 1
  )
  saveRDS(yi, out_path)
  invisible(out_path)
}

build_benchmark_inputs <- function(meta_tbl, base_dir, input_dir, rebuild_inputs = FALSE, patient_subset = NULL) {
  manifest_tbl <- build_patient_manifest(meta_tbl, base_dir, patient_subset = patient_subset)
  save_table_bundle(manifest_tbl, file.path(tables_dir, "benchmark_patient_manifest"))

  patient_ids <- sort_pid_levels(unique(manifest_tbl$pid))
  if (!length(patient_ids)) {
    stop("No benchmark patients discovered under ", base_dir)
  }

  rows <- lapply(patient_ids, function(pid) {
    patient_manifest <- manifest_tbl %>% filter(pid == !!pid)
    input_rds <- file.path(input_dir, paste0(pid, ".Rds"))
    count_matrix_path <- file.path(input_dir, paste0(pid, "_count_matrix.tsv"))
    long_df_path <- file.path(input_dir, paste0(pid, "_profile_long.tsv"))

    if (rebuild_inputs || !file.exists(input_rds) || !file.exists(count_matrix_path) || !file.exists(long_df_path)) {
      built <- build_count_matrix_from_profiles(patient_manifest)
      count_export <- data.frame(
        karyotype = rownames(built$count_matrix),
        built$count_matrix,
        check.names = FALSE
      )
      save_alfak_input(built$count_matrix, input_rds)
      write_tsv_base(count_export, count_matrix_path)
      write_tsv_base(built$long_df, long_df_path)
    }

    yi <- readRDS(input_rds)
    yi$x <- as.data.frame(yi$x)
    yi_minobs5 <- yi$x
    if (diploid_state %in% rownames(yi_minobs5)) {
      yi_minobs5 <- yi_minobs5[rownames(yi_minobs5) != diploid_state, , drop = FALSE]
    }
    input_row_count_minobs5 <- if (nrow(yi_minobs5)) {
      sum(rowSums(yi_minobs5, na.rm = TRUE) >= 5L)
    } else {
      0L
    }

    tibble(
      patient_id = pid,
      input_rds = input_rds,
      count_matrix_path = count_matrix_path,
      long_df_path = long_df_path,
      input_row_count_minobs5 = input_row_count_minobs5,
      input_row_count = nrow(yi$x),
      n_timepoints = ncol(yi$x),
      n_karyotypes = nrow(yi$x),
      total_cells = sum(as.matrix(yi$x), na.rm = TRUE),
      delta_time_labels = paste(colnames(yi$x), collapse = ","),
      effective_ids = paste(patient_manifest$effective_id[patient_manifest$exists], collapse = ","),
      stage_sequence = paste(as.character(patient_manifest$Stage[patient_manifest$exists]), collapse = " -> ")
    )
  })

  bind_rows(rows) %>%
    arrange(factor(patient_id, levels = sort_pid_levels(patient_id)))
}

extract_xval_metrics <- function(xv) {
  metrics <- list(
    xval_r2 = NA_real_,
    xval_cor = NA_real_,
    xval_rmse = NA_real_,
    xval_mae = NA_real_,
    n_xval = NA_integer_
  )

  if (is.null(xv)) {
    return(metrics)
  }

  if (is.atomic(xv) && is.numeric(xv) && is.null(dim(xv)) && length(xv) >= 1L) {
    metrics$xval_r2 <- suppressWarnings(as.numeric(xv[[1]]))
    return(metrics)
  }

  if (!is.list(xv)) {
    return(metrics)
  }

  if (!is.null(xv$R2R)) {
    metrics$xval_r2 <- suppressWarnings(as.numeric(xv$R2R))
  }

  if (is.null(xv$tmp)) {
    return(metrics)
  }

  tmp_df <- as.data.frame(xv$tmp)
  if (ncol(tmp_df) < 2) {
    return(metrics)
  }

  colnames(tmp_df)[1:2] <- c("f_est", "f_xv")
  tmp_df <- tmp_df[, c("f_est", "f_xv"), drop = FALSE]
  tmp_df$f_est <- suppressWarnings(as.numeric(tmp_df$f_est))
  tmp_df$f_xv <- suppressWarnings(as.numeric(tmp_df$f_xv))
  tmp_df <- tmp_df[is.finite(tmp_df$f_est) & is.finite(tmp_df$f_xv), , drop = FALSE]

  metrics$n_xval <- nrow(tmp_df)
  if (!nrow(tmp_df)) {
    return(metrics)
  }

  diff_vec <- tmp_df$f_est - tmp_df$f_xv
  metrics$xval_rmse <- sqrt(mean(diff_vec^2))
  metrics$xval_mae <- mean(abs(diff_vec))
  if (nrow(tmp_df) >= 2) {
    metrics$xval_cor <- suppressWarnings(stats::cor(tmp_df$f_est, tmp_df$f_xv, use = "complete.obs"))
  }
  metrics
}

refresh_cached_fit_row <- function(cached,
                                   outdir,
                                   patient_id,
                                   pm,
                                   minobs,
                                   benchmark_seed,
                                   parameter_label,
                                   nn_prior,
                                   nn_prior_grid_n,
                                   nn_prior_fit_subset,
                                   nn_prior_zero_exposure_quantile,
                                   nn_prior_zero_weight_scale,
                                   nn_prior_zero_weight_cap_ratio,
                                   nn_prior_zero_birth_fallback_weight,
                                   warning_log_path,
                                   landscape_path,
                                   bootstrap_path,
                                   posterior_path,
                                   xval_path) {
  if (!is.list(cached)) {
    cached <- list()
  }

  warning_lines <- if (file.exists(warning_log_path)) readLines(warning_log_path, warn = FALSE) else character()
  cached$patient_id <- patient_id
  cached$outdir <- outdir
  cached$pm <- pm
  cached$pm_label <- pm_to_label(pm)
  cached$minobs <- minobs
  cached$benchmark_seed <- benchmark_seed
  cached$parameter_label <- parameter_label
  cached$nn_prior <- nn_prior
  cached$nn_prior_grid_n <- nn_prior_grid_n
  cached$nn_prior_fit_subset <- nn_prior_fit_subset
  cached$nn_prior_zero_exposure_quantile <- nn_prior_zero_exposure_quantile
  cached$nn_prior_zero_weight_scale <- nn_prior_zero_weight_scale
  cached$nn_prior_zero_weight_cap_ratio <- nn_prior_zero_weight_cap_ratio
  cached$nn_prior_zero_birth_fallback_weight <- nn_prior_zero_birth_fallback_weight
  cached$cached <- TRUE
  cached$warning_count <- length(warning_lines)
  cached$lambda_endpoint_warning_count <- sum(grepl("^Grid searches over lambda", warning_lines))
  cached$warning_messages <- if (length(warning_lines)) paste(warning_lines, collapse = " || ") else NA_character_
  cached$landscape_path <- if (file.exists(landscape_path)) landscape_path else NA_character_
  cached$bootstrap_path <- if (file.exists(bootstrap_path)) bootstrap_path else NA_character_
  cached$posterior_path <- if (file.exists(posterior_path)) posterior_path else NA_character_
  cached$xval_path <- if (file.exists(xval_path)) xval_path else NA_character_
  cached
}

fit_path_component_label <- function(value, key) {
  if (is.null(value) || (length(value) == 1L && is.na(value))) {
    return(NA_character_)
  }

  switch(
    key,
    parameter_label = as.character(value),
    pm = paste0("pm_", pm_to_label(as.numeric(value))),
    minobs = paste0("MINOBS_", as.integer(value)),
    patient_id = as.character(value),
    stop("Unsupported fit path component key: ", key)
  )
}

build_fit_outdir <- function(root_dir, patient_id, path_components) {
  component_labels <- purrr::imap_chr(path_components, fit_path_component_label)
  component_labels <- component_labels[is.finite(nchar(component_labels))]
  do.call(file.path, c(list(root_dir), as.list(component_labels), list(patient_id)))
}

infer_parameter_label <- function(nn_prior) {
  paste0("nn_prior_", nn_prior)
}

build_parameter_spec_tbl <- function(parameter_labels) {
  tibble(
    parameter_label = parameter_labels,
    nn_prior = sub("^nn_prior_", "", parameter_labels)
  ) %>%
    mutate(parameter_label = factor(parameter_label, levels = parameter_labels)) %>%
    arrange(parameter_label) %>%
    mutate(parameter_label = as.character(parameter_label))
}

reconcile_fit_results_tbl <- function(fit_results_tbl, task_tbl = NULL) {
  if (is.null(fit_results_tbl) || !nrow(fit_results_tbl)) {
    return(tibble())
  }

  fit_results_tbl <- as_tibble(fit_results_tbl)
  if (!"parameter_label" %in% names(fit_results_tbl)) {
    if ("nn_prior" %in% names(fit_results_tbl)) {
      fit_results_tbl$parameter_label <- infer_parameter_label(fit_results_tbl$nn_prior)
    } else {
      fit_results_tbl$parameter_label <- NA_character_
    }
  }

  if (!is.null(task_tbl) && nrow(task_tbl)) {
    task_idx_tbl <- task_tbl %>%
      transmute(
        patient_id,
        minobs,
        pm,
        parameter_label,
        nn_prior,
        expected_outdir = outdir
      )

    fit_results_tbl <- fit_results_tbl %>%
      rename(recorded_outdir = outdir) %>%
      left_join(task_idx_tbl, by = c("patient_id", "minobs", "pm", "parameter_label", "nn_prior")) %>%
      mutate(outdir = coalesce(expected_outdir, recorded_outdir)) %>%
      select(-recorded_outdir, -expected_outdir)
  }

  fit_results_tbl %>%
    mutate(
      landscape_path = ifelse(!is.na(outdir) & nzchar(outdir), file.path(outdir, "landscape.Rds"), NA_character_),
      bootstrap_path = ifelse(!is.na(outdir) & nzchar(outdir), file.path(outdir, "bootstrap_res.Rds"), NA_character_),
      posterior_path = ifelse(!is.na(outdir) & nzchar(outdir), file.path(outdir, "landscape_posterior_samples.Rds"), NA_character_),
      xval_path = ifelse(!is.na(outdir) & nzchar(outdir), file.path(outdir, "xval.Rds"), NA_character_)
    ) %>%
    select(-any_of("fit_mode")) %>%
    relocate(parameter_label, .after = benchmark_seed)
}

run_alfak_fit <- function(patient_id,
                          input_rds,
                          outdir,
                          minobs,
                          pm,
                          nboot,
                          n0,
                          nb,
                          benchmark_seed,
                          parameter_label,
                          correct_efflux = TRUE,
                          nn_prior = "none",
                          nn_prior_grid_n = 81L,
                          nn_prior_fit_subset = "hybrid",
                          nn_prior_zero_exposure_quantile = 0.10,
                          nn_prior_zero_weight_scale = 0.50,
                          nn_prior_zero_weight_cap_ratio = NA_real_,
                          nn_prior_zero_birth_fallback_weight = 0.50,
                          force_refit = FALSE) {
  landscape_path <- file.path(outdir, "landscape.Rds")
  bootstrap_path <- file.path(outdir, "bootstrap_res.Rds")
  posterior_path <- file.path(outdir, "landscape_posterior_samples.Rds")
  xval_path <- file.path(outdir, "xval.Rds")
  summary_path <- file.path(outdir, "fit_task_result.rds")
  warning_log_path <- file.path(outdir, "fit_warnings.log")
  pm_label <- pm_to_label(pm)
  task_tag <- paste0(
    patient_id,
    " | minobs=", minobs,
    " | pm=", pm_label,
    " | parameter_label=", parameter_label,
    " | nn_prior=", nn_prior,
    " | grid=", nn_prior_grid_n,
    if (identical(nn_prior, "empirical_censored_weighted")) paste0(
      " | fit_subset=", nn_prior_fit_subset,
      " | zero_q=", signif(nn_prior_zero_exposure_quantile, 4),
      " | zero_scale=", signif(nn_prior_zero_weight_scale, 4),
      " | zero_cap=", if (is.na(nn_prior_zero_weight_cap_ratio)) "adaptive" else signif(nn_prior_zero_weight_cap_ratio, 4),
      " | zero_birth_fallback=", signif(nn_prior_zero_birth_fallback_weight, 4)
    ) else ""
  )

  if (!force_refit && file.exists(summary_path)) {
    cached <- tryCatch(readRDS(summary_path), error = function(e) NULL)
    if (is.list(cached) &&
        identical(cached$status, "ok") &&
        has_complete_alfak_outputs(outdir) &&
        weighted_prior_cache_matches(
          cached = cached,
          nn_prior = nn_prior,
          nn_prior_fit_subset = nn_prior_fit_subset,
          nn_prior_zero_exposure_quantile = nn_prior_zero_exposure_quantile,
          nn_prior_zero_weight_scale = nn_prior_zero_weight_scale,
          nn_prior_zero_weight_cap_ratio = nn_prior_zero_weight_cap_ratio,
          nn_prior_zero_birth_fallback_weight = nn_prior_zero_birth_fallback_weight
        )) {
      cached <- refresh_cached_fit_row(
        cached = cached,
        outdir = outdir,
        patient_id = patient_id,
        pm = pm,
        minobs = minobs,
        benchmark_seed = benchmark_seed,
        parameter_label = parameter_label,
        nn_prior = nn_prior,
        nn_prior_grid_n = nn_prior_grid_n,
        nn_prior_fit_subset = nn_prior_fit_subset,
        nn_prior_zero_exposure_quantile = nn_prior_zero_exposure_quantile,
        nn_prior_zero_weight_scale = nn_prior_zero_weight_scale,
        nn_prior_zero_weight_cap_ratio = nn_prior_zero_weight_cap_ratio,
        nn_prior_zero_birth_fallback_weight = nn_prior_zero_birth_fallback_weight,
        warning_log_path = warning_log_path,
        landscape_path = landscape_path,
        bootstrap_path = bootstrap_path,
        posterior_path = posterior_path,
        xval_path = xval_path
      )
      saveRDS(cached, summary_path)
      alfak_log("ALFA-K cached: ", task_tag)
      return(as_tibble(cached))
    }
  }

  if (!force_refit &&
      has_complete_alfak_outputs(outdir) &&
      !identical(nn_prior, "empirical_censored_weighted")) {
    warning_lines <- if (file.exists(warning_log_path)) readLines(warning_log_path, warn = FALSE) else character()
    xv <- tryCatch(readRDS(xval_path), error = function(e) NULL)
    cached <- c(
      list(
        patient_id = patient_id,
        outdir = outdir,
        pm = pm,
        pm_label = pm_label,
        minobs = minobs,
        benchmark_seed = benchmark_seed,
        parameter_label = parameter_label,
        nn_prior = nn_prior,
        nn_prior_grid_n = nn_prior_grid_n,
        nn_prior_fit_subset = nn_prior_fit_subset,
        nn_prior_zero_exposure_quantile = nn_prior_zero_exposure_quantile,
        nn_prior_zero_weight_scale = nn_prior_zero_weight_scale,
        nn_prior_zero_weight_cap_ratio = nn_prior_zero_weight_cap_ratio,
        nn_prior_zero_birth_fallback_weight = nn_prior_zero_birth_fallback_weight,
        status = "ok",
        cached = TRUE,
        error_message = NA_character_,
        elapsed_sec = NA_real_,
        warning_count = length(warning_lines),
        lambda_endpoint_warning_count = sum(grepl("^Grid searches over lambda", warning_lines)),
        warning_messages = if (length(warning_lines)) paste(warning_lines, collapse = " || ") else NA_character_,
        landscape_path = landscape_path,
        bootstrap_path = if (file.exists(bootstrap_path)) bootstrap_path else NA_character_,
        posterior_path = if (file.exists(posterior_path)) posterior_path else NA_character_,
        xval_path = xval_path
      ),
      extract_xval_metrics(xv)
    )
    saveRDS(cached, summary_path)
    alfak_log("ALFA-K cached: ", task_tag)
    return(as_tibble(cached))
  }

  alfak_log("ALFA-K start: ", task_tag)
  dir.create(outdir, recursive = TRUE, showWarnings = FALSE)

  yi <- readRDS(input_rds)
  yi$x <- as.data.frame(yi$x)
  if (diploid_state %in% rownames(yi$x)) {
    yi$x <- yi$x[rownames(yi$x) != diploid_state, , drop = FALSE]
  }
  if (!nrow(yi$x)) {
    stop("All rows were filtered out for input: ", input_rds)
  }
  if (max(rowSums(yi$x), na.rm = TRUE) < minobs) {
    stop("No frequent karyotypes reach minobs=", minobs, " for input: ", input_rds)
  }

  warning_messages <- character()
  started_at <- Sys.time()

  fit_row <- tryCatch({
    set.seed(benchmark_seed)
    withCallingHandlers({
      alfakR::alfak(
        yi = yi,
        outdir = outdir,
        passage_times = NULL,
        minobs = minobs,
        nboot = nboot,
        n0 = n0,
        nb = nb,
        pm = pm,
        correct_efflux = correct_efflux,
        nn_prior = nn_prior,
        nn_prior_grid_n = nn_prior_grid_n,
        nn_prior_fit_subset = nn_prior_fit_subset,
        nn_prior_zero_exposure_quantile = nn_prior_zero_exposure_quantile,
        nn_prior_zero_weight_scale = nn_prior_zero_weight_scale,
        nn_prior_zero_weight_cap_ratio = if (is.na(nn_prior_zero_weight_cap_ratio)) NULL else nn_prior_zero_weight_cap_ratio,
        nn_prior_zero_birth_fallback_weight = nn_prior_zero_birth_fallback_weight
      )
    }, warning = function(w) {
      warning_messages <<- c(warning_messages, conditionMessage(w))
      invokeRestart("muffleWarning")
    })

    xv <- readRDS(xval_path)
    elapsed_sec <- as.numeric(difftime(Sys.time(), started_at, units = "secs"))
    warning_messages <- unique(warning_messages)
    if (length(warning_messages)) {
      writeLines(warning_messages, warning_log_path)
    }

    c(
      list(
        patient_id = patient_id,
        outdir = outdir,
        pm = pm,
        pm_label = pm_label,
        minobs = minobs,
        benchmark_seed = benchmark_seed,
        parameter_label = parameter_label,
        nn_prior = nn_prior,
        nn_prior_grid_n = nn_prior_grid_n,
        nn_prior_fit_subset = nn_prior_fit_subset,
        nn_prior_zero_exposure_quantile = nn_prior_zero_exposure_quantile,
        nn_prior_zero_weight_scale = nn_prior_zero_weight_scale,
        nn_prior_zero_weight_cap_ratio = nn_prior_zero_weight_cap_ratio,
        nn_prior_zero_birth_fallback_weight = nn_prior_zero_birth_fallback_weight,
        status = "ok",
        cached = FALSE,
        error_message = NA_character_,
        elapsed_sec = elapsed_sec,
        warning_count = length(warning_messages),
        lambda_endpoint_warning_count = sum(grepl("^Grid searches over lambda", warning_messages)),
        warning_messages = if (length(warning_messages)) paste(warning_messages, collapse = " || ") else NA_character_,
        landscape_path = landscape_path,
        bootstrap_path = if (file.exists(bootstrap_path)) bootstrap_path else NA_character_,
        posterior_path = if (file.exists(posterior_path)) posterior_path else NA_character_,
        xval_path = xval_path
      ),
      extract_xval_metrics(xv)
    )
  }, error = function(e) {
    elapsed_sec <- as.numeric(difftime(Sys.time(), started_at, units = "secs"))
    warning_messages <- unique(warning_messages)
    if (length(warning_messages)) {
      writeLines(warning_messages, warning_log_path)
    }

    alfak_log("ALFA-K error: ", task_tag, " | ", conditionMessage(e))
    list(
      patient_id = patient_id,
      outdir = outdir,
      pm = pm,
      pm_label = pm_label,
      minobs = minobs,
      benchmark_seed = benchmark_seed,
      parameter_label = parameter_label,
      nn_prior = nn_prior,
      nn_prior_grid_n = nn_prior_grid_n,
      nn_prior_fit_subset = nn_prior_fit_subset,
      nn_prior_zero_exposure_quantile = nn_prior_zero_exposure_quantile,
      nn_prior_zero_weight_scale = nn_prior_zero_weight_scale,
      nn_prior_zero_weight_cap_ratio = nn_prior_zero_weight_cap_ratio,
      nn_prior_zero_birth_fallback_weight = nn_prior_zero_birth_fallback_weight,
      status = "error",
      cached = FALSE,
      error_message = conditionMessage(e),
      elapsed_sec = elapsed_sec,
      warning_count = length(warning_messages),
      lambda_endpoint_warning_count = sum(grepl("^Grid searches over lambda", warning_messages)),
      warning_messages = if (length(warning_messages)) paste(warning_messages, collapse = " || ") else NA_character_,
      xval_r2 = NA_real_,
      xval_cor = NA_real_,
      xval_rmse = NA_real_,
      xval_mae = NA_real_,
      n_xval = NA_integer_,
      landscape_path = if (file.exists(landscape_path)) landscape_path else NA_character_,
      bootstrap_path = if (file.exists(bootstrap_path)) bootstrap_path else NA_character_,
      posterior_path = if (file.exists(posterior_path)) posterior_path else NA_character_,
      xval_path = if (file.exists(xval_path)) xval_path else NA_character_
    )
  })

  saveRDS(fit_row, summary_path)
  if (identical(fit_row$status, "ok")) {
    alfak_log("ALFA-K done: ", task_tag, " | xval_r2=", signif(fit_row$xval_r2, 4))
  }
  as_tibble(fit_row)
}

task_outdir_parameter <- function(root_dir, patient_id, minobs, pm, parameter_label) {
  build_fit_outdir(
    root_dir = root_dir,
    patient_id = patient_id,
    path_components = list(
      parameter_label = parameter_label,
      pm = pm,
      minobs = minobs
    )
  )
}

build_parameter_tasks <- function(input_index_tbl,
                                  fit_root,
                                  minobs_values,
                                  pm_values,
                                  parameter_spec_tbl,
                                  nn_prior_grid_n,
                                  nn_prior_fit_subset,
                                  nn_prior_zero_exposure_quantile,
                                  nn_prior_zero_weight_scale,
                                  nn_prior_zero_weight_cap_ratio,
                                  nn_prior_zero_birth_fallback_weight,
                                  nboot,
                                  n0,
                                  nb,
                                  benchmark_seed,
                                  correct_efflux,
                                  force_refit) {
  if (is.null(parameter_spec_tbl) || !nrow(parameter_spec_tbl)) {
    return(tibble())
  }

  tidyr::crossing(
    input_index_tbl %>% select(patient_id, input_rds, input_row_count_minobs5, input_row_count),
    minobs = minobs_values,
    pm = pm_values
  ) %>%
    tidyr::crossing(parameter_spec_tbl) %>%
    mutate(
      nn_prior_grid_n = nn_prior_grid_n,
      nn_prior_fit_subset = nn_prior_fit_subset,
      nn_prior_zero_exposure_quantile = nn_prior_zero_exposure_quantile,
      nn_prior_zero_weight_scale = nn_prior_zero_weight_scale,
      nn_prior_zero_weight_cap_ratio = nn_prior_zero_weight_cap_ratio,
      nn_prior_zero_birth_fallback_weight = nn_prior_zero_birth_fallback_weight,
      outdir = purrr::pmap_chr(
        list(patient_id, minobs, pm, parameter_label),
        ~ task_outdir_parameter(fit_root, ..1, ..2, ..3, ..4)
      ),
      nboot = nboot,
      n0 = n0,
      nb = nb,
      benchmark_seed = benchmark_seed,
      correct_efflux = correct_efflux,
      force_refit = force_refit
    ) %>%
    arrange(
      desc(input_row_count_minobs5),
      desc(input_row_count),
      factor(patient_id, levels = sort_pid_levels(patient_id)),
      minobs,
      pm,
      factor(parameter_label, levels = parameter_spec_tbl$parameter_label)
    )
}

run_task_table_parallel <- function(task_tbl) {
  if (!nrow(task_tbl)) {
    return(tibble())
  }

  if (.Platform$OS.type == "unix" && n_cores_use > 1L) {
    res <- parallel::mclapply(
      seq_len(nrow(task_tbl)),
      function(i) {
        rr <- task_tbl[i, , drop = FALSE]
        run_alfak_fit(
          patient_id = rr$patient_id,
          input_rds = rr$input_rds,
          outdir = rr$outdir,
          minobs = rr$minobs,
          pm = rr$pm,
          nboot = rr$nboot,
          n0 = rr$n0,
          nb = rr$nb,
          benchmark_seed = rr$benchmark_seed,
          parameter_label = rr$parameter_label,
          correct_efflux = rr$correct_efflux,
          nn_prior = rr$nn_prior,
          nn_prior_grid_n = rr$nn_prior_grid_n,
          nn_prior_fit_subset = rr$nn_prior_fit_subset,
          nn_prior_zero_exposure_quantile = rr$nn_prior_zero_exposure_quantile,
          nn_prior_zero_weight_scale = rr$nn_prior_zero_weight_scale,
          nn_prior_zero_weight_cap_ratio = rr$nn_prior_zero_weight_cap_ratio,
          nn_prior_zero_birth_fallback_weight = rr$nn_prior_zero_birth_fallback_weight,
          force_refit = rr$force_refit
        )
      },
      mc.cores = n_cores_use,
      mc.preschedule = FALSE,
      mc.set.seed = FALSE
    )
  } else {
    res <- lapply(seq_len(nrow(task_tbl)), function(i) {
      rr <- task_tbl[i, , drop = FALSE]
      run_alfak_fit(
        patient_id = rr$patient_id,
        input_rds = rr$input_rds,
        outdir = rr$outdir,
        minobs = rr$minobs,
        pm = rr$pm,
        nboot = rr$nboot,
        n0 = rr$n0,
        nb = rr$nb,
        benchmark_seed = rr$benchmark_seed,
        parameter_label = rr$parameter_label,
        correct_efflux = rr$correct_efflux,
        nn_prior = rr$nn_prior,
        nn_prior_grid_n = rr$nn_prior_grid_n,
        nn_prior_fit_subset = rr$nn_prior_fit_subset,
        nn_prior_zero_exposure_quantile = rr$nn_prior_zero_exposure_quantile,
        nn_prior_zero_weight_scale = rr$nn_prior_zero_weight_scale,
        nn_prior_zero_weight_cap_ratio = rr$nn_prior_zero_weight_cap_ratio,
        nn_prior_zero_birth_fallback_weight = rr$nn_prior_zero_birth_fallback_weight,
        force_refit = rr$force_refit
      )
    })
  }

  bind_rows(res)
}

summarize_fit_results <- function(fit_results_tbl, group_cols) {
  if (!nrow(fit_results_tbl)) {
    return(tibble())
  }

  fit_results_tbl %>%
    group_by(across(all_of(group_cols))) %>%
    summarise(
      n_tasks = n(),
      n_ok = sum(status == "ok", na.rm = TRUE),
      n_error = sum(status == "error", na.rm = TRUE),
      success_rate = n_ok / n_tasks,
      mean_xval_r2 = mean(xval_r2, na.rm = TRUE),
      median_xval_r2 = median(xval_r2, na.rm = TRUE),
      mean_elapsed_sec = mean(elapsed_sec, na.rm = TRUE),
      median_elapsed_sec = median(elapsed_sec, na.rm = TRUE),
      total_lambda_endpoint_warnings = sum(lambda_endpoint_warning_count, na.rm = TRUE),
      .groups = "drop"
    )
}

safe_read_rds <- function(path) {
  if (is.na(path) || !nzchar(path) || !file.exists(path)) {
    return(NULL)
  }
  readRDS(path)
}

read_fit_bundle <- function(fit_dir) {
  landscape_obj <- safe_read_rds(file.path(fit_dir, "landscape.Rds"))
  list(
    bootstrap = safe_read_rds(file.path(fit_dir, "bootstrap_res.Rds")),
    landscape = landscape_obj,
    posterior = safe_read_rds(file.path(fit_dir, "landscape_posterior_samples.Rds")),
    xval = safe_read_rds(file.path(fit_dir, "xval.Rds")),
    beneficial = if (!is.null(landscape_obj)) beneficial_cnv_profile(landscape_obj) else NULL
  )
}

rank_parameter_fit_tbl <- function(fit_tbl, parameter_levels = NULL) {
  if (is.null(fit_tbl) || !nrow(fit_tbl)) {
    return(tibble())
  }

  if (is.null(parameter_levels) || !length(parameter_levels)) {
    parameter_levels <- unique(as.character(fit_tbl$parameter_label))
  }

  fit_tbl %>%
    mutate(
      status = as.character(status),
      has_positive_xval = is.finite(xval_r2) & xval_r2 > 0,
      rank_xval_r2 = dplyr::coalesce(xval_r2, -Inf),
      rank_xval_cor = dplyr::coalesce(xval_cor, -Inf),
      rank_xval_rmse = dplyr::coalesce(xval_rmse, Inf),
      rank_xval_mae = dplyr::coalesce(xval_mae, Inf)
    ) %>%
    arrange(
      factor(parameter_label, levels = parameter_levels),
      factor(patient_id, levels = sort_pid_levels(patient_id)),
      dplyr::desc(status == "ok"),
      dplyr::desc(has_positive_xval),
      dplyr::desc(rank_xval_r2),
      rank_xval_rmse,
      rank_xval_mae,
      dplyr::desc(rank_xval_cor),
      minobs,
      pm
    )
}

select_best_parameter_fit_tbl <- function(fit_tbl, parameter_levels = NULL) {
  ranked_tbl <- rank_parameter_fit_tbl(fit_tbl, parameter_levels = parameter_levels)
  if (!nrow(ranked_tbl)) {
    return(tibble())
  }

  best_tbl <- ranked_tbl %>%
    filter(status == "ok") %>%
    group_by(parameter_label, patient_id) %>%
    slice_head(n = 1) %>%
    ungroup() %>%
    select(-has_positive_xval, -rank_xval_r2, -rank_xval_cor, -rank_xval_rmse, -rank_xval_mae)

  best_tbl
}

load_selected_landscapes_by_parameter <- function(selected_fit_tbl, parameter_levels = NULL) {
  if (is.null(selected_fit_tbl) || !nrow(selected_fit_tbl)) {
    return(list())
  }

  if (is.null(parameter_levels) || !length(parameter_levels)) {
    parameter_levels <- unique(as.character(selected_fit_tbl$parameter_label))
  }

  out <- setNames(vector("list", length(parameter_levels)), parameter_levels)
  for (parameter_label_name in parameter_levels) {
    sub_tbl <- selected_fit_tbl %>%
      filter(parameter_label == parameter_label_name) %>%
      arrange(factor(patient_id, levels = sort_pid_levels(patient_id)))

    if (!nrow(sub_tbl)) {
      out[[parameter_label_name]] <- list()
      next
    }

    landscape_list <- lapply(sub_tbl$outdir, function(outdir) {
      read_fit_bundle(outdir)$landscape
    })
    names(landscape_list) <- sub_tbl$patient_id
    out[[parameter_label_name]] <- landscape_list
  }

  out
}

save_beneficial_heatmap_png <- function(beneficial_mat, png_path) {
  if (!nrow(beneficial_mat) || !ncol(beneficial_mat)) {
    return(invisible(png_path))
  }

  panel_d_mat <- beneficial_mat
  panel_d_mat[!is.finite(panel_d_mat)] <- 0
  col_fun_d <- circlize::colorRamp2(c(0, 0.5, 1), c("blue", "#EEEEEE", "red"))

  ht_d <- ComplexHeatmap::Heatmap(
    panel_d_mat,
    name = "Beneficial\nkaryotypes\nproportion",
    col = col_fun_d,
    cluster_rows = nrow(panel_d_mat) > 1,
    cluster_columns = ncol(panel_d_mat) > 1,
    na_col = "#BDBDBD",
    show_row_names = TRUE,
    show_column_names = TRUE,
    row_names_gp = grid::gpar(fontsize = 8),
    column_names_gp = grid::gpar(fontsize = 7),
    rect_gp = grid::gpar(col = "white", lwd = 1)
  )

  plot_height <- max(4.5, 0.28 * nrow(panel_d_mat) + 1.5)
  grDevices::png(png_path, width = 10 * 150, height = plot_height * 150, res = 150)
  ComplexHeatmap::draw(ht_d, heatmap_legend_side = "right")
  grDevices::dev.off()
  invisible(png_path)
}

build_parameter_beneficial_artifacts <- function(selected_landscapes_by_parameter, parameter_levels, tables_dir, figures_dir) {
  if (!length(parameter_levels)) {
    return(tibble())
  }

  bind_rows(lapply(parameter_levels, function(parameter_label_name) {
    landscape_list <- selected_landscapes_by_parameter[[parameter_label_name]]
    if (is.null(landscape_list) || !length(landscape_list)) {
      return(tibble(
        parameter_label = parameter_label_name,
        n_patients = 0L,
        proportion_matrix_path = NA_character_,
        valid_n_matrix_path = NA_character_,
        heatmap_png_path = NA_character_
      ))
    }

    beneficial_profiles <- lapply(landscape_list, beneficial_cnv_profile)
    beneficial_list <- lapply(beneficial_profiles, function(x) x$proportion)
    beneficial_valid_n_list <- lapply(beneficial_profiles, function(x) x$valid_n)

    beneficial_mat <- do.call(rbind, beneficial_list)
    beneficial_mat <- beneficial_mat[sort_pid_levels(rownames(beneficial_mat)), , drop = FALSE]
    beneficial_valid_n_mat <- do.call(rbind, beneficial_valid_n_list)
    beneficial_valid_n_mat <- beneficial_valid_n_mat[sort_pid_levels(rownames(beneficial_valid_n_mat)), , drop = FALSE]

    proportion_stem <- file.path(tables_dir, paste0("parameter_", parameter_label_name, "_beneficial_proportion_matrix"))
    valid_n_stem <- file.path(tables_dir, paste0("parameter_", parameter_label_name, "_beneficial_valid_n_matrix"))
    heatmap_png_path <- file.path(figures_dir, paste0("beneficial_heatmap_", parameter_label_name, ".png"))

    save_table_bundle(
      tibble::rownames_to_column(as.data.frame(beneficial_mat), var = "Patient"),
      proportion_stem
    )
    save_table_bundle(
      tibble::rownames_to_column(as.data.frame(beneficial_valid_n_mat), var = "Patient"),
      valid_n_stem
    )
    save_beneficial_heatmap_png(beneficial_mat, heatmap_png_path)

    tibble(
      parameter_label = parameter_label_name,
      n_patients = nrow(beneficial_mat),
      proportion_matrix_path = paste0(proportion_stem, ".tsv"),
      valid_n_matrix_path = paste0(valid_n_stem, ".tsv"),
      heatmap_png_path = heatmap_png_path
    )
  }))
}

flatten_pair_values <- function(lhs, rhs) {
  if (is.null(lhs) || is.null(rhs)) {
    return(list(lhs = numeric(), rhs = numeric(), lhs_total = 0L, rhs_total = 0L, n_used = 0L))
  }

  if (!is.null(dim(lhs)) && !is.null(dim(rhs))) {
    lhs_mat <- as.matrix(lhs)
    rhs_mat <- as.matrix(rhs)

    if (!is.null(colnames(lhs_mat)) && !is.null(colnames(rhs_mat))) {
      common_cols <- intersect(colnames(lhs_mat), colnames(rhs_mat))
      if (length(common_cols)) {
        lhs_mat <- lhs_mat[, common_cols, drop = FALSE]
        rhs_mat <- rhs_mat[, common_cols, drop = FALSE]
      }
    }

    n_rows <- min(nrow(lhs_mat), nrow(rhs_mat))
    if (n_rows < 1L) {
      return(list(lhs = numeric(), rhs = numeric(), lhs_total = length(lhs_mat), rhs_total = length(rhs_mat), n_used = 0L))
    }
    lhs_vec <- as.numeric(lhs_mat[seq_len(n_rows), , drop = FALSE])
    rhs_vec <- as.numeric(rhs_mat[seq_len(n_rows), , drop = FALSE])
  } else {
    lhs_vec <- as.numeric(lhs)
    rhs_vec <- as.numeric(rhs)
  }

  n_used <- min(length(lhs_vec), length(rhs_vec))
  if (n_used < 1L) {
    return(list(lhs = numeric(), rhs = numeric(), lhs_total = length(lhs_vec), rhs_total = length(rhs_vec), n_used = 0L))
  }

  list(
    lhs = lhs_vec[seq_len(n_used)],
    rhs = rhs_vec[seq_len(n_used)],
    lhs_total = length(lhs_vec),
    rhs_total = length(rhs_vec),
    n_used = n_used
  )
}

compare_numeric_objects <- function(lhs, rhs, metric_name) {
  aligned <- flatten_pair_values(lhs, rhs)
  lhs_vec <- aligned$lhs
  rhs_vec <- aligned$rhs
  if (!length(lhs_vec) || !length(rhs_vec)) {
    return(tibble(
      metric = metric_name,
      n = 0L,
      lhs_mean = NA_real_,
      rhs_mean = NA_real_,
      mean_diff = NA_real_,
      mean_abs_diff = NA_real_,
      max_abs_diff = NA_real_,
      correlation = NA_real_,
      lhs_negative = NA_integer_,
      rhs_negative = NA_integer_,
      lhs_total = aligned$lhs_total,
      rhs_total = aligned$rhs_total
    ))
  }

  diff_vec <- rhs_vec - lhs_vec
  tibble(
    metric = metric_name,
    n = aligned$n_used,
    lhs_mean = mean(lhs_vec, na.rm = TRUE),
    rhs_mean = mean(rhs_vec, na.rm = TRUE),
    mean_diff = mean(diff_vec, na.rm = TRUE),
    mean_abs_diff = mean(abs(diff_vec), na.rm = TRUE),
    max_abs_diff = max(abs(diff_vec), na.rm = TRUE),
    correlation = if (aligned$n_used >= 2L) {
      suppressWarnings(stats::cor(lhs_vec, rhs_vec, use = "complete.obs"))
    } else {
      NA_real_
    },
    lhs_negative = sum(lhs_vec < 0, na.rm = TRUE),
    rhs_negative = sum(rhs_vec < 0, na.rm = TRUE),
    lhs_total = aligned$lhs_total,
    rhs_total = aligned$rhs_total
  )
}

summarize_sign_changes <- function(lhs, rhs, metric_name) {
  aligned <- flatten_pair_values(lhs, rhs)
  lhs_vec <- aligned$lhs
  rhs_vec <- aligned$rhs
  if (!length(lhs_vec) || !length(rhs_vec)) {
    return(tibble(
      metric = metric_name,
      n = 0L,
      lhs_negative = NA_integer_,
      rhs_negative = NA_integer_,
      neg_to_nonneg = NA_integer_,
      nonneg_to_neg = NA_integer_,
      across_zero = NA_integer_
    ))
  }

  tibble(
    metric = metric_name,
    n = aligned$n_used,
    lhs_negative = sum(lhs_vec < 0, na.rm = TRUE),
    rhs_negative = sum(rhs_vec < 0, na.rm = TRUE),
    neg_to_nonneg = sum(lhs_vec < 0 & rhs_vec >= 0, na.rm = TRUE),
    nonneg_to_neg = sum(lhs_vec >= 0 & rhs_vec < 0, na.rm = TRUE),
    across_zero = sum(lhs_vec * rhs_vec < 0, na.rm = TRUE)
  )
}

merge_landscape_pair <- function(lhs_landscape, rhs_landscape) {
  lhs_tbl <- lhs_landscape %>%
    transmute(
      k,
      lhs_mean = mean,
      lhs_median = median,
      lhs_sd = sd,
      lhs_fq = as.logical(fq),
      lhs_nn = as.logical(nn)
    )
  rhs_tbl <- rhs_landscape %>%
    transmute(
      k,
      rhs_mean = mean,
      rhs_median = median,
      rhs_sd = sd,
      rhs_fq = as.logical(fq),
      rhs_nn = as.logical(nn)
    )

  inner_join(lhs_tbl, rhs_tbl, by = "k") %>%
    mutate(
      state_group = case_when(
        lhs_fq | rhs_fq ~ "fq",
        lhs_nn | rhs_nn ~ "nn",
        TRUE ~ "other"
      )
    )
}

summarize_landscape_groups <- function(merged_landscape, lhs_col, rhs_col, metric_name) {
  if (!nrow(merged_landscape)) {
    return(tibble())
  }

  diff_vec <- merged_landscape[[rhs_col]] - merged_landscape[[lhs_col]]
  tmp <- merged_landscape %>%
    mutate(
      lhs_value = .data[[lhs_col]],
      rhs_value = .data[[rhs_col]],
      diff_value = diff_vec
    ) %>%
    group_by(state_group) %>%
    summarise(
      metric = metric_name,
      n = n(),
      lhs_mean = mean(lhs_value, na.rm = TRUE),
      rhs_mean = mean(rhs_value, na.rm = TRUE),
      mean_diff = mean(diff_value, na.rm = TRUE),
      mean_abs_diff = mean(abs(diff_value), na.rm = TRUE),
      max_abs_diff = max(abs(diff_value), na.rm = TRUE),
      correlation = if (n() >= 2L) {
        suppressWarnings(stats::cor(lhs_value, rhs_value, use = "complete.obs"))
      } else {
        NA_real_
      },
      .groups = "drop"
    )

  tmp %>% select(metric, state_group, everything())
}

top_landscape_shifts <- function(merged_landscape, top_n = 15L) {
  if (!nrow(merged_landscape)) {
    return(tibble())
  }

  merged_landscape %>%
    transmute(
      k,
      state_group,
      lhs_mean,
      rhs_mean,
      diff = rhs_mean - lhs_mean
    ) %>%
    arrange(desc(abs(diff))) %>%
    slice_head(n = top_n)
}

compare_fit_pair <- function(lhs_dir,
                             rhs_dir,
                             patient_id,
                             minobs,
                             pm,
                             lhs_label,
                             rhs_label,
                             lhs_setting,
                             rhs_setting,
                             comparison_set,
                             top_n = 15L,
                             include_posterior = TRUE) {
  lhs_bundle <- read_fit_bundle(lhs_dir)
  rhs_bundle <- read_fit_bundle(rhs_dir)

  component_tbl <- bind_rows(
    compare_numeric_objects(lhs_bundle$bootstrap$initial_fitness, rhs_bundle$bootstrap$initial_fitness, "bootstrap_initial_fitness"),
    compare_numeric_objects(lhs_bundle$bootstrap$final_fitness, rhs_bundle$bootstrap$final_fitness, "bootstrap_final_fitness"),
    compare_numeric_objects(lhs_bundle$bootstrap$initial_frequencies, rhs_bundle$bootstrap$initial_frequencies, "bootstrap_initial_frequencies"),
    compare_numeric_objects(lhs_bundle$bootstrap$final_frequencies, rhs_bundle$bootstrap$final_frequencies, "bootstrap_final_frequencies"),
    compare_numeric_objects(lhs_bundle$bootstrap$nn_fitness, rhs_bundle$bootstrap$nn_fitness, "bootstrap_nn_fitness"),
    compare_numeric_objects(lhs_bundle$landscape$mean, rhs_bundle$landscape$mean, "landscape_mean"),
    compare_numeric_objects(lhs_bundle$landscape$median, rhs_bundle$landscape$median, "landscape_median"),
    compare_numeric_objects(lhs_bundle$landscape$sd, rhs_bundle$landscape$sd, "landscape_sd"),
    compare_numeric_objects(lhs_bundle$beneficial$proportion, rhs_bundle$beneficial$proportion, "beneficial_proportion"),
    compare_numeric_objects(lhs_bundle$beneficial$valid_n, rhs_bundle$beneficial$valid_n, "beneficial_valid_n"),
    if (include_posterior) {
      compare_numeric_objects(lhs_bundle$posterior, rhs_bundle$posterior, "landscape_posterior_samples")
    },
    compare_numeric_objects(
      extract_xval_metrics(lhs_bundle$xval)$xval_r2,
      extract_xval_metrics(rhs_bundle$xval)$xval_r2,
      "xval_r2"
    )
  ) %>%
    mutate(
      comparison_set = comparison_set,
      patient_id = patient_id,
      minobs = minobs,
      pm = pm,
      lhs_label = lhs_label,
      rhs_label = rhs_label,
      lhs_setting = lhs_setting,
      rhs_setting = rhs_setting
    ) %>%
    relocate(comparison_set, patient_id, minobs, pm, lhs_label, rhs_label, lhs_setting, rhs_setting, metric)

  sign_tbl <- bind_rows(
    summarize_sign_changes(lhs_bundle$bootstrap$nn_fitness, rhs_bundle$bootstrap$nn_fitness, "bootstrap_nn_fitness"),
    summarize_sign_changes(lhs_bundle$landscape$mean, rhs_bundle$landscape$mean, "landscape_mean"),
    summarize_sign_changes(lhs_bundle$landscape$median, rhs_bundle$landscape$median, "landscape_median")
  ) %>%
    mutate(
      comparison_set = comparison_set,
      patient_id = patient_id,
      minobs = minobs,
      pm = pm,
      lhs_label = lhs_label,
      rhs_label = rhs_label,
      lhs_setting = lhs_setting,
      rhs_setting = rhs_setting
    ) %>%
    relocate(comparison_set, patient_id, minobs, pm, lhs_label, rhs_label, lhs_setting, rhs_setting, metric)

  merged_landscape <- if (!is.null(lhs_bundle$landscape) && !is.null(rhs_bundle$landscape)) {
    merge_landscape_pair(lhs_bundle$landscape, rhs_bundle$landscape)
  } else {
    tibble()
  }
  landscape_group_tbl <- if (nrow(merged_landscape)) {
    bind_rows(
      summarize_landscape_groups(merged_landscape, "lhs_mean", "rhs_mean", "landscape_mean"),
      summarize_landscape_groups(merged_landscape, "lhs_median", "rhs_median", "landscape_median"),
      summarize_landscape_groups(merged_landscape, "lhs_sd", "rhs_sd", "landscape_sd")
    ) %>%
      mutate(
        comparison_set = comparison_set,
        patient_id = patient_id,
        minobs = minobs,
        pm = pm,
        lhs_label = lhs_label,
        rhs_label = rhs_label,
        lhs_setting = lhs_setting,
        rhs_setting = rhs_setting
      ) %>%
      relocate(comparison_set, patient_id, minobs, pm, lhs_label, rhs_label, lhs_setting, rhs_setting, metric, state_group)
  } else {
    tibble()
  }

  top_shift_tbl <- if (nrow(merged_landscape)) {
    top_landscape_shifts(merged_landscape, top_n = top_n) %>%
      mutate(
        comparison_set = comparison_set,
        patient_id = patient_id,
        minobs = minobs,
        pm = pm,
        lhs_label = lhs_label,
        rhs_label = rhs_label,
        lhs_setting = lhs_setting,
        rhs_setting = rhs_setting
      ) %>%
      relocate(comparison_set, patient_id, minobs, pm, lhs_label, rhs_label, lhs_setting, rhs_setting)
  } else {
    tibble()
  }

  list(
    component = component_tbl,
    sign = sign_tbl,
    landscape_group = landscape_group_tbl,
    top_shift = top_shift_tbl
  )
}

build_pairwise_comparisons <- function(results_tbl,
                                       pair_values,
                                       setting_col,
                                       comparison_set,
                                       top_n = 15L,
                                       include_posterior = TRUE) {
  if (!nrow(results_tbl) || length(pair_values) < 2L) {
    return(list(
      component = tibble(),
      sign = tibble(),
      landscape_group = tibble(),
      top_shift = tibble()
    ))
  }

  pair_defs <- combn(pair_values, 2, simplify = FALSE)
  all_component <- list()
  all_sign <- list()
  all_group <- list()
  all_shift <- list()

  for (pair in pair_defs) {
    lhs_val <- pair[[1]]
    rhs_val <- pair[[2]]

    lhs_idx <- results_tbl %>%
      filter(status == "ok", .data[[setting_col]] == lhs_val) %>%
      transmute(
        patient_id,
        minobs,
        pm,
        lhs_setting = .data[[setting_col]],
        lhs_label = if (setting_col == "nn_prior_grid_n") paste0("grid_", .data[[setting_col]]) else as.character(.data[[setting_col]]),
        outdir_lhs = outdir
      )

    rhs_idx <- results_tbl %>%
      filter(status == "ok", .data[[setting_col]] == rhs_val) %>%
      transmute(
        patient_id,
        minobs,
        pm,
        rhs_setting = .data[[setting_col]],
        rhs_label = if (setting_col == "nn_prior_grid_n") paste0("grid_", .data[[setting_col]]) else as.character(.data[[setting_col]]),
        outdir_rhs = outdir
      )

    joined <- inner_join(lhs_idx, rhs_idx, by = c("patient_id", "minobs", "pm"))
    if (!nrow(joined)) {
      next
    }

    pair_res <- lapply(seq_len(nrow(joined)), function(i) {
      rr <- joined[i, , drop = FALSE]
      compare_fit_pair(
        lhs_dir = rr$outdir_lhs,
        rhs_dir = rr$outdir_rhs,
        patient_id = rr$patient_id,
        minobs = rr$minobs,
        pm = rr$pm,
        lhs_label = rr$lhs_label,
        rhs_label = rr$rhs_label,
        lhs_setting = rr$lhs_setting,
        rhs_setting = rr$rhs_setting,
        comparison_set = comparison_set,
        top_n = top_n,
        include_posterior = include_posterior
      )
    })

    all_component <- c(all_component, lapply(pair_res, `[[`, "component"))
    all_sign <- c(all_sign, lapply(pair_res, `[[`, "sign"))
    all_group <- c(all_group, lapply(pair_res, `[[`, "landscape_group"))
    all_shift <- c(all_shift, lapply(pair_res, `[[`, "top_shift"))
  }

  list(
    component = bind_rows(all_component),
    sign = bind_rows(all_sign),
    landscape_group = bind_rows(all_group),
    top_shift = bind_rows(all_shift)
  )
}

```

## Build ALFA-K Inputs

```{r build-inputs}
meta_tbl <- readxl::read_xlsx(meta_path)
input_index_tbl <- build_benchmark_inputs(
  meta_tbl = meta_tbl,
  base_dir = data_dir,
  input_dir = input_dir,
  rebuild_inputs = rebuild_inputs_use,
  patient_subset = patient_subset_use
)
save_table_bundle(input_index_tbl, file.path(tables_dir, "benchmark_input_index"))

input_index_tbl
```

```{r input-summary}
render_tbl(
  input_index_tbl %>%
    select(patient_id, n_timepoints, n_karyotypes, total_cells, delta_time_labels, stage_sequence),
  caption = "Benchmark input summary"
)
```

## Parameter-Label Benchmark

```{r parameter-grid-text, results='asis'}
cat(sprintf("**Fixed benchmark-wide `nn_prior_grid_n`: `%d`**\n\n", selected_grid_n_use))
cat(sprintf("**Compared `nn_prior` labels:** `%s`\n\n", paste(parameter_levels_use, collapse = "`, `")))
```

```{r parameter-tasks}
parameter_spec_tbl <- build_parameter_spec_tbl(
  parameter_labels = parameter_levels_use
)
save_table_bundle(parameter_spec_tbl, file.path(tables_dir, "parameter_spec"))

parameter_tasks_tbl <- build_parameter_tasks(
  input_index_tbl = input_index_tbl,
  fit_root = fit_dir,
  minobs_values = minobs_values_use,
  pm_values = pm_values_use,
  parameter_spec_tbl = parameter_spec_tbl,
  nn_prior_grid_n = selected_grid_n_use,
  nn_prior_fit_subset = nn_prior_fit_subset_use,
  nn_prior_zero_exposure_quantile = nn_prior_zero_exposure_quantile_use,
  nn_prior_zero_weight_scale = nn_prior_zero_weight_scale_use,
  nn_prior_zero_weight_cap_ratio = nn_prior_zero_weight_cap_ratio_use,
  nn_prior_zero_birth_fallback_weight = nn_prior_zero_birth_fallback_weight_use,
  nboot = nboot_use,
  n0 = n0_use,
  nb = nb_use,
  benchmark_seed = benchmark_seed_use,
  correct_efflux = correct_efflux_use,
  force_refit = force_refit_use
)
save_table_bundle(parameter_tasks_tbl, file.path(tables_dir, "parameter_tasks"))

parameter_tasks_tbl %>%
  count(parameter_label, name = "n_tasks")
```

```{r parameter-run}
parameter_results_path <- file.path(tables_dir, "parameter_fit_results")

if (run_benchmark_use && nrow(parameter_tasks_tbl)) {
  parameter_results_all_tbl <- run_task_table_parallel(parameter_tasks_tbl)
} else {
  parameter_results_all_tbl <- load_saved_table(parameter_results_path)
  if (is.null(parameter_results_all_tbl)) {
    parameter_results_all_tbl <- tibble()
  }
}

parameter_results_all_tbl <- reconcile_fit_results_tbl(
  fit_results_tbl = parameter_results_all_tbl,
  task_tbl = parameter_tasks_tbl
)
parameter_results_all_tbl <- parameter_results_all_tbl %>%
  filter(parameter_label %in% parameter_levels_use)
save_table_bundle(parameter_results_all_tbl, parameter_results_path)

if (nrow(parameter_results_all_tbl)) {
  parameter_results_all_tbl <- parameter_results_all_tbl %>%
    arrange(factor(patient_id, levels = sort_pid_levels(patient_id)), minobs, pm, factor(parameter_label, levels = parameter_levels_use))
}

save_table_bundle(parameter_results_all_tbl, file.path(tables_dir, "parameter_all_results"))
parameter_results_all_tbl
```

```{r parameter-summary}
parameter_fit_summary_tbl <- summarize_fit_results(parameter_results_all_tbl, group_cols = c("parameter_label"))
save_table_bundle(parameter_fit_summary_tbl, file.path(tables_dir, "parameter_fit_summary"))

if (nrow(parameter_fit_summary_tbl)) {
  render_tbl(parameter_fit_summary_tbl, caption = "Fit summary by parameter label")
} else {
  render_tbl(empty_note_tbl("No parameter-label fit results are available yet."), caption = "Fit summary by parameter label")
}
```

## Selected Fits for All-Sample Panels

```{r parameter-selected-fits}
parameter_selected_fit_tbl <- select_best_parameter_fit_tbl(
  parameter_results_all_tbl,
  parameter_levels = parameter_levels_use
)
save_table_bundle(parameter_selected_fit_tbl, file.path(tables_dir, "parameter_selected_fit_index"))

parameter_selected_landscapes <- load_selected_landscapes_by_parameter(
  parameter_selected_fit_tbl,
  parameter_levels = parameter_levels_use
)

parameter_beneficial_artifact_tbl <- build_parameter_beneficial_artifacts(
  selected_landscapes_by_parameter = parameter_selected_landscapes,
  parameter_levels = parameter_levels_use,
  tables_dir = tables_dir,
  figures_dir = figures_dir
)
save_table_bundle(parameter_beneficial_artifact_tbl, file.path(tables_dir, "parameter_beneficial_all_patient_artifacts"))

parameter_selected_fit_tbl
```

```{r parameter-selected-fits-render}
parameter_selected_fit_render_tbl <- if (nrow(parameter_selected_fit_tbl)) {
  parameter_selected_fit_tbl %>%
    transmute(
      parameter_label,
      patient_id,
      nn_prior,
      minobs,
      pm,
      xval_r2,
      elapsed_sec,
      cached
    )
} else {
  tibble()
}

if (nrow(parameter_selected_fit_render_tbl)) {
  render_tbl(
    parameter_selected_fit_render_tbl,
    caption = "Best successful fit per patient within each parameter label, used for all-sample heatmaps"
  )
} else {
  render_tbl(
    empty_note_tbl("No successful patient-level fits are available for the all-sample heatmaps yet."),
    caption = "Best successful fit per patient within each parameter label"
  )
}
```

## All-Sample Beneficial Karyotype Heatmaps

```{r parameter-beneficial-artifacts-render}
parameter_beneficial_summary_render_tbl <- if (nrow(parameter_beneficial_artifact_tbl)) {
  parameter_beneficial_artifact_tbl %>%
    select(parameter_label, n_patients)
} else {
  tibble()
}

if (nrow(parameter_beneficial_summary_render_tbl)) {
  render_tbl(
    parameter_beneficial_summary_render_tbl,
    caption = "All-sample beneficial-heatmap coverage by parameter label"
  )
} else {
  render_tbl(
    empty_note_tbl("No all-sample beneficial heatmaps are available yet."),
    caption = "All-sample beneficial-heatmap coverage by parameter label"
  )
}
```

```{r parameter-beneficial-heatmaps-display, echo=FALSE, results='asis'}
if (nrow(parameter_beneficial_artifact_tbl)) {
  for (i in seq_len(nrow(parameter_beneficial_artifact_tbl))) {
    rr <- parameter_beneficial_artifact_tbl[i, , drop = FALSE]
    cat("### ", rr$parameter_label, "\n\n", sep = "")
    if (is.na(rr$heatmap_png_path) || !nzchar(rr$heatmap_png_path) || !file.exists(rr$heatmap_png_path)) {
      cat("No all-sample beneficial heatmap is available for this parameter label.\n\n")
      next
    }
    emit_report_image(
      rr$heatmap_png_path,
      alt = paste("All-sample beneficial heatmap for", rr$parameter_label)
    )
  }
} else {
  cat("No all-sample beneficial heatmaps are available yet.\n")
}
```

## Parameter-Label Pairwise Comparisons

```{r parameter-compare}
parameter_pair_values_use <- parameter_levels_use

parameter_pair_results <- build_pairwise_comparisons(
  results_tbl = parameter_results_all_tbl,
  pair_values = parameter_pair_values_use,
  setting_col = "parameter_label",
  comparison_set = "parameter",
  top_n = top_shift_n_use,
  include_posterior = include_posterior_use
)

save_table_bundle(parameter_pair_results$component, file.path(tables_dir, "parameter_pair_component_summary"))
save_table_bundle(parameter_pair_results$sign, file.path(tables_dir, "parameter_pair_sign_summary"))
save_table_bundle(parameter_pair_results$landscape_group, file.path(tables_dir, "parameter_pair_landscape_group_summary"))
save_table_bundle(parameter_pair_results$top_shift, file.path(tables_dir, "parameter_pair_top_landscape_shifts"))

parameter_pair_overview_tbl <- if (nrow(parameter_pair_results$component)) {
  parameter_pair_results$component %>%
    filter(metric %in% c("landscape_mean", "bootstrap_nn_fitness", "xval_r2")) %>%
    group_by(lhs_label, rhs_label, metric) %>%
    summarise(
      n_pairs = n(),
      mean_abs_diff = mean(mean_abs_diff, na.rm = TRUE),
      max_abs_diff = max(max_abs_diff, na.rm = TRUE),
      mean_correlation = mean(correlation, na.rm = TRUE),
      .groups = "drop"
    )
} else {
  tibble()
}
save_table_bundle(parameter_pair_overview_tbl, file.path(tables_dir, "parameter_pair_overview"))

parameter_pair_overview_tbl
```

```{r parameter-overview-render}
if (nrow(parameter_pair_overview_tbl)) {
  render_tbl(parameter_pair_overview_tbl, caption = "Pairwise comparison overview by parameter label")
} else {
  render_tbl(empty_note_tbl("No successful parameter-label pairwise comparisons are available yet."), caption = "Pairwise comparison overview by parameter label")
}
```

```{r parameter-top-shifts}
parameter_top_shift_preview_tbl <- if (nrow(parameter_pair_results$top_shift)) {
  parameter_pair_results$top_shift %>%
    group_by(lhs_label, rhs_label) %>%
    slice_head(n = min(10L, top_shift_n_use)) %>%
    ungroup()
} else {
  tibble()
}

if (nrow(parameter_top_shift_preview_tbl)) {
  render_tbl(parameter_top_shift_preview_tbl, caption = "Top landscape shifts for parameter-label comparisons")
} else {
  render_tbl(empty_note_tbl("No top-shift table is available for the parameter-label benchmark."), caption = "Top landscape shifts for parameter-label comparisons")
}
```

## Global Landscape and Beneficial Comparison

```{r parameter-global-landscape}
parameter_global_landscape_overview_tbl <- if (nrow(parameter_pair_results$component)) {
  parameter_pair_results$component %>%
    filter(metric %in% c("landscape_mean", "landscape_median", "landscape_sd", "beneficial_proportion", "beneficial_valid_n")) %>%
    group_by(lhs_label, rhs_label, metric) %>%
    summarise(
      n_pairs = n(),
      mean_diff = mean(mean_diff, na.rm = TRUE),
      mean_abs_diff = mean(mean_abs_diff, na.rm = TRUE),
      max_abs_diff = max(max_abs_diff, na.rm = TRUE),
      mean_correlation = mean(correlation, na.rm = TRUE),
      .groups = "drop"
    )
} else {
  tibble()
}
save_table_bundle(parameter_global_landscape_overview_tbl, file.path(tables_dir, "parameter_global_landscape_overview"))

parameter_global_landscape_state_tbl <- if (nrow(parameter_pair_results$landscape_group)) {
  parameter_pair_results$landscape_group %>%
    filter(metric == "landscape_mean") %>%
    group_by(lhs_label, rhs_label, state_group) %>%
    summarise(
      n_pairs = n(),
      mean_diff = mean(mean_diff, na.rm = TRUE),
      mean_abs_diff = mean(mean_abs_diff, na.rm = TRUE),
      max_abs_diff = max(max_abs_diff, na.rm = TRUE),
      mean_correlation = mean(correlation, na.rm = TRUE),
      .groups = "drop"
    )
} else {
  tibble()
}
save_table_bundle(parameter_global_landscape_state_tbl, file.path(tables_dir, "parameter_global_landscape_state_summary"))

list(
  landscape_overview = parameter_global_landscape_overview_tbl,
  landscape_by_state = parameter_global_landscape_state_tbl
)
```

```{r parameter-global-landscape-render}
if (nrow(parameter_global_landscape_overview_tbl)) {
  render_tbl(parameter_global_landscape_overview_tbl, caption = "Benchmark-wide landscape and beneficial-proportion comparison summary")
} else {
  render_tbl(empty_note_tbl("No benchmark-wide landscape summary is available yet."), caption = "Benchmark-wide landscape summary")
}
```

```{r parameter-global-landscape-state-render}
if (nrow(parameter_global_landscape_state_tbl)) {
  render_tbl(parameter_global_landscape_state_tbl, caption = "Benchmark-wide landscape mean differences split by fq / nn / other states")
} else {
  render_tbl(empty_note_tbl("No state-group landscape summary is available for the benchmark yet."), caption = "Benchmark-wide landscape mean by state group")
}
```

```{r parameter-figures, eval=render_figures_use}
if (nrow(parameter_fit_summary_tbl)) {
  p_parameter_runtime <- ggplot(parameter_results_all_tbl %>% filter(status == "ok") %>% mutate(parameter_label = factor(parameter_label, levels = parameter_levels_use)), aes(x = parameter_label, y = elapsed_sec)) +
    geom_boxplot(outlier.shape = NA, fill = "#E8D8E8", color = "#633A63") +
    geom_jitter(width = 0.15, alpha = 0.5, size = 1.6, color = "#633A63") +
    labs(
      title = paste0("Runtime by parameter label (grid = ", selected_grid_n_use, ")"),
      x = "Parameter label",
      y = "Elapsed seconds"
    ) +
    theme_bw(base_size = 12) +
    theme(axis.text.x = element_text(angle = 20, hjust = 1))
  print(p_parameter_runtime)
  ggsave(file.path(figures_dir, "runtime_by_parameter_label.png"), p_parameter_runtime, width = 9, height = 5, dpi = 150)

  p_parameter_xval <- ggplot(parameter_results_all_tbl %>% filter(status == "ok") %>% mutate(parameter_label = factor(parameter_label, levels = parameter_levels_use)), aes(x = parameter_label, y = xval_r2)) +
    geom_boxplot(outlier.shape = NA, fill = "#D8EBCF", color = "#31572C") +
    geom_jitter(width = 0.15, alpha = 0.5, size = 1.6, color = "#31572C") +
    labs(
      title = paste0("xval by parameter label (grid = ", selected_grid_n_use, ")"),
      x = "Parameter label",
      y = "xval_r2"
    ) +
    theme_bw(base_size = 12) +
    theme(axis.text.x = element_text(angle = 20, hjust = 1))
  print(p_parameter_xval)
  ggsave(file.path(figures_dir, "xval_by_parameter_label.png"), p_parameter_xval, width = 9, height = 5, dpi = 150)
}

if (nrow(parameter_global_landscape_overview_tbl)) {
  p_parameter_global_diff <- parameter_global_landscape_overview_tbl %>%
    mutate(comparison = paste(lhs_label, rhs_label, sep = " vs ")) %>%
    ggplot(aes(x = comparison, y = mean_abs_diff, fill = comparison)) +
    geom_col(width = 0.7, alpha = 0.9, show.legend = FALSE) +
    facet_wrap(~ metric, scales = "free_y", ncol = 1) +
    labs(
      title = "Benchmark-wide mean absolute differences",
      x = NULL,
      y = "Mean absolute difference"
    ) +
    theme_bw(base_size = 12) +
    theme(axis.text.x = element_text(angle = 20, hjust = 1))
  print(p_parameter_global_diff)
  ggsave(file.path(figures_dir, "global_landscape_and_beneficial_diff.png"), p_parameter_global_diff, width = 9, height = 10, dpi = 150)
}

if (nrow(parameter_global_landscape_state_tbl)) {
  p_parameter_state_diff <- parameter_global_landscape_state_tbl %>%
    mutate(comparison = paste(lhs_label, rhs_label, sep = " vs ")) %>%
    ggplot(aes(x = state_group, y = mean_abs_diff, fill = comparison)) +
    geom_col(position = position_dodge(width = 0.75), width = 0.65) +
    labs(
      title = "Benchmark-wide landscape mean differences by state group",
      x = "State group",
      y = "Mean absolute difference",
      fill = "Comparison"
    ) +
    theme_bw(base_size = 12)
  print(p_parameter_state_diff)
  ggsave(file.path(figures_dir, "global_landscape_state_diff.png"), p_parameter_state_diff, width = 9, height = 5, dpi = 150)
}
```

## Focus Patient Four-Way Comparison

```{r focus-setup}
focus_fit_index_tbl <- parameter_results_all_tbl %>%
  filter(
    patient_id == focus_pid_use,
    minobs == focus_minobs_use,
    dplyr::near(pm, focus_pm_use),
    parameter_label %in% parameter_levels_use
  ) %>%
  distinct(parameter_label, .keep_all = TRUE) %>%
  arrange(match(parameter_label, parameter_levels_use))
save_table_bundle(focus_fit_index_tbl, file.path(tables_dir, "focus_parameter_fit_index"))

focus_parameter_bundles <- build_focus_parameter_bundles(
  results_tbl = parameter_results_all_tbl,
  focus_pid = focus_pid_use,
  focus_minobs = focus_minobs_use,
  focus_pm = focus_pm_use,
  parameter_levels = parameter_levels_use
)
focus_parameter_levels_available <- names(focus_parameter_bundles)

focus_landscape_long_tbl <- landscape_long_from_bundles(focus_parameter_bundles)
focus_landscape_summary_tbl <- summarize_landscape_by_parameter(focus_landscape_long_tbl)
focus_landscape_variation_tbl <- summarize_focus_landscape_variation(
  focus_landscape_long_tbl,
  parameter_levels = focus_parameter_levels_available,
  top_n = top_shift_n_use
)
focus_parity_tbl <- build_focus_parity_tbl(
  focus_parameter_bundles,
  parameter_levels = focus_parameter_levels_available,
  value_col = "mean"
)

focus_pairwise_component_tbl <- if (nrow(parameter_pair_results$component)) {
  parameter_pair_results$component %>%
    filter(
      patient_id == focus_pid_use,
      minobs == focus_minobs_use,
      dplyr::near(pm, focus_pm_use),
      metric %in% c("landscape_mean", "landscape_median", "landscape_sd", "xval_r2", "beneficial_proportion", "beneficial_valid_n")
    ) %>%
    arrange(match(lhs_label, focus_parameter_levels_available), match(rhs_label, focus_parameter_levels_available), metric)
} else {
  tibble()
}
save_table_bundle(focus_pairwise_component_tbl, file.path(tables_dir, "focus_pairwise_component_summary"))

focus_pairwise_landscape_group_tbl <- if (nrow(parameter_pair_results$landscape_group)) {
  parameter_pair_results$landscape_group %>%
    filter(
      patient_id == focus_pid_use,
      minobs == focus_minobs_use,
      dplyr::near(pm, focus_pm_use),
      metric == "landscape_mean"
    ) %>%
    arrange(match(lhs_label, focus_parameter_levels_available), match(rhs_label, focus_parameter_levels_available), state_group)
} else {
  tibble()
}
save_table_bundle(focus_pairwise_landscape_group_tbl, file.path(tables_dir, "focus_pairwise_landscape_group_summary"))

focus_top_shift_tbl <- if (nrow(parameter_pair_results$top_shift)) {
  parameter_pair_results$top_shift %>%
    filter(
      patient_id == focus_pid_use,
      minobs == focus_minobs_use,
      dplyr::near(pm, focus_pm_use)
    ) %>%
    group_by(lhs_label, rhs_label) %>%
    slice_head(n = min(10L, top_shift_n_use)) %>%
    ungroup()
} else {
  tibble()
}
save_table_bundle(focus_top_shift_tbl, file.path(tables_dir, "focus_pairwise_top_landscape_shifts"))

focus_beneficial_profiles <- lapply(focus_parameter_levels_available, function(parameter_name) {
  focus_parameter_bundles[[parameter_name]]$beneficial
})
names(focus_beneficial_profiles) <- focus_parameter_levels_available

focus_beneficial_long_tbl <- beneficial_long_from_profiles(
  beneficial_profiles = focus_beneficial_profiles,
  parameter_levels = focus_parameter_levels_available
)
focus_beneficial_summary_tbl <- summarize_beneficial_by_parameter(focus_beneficial_long_tbl)
focus_beneficial_shift_tbl <- build_focus_beneficial_shift_tbl(
  focus_beneficial_long_tbl,
  parameter_levels = focus_parameter_levels_available,
  top_n = top_shift_n_use
)
focus_beneficial_proportion_matrix_tbl <- if (nrow(focus_beneficial_long_tbl)) {
  focus_beneficial_long_tbl %>%
    mutate(move = as.character(move)) %>%
    select(parameter_label, move, proportion) %>%
    pivot_wider(names_from = move, values_from = proportion)
} else {
  tibble()
}
focus_beneficial_valid_n_matrix_tbl <- if (nrow(focus_beneficial_long_tbl)) {
  focus_beneficial_long_tbl %>%
    mutate(move = as.character(move)) %>%
    select(parameter_label, move, valid_n) %>%
    pivot_wider(names_from = move, values_from = valid_n)
} else {
  tibble()
}

focus_umap_parameter_levels <- intersect(
  parameter_levels_use,
  focus_parameter_levels_available
)
focus_umap_artifact_tbl <- bind_rows(lapply(focus_umap_parameter_levels, function(parameter_label_name) {
  landscape_df <- focus_parameter_bundles[[parameter_label_name]]$landscape
  if (is.null(landscape_df) || !nrow(landscape_df)) {
    return(tibble(
      parameter_label = parameter_label_name,
      scale_mode = "relative",
      n_karyotypes = 0L,
      png_path = NA_character_
    ))
  }

  plot_obj <- make_focus_umap_plot(
    landscape_df = landscape_df,
    patient_id = focus_pid_use,
    parameter_label = parameter_label_name,
    scale_mode = "relative"
  )
  png_path <- file.path(figures_dir, paste0("focus_", focus_pid_use, "_", parameter_label_name, "_relative_umap.png"))
  ggsave(png_path, plot_obj, width = 5.5, height = 5, dpi = 150)

  tibble(
    parameter_label = parameter_label_name,
    scale_mode = "relative",
    n_karyotypes = nrow(landscape_df),
    png_path = png_path
  )
}))

save_table_bundle(focus_landscape_summary_tbl, file.path(tables_dir, "focus_landscape_summary"))
save_table_bundle(focus_landscape_variation_tbl, file.path(tables_dir, "focus_top_variable_karyotypes"))
save_table_bundle(focus_beneficial_summary_tbl, file.path(tables_dir, "focus_beneficial_summary"))
save_table_bundle(focus_beneficial_shift_tbl, file.path(tables_dir, "focus_beneficial_top_shifts"))
save_table_bundle(focus_beneficial_proportion_matrix_tbl, file.path(tables_dir, "focus_beneficial_proportion_matrix"))
save_table_bundle(focus_beneficial_valid_n_matrix_tbl, file.path(tables_dir, "focus_beneficial_valid_n_matrix"))
save_table_bundle(focus_umap_artifact_tbl, file.path(tables_dir, "focus_umap_artifacts"))

list(
  focus_fit_index = focus_fit_index_tbl,
  focus_landscape_summary = focus_landscape_summary_tbl,
  focus_pairwise_component = focus_pairwise_component_tbl,
  focus_beneficial_summary = focus_beneficial_summary_tbl,
  focus_umap_artifacts = focus_umap_artifact_tbl
)
```

```{r focus-context, results='asis'}
cat(
  sprintf(
    "Focus comparison uses patient `%s`, `minobs = %d`, and `pm = %s`.\n\n",
    focus_pid_use,
    focus_minobs_use,
    pm_to_label(focus_pm_use)
  )
)
```

### Focus Fit Index

```{r focus-fit-index-render}
focus_fit_index_render_tbl <- if (nrow(focus_fit_index_tbl)) {
  focus_fit_index_tbl %>%
    transmute(
      parameter_label,
      nn_prior,
      status,
      xval_r2,
      elapsed_sec,
      cached,
      warning_count,
      error_message
    )
} else {
  tibble()
}

if (nrow(focus_fit_index_render_tbl)) {
  render_tbl(focus_fit_index_render_tbl, caption = "Focus patient fit index by parameter label")
} else {
  render_tbl(empty_note_tbl("The focus patient combination is not available in the benchmark results yet."), caption = "Focus patient fit index")
}
```

### Focus Landscape Summary

```{r focus-landscape-summary-render}
if (nrow(focus_landscape_summary_tbl)) {
  render_tbl(focus_landscape_summary_tbl, caption = "Focus patient landscape summary by parameter label")
} else {
  render_tbl(empty_note_tbl("No focus-patient landscape summary is available yet."), caption = "Focus patient landscape summary")
}
```

### Focus UMAP

```{r focus-umap-render, echo=FALSE, results='asis'}
if (nrow(focus_umap_artifact_tbl)) {
  for (i in seq_len(nrow(focus_umap_artifact_tbl))) {
    rr <- focus_umap_artifact_tbl[i, , drop = FALSE]
    cat("#### ", rr$parameter_label, "\n\n", sep = "")
    if (is.na(rr$png_path) || !nzchar(rr$png_path) || !file.exists(rr$png_path)) {
      cat("No UMAP is available for this parameter label.\n\n")
      next
    }
    emit_report_image(
      rr$png_path,
      alt = paste("Focus patient UMAP for", rr$parameter_label)
    )
  }
} else {
  cat("No focus-patient UMAP plots are available yet.\n")
}
```

```{r focus-pairwise-component-render}
if (nrow(focus_pairwise_component_tbl)) {
  render_tbl(focus_pairwise_component_tbl, caption = "Focus patient pairwise landscape, xval, and beneficial-proportion summary")
} else {
  render_tbl(empty_note_tbl("No pairwise focus-patient comparison is available yet."), caption = "Focus patient pairwise comparison summary")
}
```

```{r focus-landscape-group-render}
if (nrow(focus_pairwise_landscape_group_tbl)) {
  render_tbl(focus_pairwise_landscape_group_tbl, caption = "Focus patient pairwise landscape mean differences split by fq / nn / other")
} else {
  render_tbl(empty_note_tbl("No focus-patient state-group landscape summary is available yet."), caption = "Focus patient landscape mean by state group")
}
```

```{r focus-top-variable-karyotypes-render}
if (nrow(focus_landscape_variation_tbl)) {
  render_tbl(focus_landscape_variation_tbl, caption = "Focus patient karyotypes with the largest four-way landscape mean spread")
} else {
  render_tbl(empty_note_tbl("No focus-patient landscape spread table is available yet."), caption = "Focus patient top variable karyotypes")
}
```

```{r focus-top-shifts-render}
if (nrow(focus_top_shift_tbl)) {
  render_tbl(focus_top_shift_tbl, caption = "Top focus-patient pairwise landscape mean shifts")
} else {
  render_tbl(empty_note_tbl("No focus-patient top-shift table is available yet."), caption = "Focus patient top landscape shifts")
}
```

### Focus Beneficial Karyotype Proportion

```{r focus-beneficial-summary-render}
if (nrow(focus_beneficial_summary_tbl)) {
  render_tbl(focus_beneficial_summary_tbl, caption = "Focus patient beneficial-karyotype summary by parameter label")
} else {
  render_tbl(empty_note_tbl("No focus-patient beneficial-karyotype summary is available yet."), caption = "Focus patient beneficial-karyotype summary")
}
```

```{r focus-beneficial-shifts-render}
if (nrow(focus_beneficial_shift_tbl)) {
  render_tbl(focus_beneficial_shift_tbl, caption = "Focus patient moves with the largest pairwise beneficial-proportion shifts")
} else {
  render_tbl(empty_note_tbl("No focus-patient beneficial-proportion shift table is available yet."), caption = "Focus patient beneficial-proportion shifts")
}
```

```{r focus-figures, eval=render_figures_use}
if (nrow(focus_landscape_long_tbl)) {
  focus_landscape_density_long_tbl <- focus_landscape_long_tbl %>%
    pivot_longer(cols = c(mean, median), names_to = "metric", values_to = "value") %>%
    mutate(parameter_label = factor(parameter_label, levels = focus_parameter_levels_available))

  p_focus_landscape_density <- ggplot(focus_landscape_density_long_tbl, aes(x = value, color = parameter_label, fill = parameter_label)) +
    geom_density(alpha = 0.18, adjust = 1) +
    facet_wrap(~ metric, scales = "free", ncol = 2) +
    labs(
      title = paste0(focus_pid_use, " landscape distributions by parameter label"),
      subtitle = paste0("minobs = ", focus_minobs_use, ", pm = ", pm_to_label(focus_pm_use)),
      x = "Landscape value",
      y = "Density",
      color = "Parameter label",
      fill = "Parameter label"
    ) +
    theme_bw(base_size = 12)
  print(p_focus_landscape_density)
  ggsave(file.path(figures_dir, paste0("focus_", focus_pid_use, "_landscape_density_by_parameter_label.png")), p_focus_landscape_density, width = 10, height = 5, dpi = 150)
}

if (nrow(focus_parity_tbl)) {
  p_focus_parity <- ggplot(focus_parity_tbl, aes(x = lhs_value, y = rhs_value, color = state_group)) +
    geom_abline(slope = 1, intercept = 0, linetype = 2, color = "grey45") +
    geom_point(alpha = 0.28, size = 0.9) +
    facet_wrap(~ comparison, scales = "free", ncol = 2) +
    scale_color_manual(values = c(fq = "#C53030", nn = "#2B6CB0", other = "#6B7280")) +
    labs(
      title = paste0(focus_pid_use, " pairwise landscape parity plots"),
      subtitle = "Each facet shows matched landscape means on shared karyotypes",
      x = "Left-hand landscape mean",
      y = "Right-hand landscape mean",
      color = "State group"
    ) +
    theme_bw(base_size = 12)
  print(p_focus_parity)
  ggsave(file.path(figures_dir, paste0("focus_", focus_pid_use, "_landscape_parity_by_parameter_label.png")), p_focus_parity, width = 10, height = 8, dpi = 150)
}

if (nrow(focus_beneficial_long_tbl)) {
  p_focus_beneficial <- focus_beneficial_long_tbl %>%
    mutate(parameter_label = factor(parameter_label, levels = focus_parameter_levels_available)) %>%
    ggplot(aes(x = move, y = parameter_label, fill = proportion)) +
    geom_tile(color = "white", linewidth = 0.15) +
    scale_fill_gradient2(
      low = "#2B6CB0",
      mid = "#F7F7F7",
      high = "#C53030",
      midpoint = 0.5,
      limits = c(0, 1),
      na.value = "grey85"
    ) +
    labs(
      title = paste0(focus_pid_use, " beneficial-karyotype proportion by parameter label"),
      subtitle = paste0("minobs = ", focus_minobs_use, ", pm = ", pm_to_label(focus_pm_use)),
      x = "Chromosome move",
      y = NULL,
      fill = "Beneficial\nproportion"
    ) +
    theme_bw(base_size = 11) +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
      panel.grid = element_blank()
    )
  print(p_focus_beneficial)
  ggsave(file.path(figures_dir, paste0("focus_", focus_pid_use, "_beneficial_proportion_heatmap.png")), p_focus_beneficial, width = 14, height = 3.5, dpi = 150)
}
```

## Artifact Index

```{r artifact-index}
artifact_index_tbl <- tibble(
  artifact = c(
    "benchmark_input_index",
    "parameter_spec",
    "parameter_tasks",
    "parameter_fit_results",
    "parameter_all_results",
    "parameter_fit_summary",
    "parameter_selected_fit_index",
    "parameter_beneficial_all_patient_artifacts",
    "parameter_pair_component_summary",
    "parameter_pair_overview",
    "parameter_pair_landscape_group_summary",
    "parameter_pair_sign_summary",
    "parameter_pair_top_landscape_shifts",
    "parameter_global_landscape_overview",
    "parameter_global_landscape_state_summary",
    "focus_parameter_fit_index",
    "focus_landscape_summary",
    "focus_pairwise_component_summary",
    "focus_pairwise_landscape_group_summary",
    "focus_pairwise_top_landscape_shifts",
    "focus_top_variable_karyotypes",
    "focus_beneficial_summary",
    "focus_beneficial_top_shifts",
    "focus_beneficial_proportion_matrix",
    "focus_beneficial_valid_n_matrix",
    "focus_umap_artifacts"
  ),
  path = file.path(
    tables_dir,
    c(
      "benchmark_input_index.tsv",
      "parameter_spec.tsv",
      "parameter_tasks.tsv",
      "parameter_fit_results.tsv",
      "parameter_all_results.tsv",
      "parameter_fit_summary.tsv",
      "parameter_selected_fit_index.tsv",
      "parameter_beneficial_all_patient_artifacts.tsv",
      "parameter_pair_component_summary.tsv",
      "parameter_pair_overview.tsv",
      "parameter_pair_landscape_group_summary.tsv",
      "parameter_pair_sign_summary.tsv",
      "parameter_pair_top_landscape_shifts.tsv",
      "parameter_global_landscape_overview.tsv",
      "parameter_global_landscape_state_summary.tsv",
      "focus_parameter_fit_index.tsv",
      "focus_landscape_summary.tsv",
      "focus_pairwise_component_summary.tsv",
      "focus_pairwise_landscape_group_summary.tsv",
      "focus_pairwise_top_landscape_shifts.tsv",
      "focus_top_variable_karyotypes.tsv",
      "focus_beneficial_summary.tsv",
      "focus_beneficial_top_shifts.tsv",
      "focus_beneficial_proportion_matrix.tsv",
      "focus_beneficial_valid_n_matrix.tsv",
      "focus_umap_artifacts.tsv"
    )
  )
)
save_table_bundle(artifact_index_tbl, file.path(tables_dir, "artifact_index"))

render_tbl(artifact_index_tbl, caption = "Benchmark artifact index")
```
